---
title: 'The center-periphery hypothesis revisited: a meta-analysis'
author: "Jacob Moutouama and Orou Gaoue"
date: "`r Sys.Date()`"
output: 
  rmdformats::readthedown:
    code_folding: hide
    self_contained: true
    number_sections: true
    thumbnails: true
    lightbox: true
    gallery: true
    keep_md: true
    highlight: tango
    df_print: kable 
    toc_depth: 3
    fig_width: 8
    fig_height: 8
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(external = TRUE, echo = F, warning = FALSE, fig.pos = "H", global.device = TRUE)
a4width <- 8.3
a4height <- 11.7
options(knitr.table.format = "latex")
```

This is the code for the paper 'The center-periphery hypothesis revisited: a meta-analysis'.

<style type="text/css">

body{ /* Normal  */
      font-size: 16px;
      font-family: "Times New Roman"
  }
td {  /* Table  */
  font-size: 14px;
}
h1.title {
  font-size: 34px;
  color: Black;
  font-family: "Arial";
}
h1 { /* Header 1 */
  font-size: 26px;
  color: DarkBlue;
  font-family: "Arial";
}
h2 { /* Header 2 */
    font-size: 22px;
    font-family: "Arial";
    color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family: "Arial";
  color: DarkBlue;
}

h4 { /* Header 4 */
  font-size: 16px;
  font-family: "Arial";
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 12px;
}
</style>


**Variables**:\

**eff_size:** Hedges's d and Z transform\
**vi:**      sampling error variance\
**se:**       sampling error\
**mcenter, sdcenter, ncenter:** mean for populations at the range center, standard deviation at the range center, sample size for populations at the range center\
**mperiphery,  sdperiphery,  nperiphery:**  mean for populations at the range periphery, standard deviation at the range periphery, sample size for populations at the range periphery\
**correlation coefficient:**  Relation between population abundance and distance from geographic or niche centroid \
**Years:** Year in which the article was published. 

**Fixed effects:**

1. **study design:**
  + **Method:**  methods that were used to test  the center periphery hypothesis ( geographic distance, euclidean distance, mahanalobis distance)
  + **Kingdom:**  Animal vs Plant
2. **biological traits:**
  + **lifespan:**  the length of time for which an animal exists
  + **Pollination mode:**  Self vs Cross pollinated

**Random effects:**\
**Author**: Author name.\
**Species**  : Identity of species [phylogenetic part].\


```{r eval=TRUE,echo=T}
rm(list = ls(all = TRUE))
```

```{r}
# Load the required packages
library(taxize)
library(ape)
library(phytools)
library(ade4)
library(ggplot2)
library(RColorBrewer)
library(dplyr)
library(tidytree)
library(aplot)
library(cowplot) 
library(ggpubr)
library(ggsci)
library(R.rsp)
library(flextable)
library(effectsize)
library(metafor)
library(esc)
library(knitr)
library(tidyverse)
library(brms)
library(coda)
library(modelr)
library(gridExtra)
library(pBrackets)
library(performance)
library(kableExtra)
library(tidybayes)
library(formattable)
library(grid)
library(rstan)
rstan_options(auto_write = TRUE)
```

Store the Figures and the table 
```{r}
dir.create('/Users/jm200/Library/CloudStorage/Dropbox/PhD Project/Chapter 4/Manuscript/Figures',showWarnings = F)
dir.create('/Users/jm200/Library/CloudStorage/Dropbox/PhD Project/Chapter 4/Manuscript/Tables',showWarnings = F)
```

Function to export the model summary 
```{r}
Ftable <- function(model, reorder = F, order, percent = F){
  table <- summary(model)$fixed
  table2 <-summary(model, prob = 0.9)$fixed %>% as_tibble()
  names <- rownames(table)
  table <- table %>% 
    as_tibble() %>% 
    dplyr::rename('L95%CL' = `l-95% CI`, 'U95%CL' = `u-95% CI`) %>% 
    select(-Rhat, -Bulk_ESS, -Tail_ESS)
  table$`L90%CL` <- table2$`l-90% CI`
  table$`U90%CL` <- table2$`u-90% CI`
  
  table <-table %>% 
    mutate(Estimate = ifelse(`L95%CL`	* `U95%CL` > 0, 
                             paste(sprintf('%.3f', round(Estimate, 3)),  '*', sep = ''), 
                             ifelse(`L90%CL`	* `U90%CL` > 0, 
                                    paste(sprintf('%.3f', round(Estimate, 3)), '\206', sep = ''), round(Estimate, 3))),
           rowname = names) %>% 
    dplyr::rename('SE    ' = Est.Error)  %>% 
    column_to_rownames(var = 'rowname') 

  if (reorder == T){
    table <- table[order,]
  }
  
  table <- table %>% 
    kable(digits = 3, escape = F, table.attr = "style = \"color: black;\"") %>% 
    kable_styling(position = "left")
  return(table)
}


```

Theme to use for all the graphs
```{r}
theme <-  theme(panel.background = element_blank(),
                panel.border = element_rect(fill = NA), 
                strip.background = element_blank(),
                axis.text.x = element_text(size = 10, colour = 'black'),
                axis.text.y = element_text(size = 10, colour = 'black'),
                #text        = element_text(size = 8),
                axis.title.y = element_text(size = 10),# size of y lab
                axis.title.x = element_text(size = 10),# size of X lab
                legend.position     = c(0.1,0.8),
                legend.title        = element_text(size = 9),
                legend.text         = element_text(size = 8),
                legend.key.height   = unit(0.8,'line'),
                legend.background   = element_blank(),
                legend.key          = element_rect(colour = NA, fill = NA),
                plot.title  = element_text(hjust = 0.5),
                plot.margin = margin(t = 0.5, r = 0.5, b = 0.5, l = 0.5, unit = "cm"))
```

# What is the magnitude of the cumulative effect size across studies for CPH support, and is it significantly different from zero?  

## Population growth rate and CPH

```{r}
Lambdacp <- read.csv("/Users/jm200/Library/CloudStorage/Dropbox/PhD Project/Chapter 4/Manuscript/Data/Lambda.csv", header = T)
```

### Standardized Mean Difference: Hedges’s d (SDM)

```{r}
datlambdacp <- escalc("SMDH", m1i = mcenter, m2i = mperiphery, sd1i = sdcenter, sd2i = sdperiphery, n1i = ncenter, n2i = nperiphery, slab = Author, data = Lambdacp)
Lambdar <- read.csv("/Users/jm200/Library/CloudStorage/Dropbox/PhD Project/Chapter 4/Manuscript/Data/Lambdar.csv", header = T)
datlambda<-rbind(datlambdacp,Lambdar)
datlambda$se<-sqrt(datlambda$vi)
# write.csv(datlambda,file = "/Users/jm200/Library/CloudStorage/Dropbox/PhD Project/Chapter 4/Manuscript/Data/Figshare/Lambda.csv")
```

### Phylogeny tree 

```{r,eval=FALSE}
sp.lambda <- datlambda[, 4]
spnames <- as.character(sp.lambda)
species1 <- unique(spnames)
# length(species1)
# species2 <- gsub("\n", "", species1)
species2 <- gsub("\xca", " ", species1) 
temp <- gnr_resolve(species2, best_match_only = TRUE, canonical = TRUE)
dat <- merge(data.frame(species2), temp[,c("user_supplied_name", "matched_name2")], 
               by.x = "species2", by.y = "user_supplied_name", all.x = TRUE)
# Retrieve classifications)
resolved_names <- classification(dat[, 2], db = "ncbi")
resolved_names <- unique(resolved_names)
lapply(resolved_names, head)
class(resolved_names)
# Building a phylogenetic tree
lambdatree <- class2tree(resolved_names)
(lambdaphy <- lambdatree$phylo)
```

### Computes the expected variances and covariances of branch length 

```{r}
datlambda$Species<-gsub("\xca", " ", datlambda$Species) 
lambdaphyclean<-drop.tip(lambdaphy, lambdaphy$tip.label[-na.omit(match(datlambda$Species, lambdaphy$tip.label))])
A<-ape::vcv.phylo(lambdaphy)
```

### Grand Mean effect size and significance 

```{r}
Lambdamodel <- brm(
  yi | se(se, sigma = TRUE) ~ 1  +  (1 | Author) + (1|gr(Species, cov = A)),
  data = datlambda[-c(11,18),],
  family = gaussian(),
  control = list(adapt_delta = 0.9999, max_treedepth = 15),
                   iter = 3000, cores = 4,
  data2 = list(A = A)
)
```

### Predictive check for population growth rate

```{r}
pp_check(Lambdamodel)+
  xlab("Growth rate status")+ylab("Density")+
  ggtitle((""))+theme->ppc_growthrate
```

### MCMC Mixing

```{r}
plot(Lambdamodel)
```

### Value for Grand Mean effect size

```{r}
Lambdamodel %>% Ftable(percent = T)
summary(Lambdamodel)
```

### Plot for Grand Mean effect size

```{r}
# Average effect
out_Lambda <- spread_draws(Lambdamodel, b_Intercept) %>% 
  mutate(study = "Population\ngrowth rate")
# Data frame of summary numbers
out_all_sum_lambda <- group_by(out_Lambda, study) %>% 
  mean_qi(b_Intercept)
# Draw plot
Fig1a<-out_Lambda %>%   
  ggplot(aes(b_Intercept, study)) +
  labs(x = "", y = '')+
  xlim(-2,2)+
  # Zero!
  geom_vline(xintercept = 0, linewidth = .25, lty = 2) +
  stat_halfeye(.width = c(.8, .95), fill = "#FFC34D",alpha = 0.9) +
  # Add text labels
  geom_text(
    data = mutate_if(out_all_sum_lambda, is.numeric, round, 2),
    aes(label = str_glue("{b_Intercept} [{.lower}, {.upper}]"), x =2.1,y=1.05),
    hjust = "inward"
  ) +
  theme
```

### Heterogeneity for population growth rate

```{r}
dathlambda<-datlambda[-c(11,18),]
var_lambda <- sum(1/dathlambda$vi)*(nrow(dathlambda)-1)/((sum(1/dathlambda$vi))^2-sum((1/dathlambda$vi)^2))
var_samp_lambda <- Lambdamodel %>% 
  posterior_samples() %>% 
  select(starts_with('sd'), sigma) %>% 
  transmute_all(.funs = list(sq = ~(. ^ 2))) %>% # sd to variance
  mutate(sampling = var_lambda) %>%  # sampling variance
  mutate(total_var = rowSums(.),
         no_sample = total_var - sampling)
colnames(var_samp_lambda) <- colnames(var_samp_lambda) %>% 
  str_remove_all('__Intercept|_sq|sd_')
```

```{r}
# I^2 and lambda
I_square_lambda <- hypothesis(var_samp_lambda, c('Author/total_var= 0',
                                   'Species/total_var = 0',
                                   'sigma/total_var = 0',
                                   'Species/no_sample = 0'))$hypothesis %>% 
  as_tibble() %>% 
  select(-Evid.Ratio, -Post.Prob, -Star)

I_square_lambda[,1] <- c('Author', 'Species', 'sigma','Phylogenetic signal')
colnames(I_square_lambda) <- c('', 'Estimate', 'SE', 'L95%CL', 'U95%CL')
I_square_lambda <- I_square_lambda %>% 
  kable(digits = 4, escape = F, table.attr = "style = \"color: black;\"") %>% 
  kable_styling()
I_square_lambda
```

## Genetics and CPH

```{r}
gencp <- read.csv("/Users/jm200/Library/CloudStorage/Dropbox/PhD Project/Chapter 4/Manuscript/Data/genetic.csv", header = T)
genr <- read.csv("/Users/jm200/Library/CloudStorage/Dropbox/PhD Project/Chapter 4/Manuscript/Data/geneticr.csv", header = T)

```

### Standardized Mean Difference: Hedges’s d (SDM)

```{r}
datgen <- escalc("SMDH", m1i = mcenter, m2i = mperiphery, sd1i = sdcenter, sd2i = sdperiphery, n1i = ncenter, n2i = nperiphery, slab = Author, data = gencp)
datgencp<-rbind(datgen,genr)
datgencp <- datgen[!apply(datgen[,c("yi", "vi")], 1, anyNA),] 
datgencp$se<-sqrt(datgencp$vi)
write.csv(datgencp,file = "/Users/jm200/Library/CloudStorage/Dropbox/PhD Project/Chapter 4/Manuscript/Data/Figshare/Genetic.csv")
```

### Grand Mean effect size and significance 

```{r, warning=FALSE}
genamodel <- brm(
  yi | se(se, sigma = TRUE) ~ 1 + (1 | Author) + (1|Species),
  control = list(adapt_delta = 0.9999, max_treedepth = 20),
                   iter = 5000, cores = 4,
  data = datgencp
)
```

### Predictive check for genetic diversity

```{r}
pp_check(genamodel)+
  xlab("Genetic diversity status")+ylab("Density")+
  ggtitle((""))+theme->ppc_genetic
```

### MCMC Mixing

```{r}
plot(genamodel)
```

### Value for Grand Mean effect size for genetic diversity

```{r}
genamodel %>% Ftable(percent = T)
```

# Heterogeinity for genetic diversity 

```{r}
var_gen <- sum(1/datgencp$vi)*(nrow(datgencp)-1)/((sum(1/datgencp$vi))^2-sum((1/datgencp$vi)^2))
var_samp_gen <- genamodel %>% 
  posterior_samples() %>% 
  select(starts_with('sd'), sigma) %>% 
  transmute_all(.funs = list(sq = ~(. ^ 2))) %>% # sd to variance
  mutate(sampling = var_gen) %>%  # sampling variance
  mutate(total_var = rowSums(.),
         no_sample = total_var - sampling)
colnames(var_samp_gen) <- colnames(var_samp_gen) %>% 
  str_remove_all('__Intercept|_sq|sd_')
```

```{r}
# I^2 and lambda
I_square_gen <- hypothesis(var_samp_gen, c('Author/total_var= 0',
                                   'Species/total_var = 0',
                                   'sigma/total_var = 0',
                                   'Species/no_sample = 0'))$hypothesis %>% 
  as_tibble() %>% 
  select(-Evid.Ratio, -Post.Prob, -Star)

I_square_gen[,1] <- c('Author', 'Species', 'sigma','Phylogenetic signal')
colnames(I_square_gen) <- c('', 'Estimate', 'SE', 'L95%CL', 'U95%CL')
I_square_gen <- I_square_gen %>% 
  kable(digits = 4, escape = F, table.attr = "style = \"color: black;\"") %>% 
  kable_styling()
I_square_gen
```

### Plot for Grand Mean effect size

```{r}
# Average effect
out_f_gen <- spread_draws(genamodel, b_Intercept) %>% 
  mutate(study = "Genetic \n diversity")
# Data frame of summary numbers
out_all_sum_gen <- group_by(out_f_gen, study) %>% 
  mean_qi(b_Intercept)
# Draw plot
Fig1b<-out_f_gen %>%   
  ggplot(aes(b_Intercept, study)) +
  labs(x = "", y = '')+
  xlim(-2,2)+
  # Zero!
  geom_vline(xintercept = 0, linewidth = .25, lty = 2) +
  stat_halfeye(.width = c(.8, .95), fill = "#FFC34D",alpha = 0.9) +
  # Add text labels
  geom_text(
    data = mutate_if(out_all_sum_gen, is.numeric, round, 2),
    aes(label = str_glue("{b_Intercept} [{.lower}, {.upper}]"), x =1.6),
    hjust = "inward"
  ) +
  theme
  
```

## Population abundance and CPH

```{r}
Abundcp <- read.csv("/Users/jm200/Library/CloudStorage/Dropbox/PhD Project/Chapter 4/Manuscript/Data/Abundance.csv", header = T)
Abundcp$Species<-gsub("spp.","",as.character(Abundcp$Species))
Abundcp$Species<-gsub("sp.","",as.character(Abundcp$Species))
```

### Fisher’s Z-transformed-r

```{r}
databun <- escalc("ZCOR", ri = correlation.coefficient, ni = sample_size, data = Abundcp)
Abundcp <- read.csv("/Users/jm200/Library/CloudStorage/Dropbox/PhD Project/Chapter 4/Manuscript/Data/Abundancez.csv", header = T)
databuncp<-rbind(databun,Abundcp)
databuncp<-databun[!apply(databun[,c("yi", "vi")], 1, anyNA),] 
databuncp$se<-sqrt(databuncp$vi)
write.csv(databuncp,file = "/Users/jm200/Library/CloudStorage/Dropbox/PhD Project/Chapter 4/Manuscript/Data/Figshare/Abundance.csv")
```

### Grand Mean effect size and significance 

```{r}
abunmodel <- brm(
  yi | se(se, sigma = TRUE) ~ 1 + (1 | Author) + (1|Species),
  data = databuncp
)
```

### Predictive check for population abundance 

```{r}
pp_check(abunmodel)+
  xlab("Abundance status")+ylab("Density")+
  ggtitle((""))+theme->ppc_abundance
```

```{r}
plot(abunmodel)
```

```{r}
abunmodel %>% Ftable(percent = T)
```

```{r}
# Average effect
out_f_abun <- spread_draws(abunmodel, b_Intercept) %>% 
  mutate(study = "Abundance")
# Data frame of summary numbers
out_all_sum_abun <- group_by(out_f_abun, study) %>% 
  mean_qi(b_Intercept)
# Draw plot
Fig1c<-out_f_abun %>%   
  ggplot(aes(b_Intercept, study)) +
  labs(x = "", y = '')+
  xlim(-2,2)+
  # Zero!
  geom_vline(xintercept = 0, linewidth = .25, lty = 2) +
  stat_halfeye(.width = c(.8, .95), fill = "#FFC34D",alpha = 0.9) +
  # Add text labels
  geom_text(
    data = mutate_if(out_all_sum_abun, is.numeric, round, 2),
    aes(label = str_glue("{b_Intercept} [{.lower}, {.upper}]"), x =1.6),
    hjust = "inward"
  ) +
  theme
```

### Heterogeinity for abundance data
```{r}
var_abun <- sum(1/databuncp$vi)*(nrow(databuncp)-1)/((sum(1/databuncp$vi))^2-sum((1/databuncp$vi)^2))
var_samp_abun <- abunmodel %>% 
  posterior_samples() %>% 
  select(starts_with('sd'), sigma) %>% 
  transmute_all(.funs = list(sq = ~(. ^ 2))) %>% # sd to variance
  mutate(sampling = var_abun) %>%  # sampling variance
  mutate(total_var = rowSums(.),
         no_sample = total_var - sampling)
colnames(var_samp_abun) <- colnames(var_samp_abun) %>% 
  str_remove_all('__Intercept|_sq|sd_')
```

```{r}
# I^2 and lambda
I_square_abun <- hypothesis(var_samp_abun, c('Author/total_var= 0',
                                   'Species/total_var = 0',
                                   'sigma/total_var = 0',
                                   'Species/no_sample = 0'))$hypothesis %>% 
  as_tibble() %>% 
  select(-Evid.Ratio, -Post.Prob, -Star)

I_square_abun[,1] <- c('Author', 'Species', 'sigma','Phylogenetic signal')
colnames(I_square_abun) <- c('', 'Estimate', 'SE', 'L95%CL', 'U95%CL')
I_square_abun <- I_square_abun %>% 
  kable(digits = 4, escape = F, table.attr = "style = \"color: black;\"") %>% 
  kable_styling()
I_square_abun
```

## Testing if adding data from Dallas et al , 2017  improve the results

```{r}
AbundcpDallas <- read.csv("/Users/jm200/Library/CloudStorage/Dropbox/PhD Project/Chapter 4/Manuscript/Data/Abundance_withoutDallas.csv", header = T)
AbundcpDallas$Species<-gsub("spp.","",as.character(AbundcpDallas$Species))
AbundcpDallas$Species<-gsub("sp.","",as.character(AbundcpDallas$Species))
```

### Fisher’s Z-transformed-r

```{r}
databunDallas <- escalc("ZCOR", ri = correlation.coefficient, ni = sample_size, data = AbundcpDallas)
AbundcpDallas <- read.csv("/Users/jm200/Library/CloudStorage/Dropbox/PhD Project/Chapter 4/Data/Data/Abundancez.csv", header = T)
databuncpDallas<-rbind(databunDallas,AbundcpDallas)
databuncpDallas<-databun[!apply(databun[,c("yi", "vi")], 1, anyNA),] 
databuncpDallas$se<-sqrt(databuncpDallas$vi)
# write.csv(databuncpDallas,file = "/Users/jm200/Library/CloudStorage/Dropbox/PhD Project/Chapter 4/Manuscript/Data/Abundance_WithoutDallas.csv")
```

### Grand Mean effect size and significance 

```{r, warning=FALSE}
abunmodelDallas <- brm(
  yi | se(se, sigma = TRUE) ~ 1 + (1 | Author) + (1|Species),
  data = databuncpDallas
)
```


```{r}
pp_check(abunmodelDallas)
```

```{r}
plot(abunmodelDallas)
```

```{r}
abunmodelDallas %>% Ftable(percent = T)
```

### Predictive check for all the models (abundance, population growth rate, population abundance)

```{r}
library(scater)
# print figure
pdf("/Users/jm200/Library/CloudStorage/Dropbox/PhD Project/Chapter 4/Manuscript/Figures/PPC.pdf",useDingbats = F,height=7,width=6)
multiplot(ppc_growthrate,ppc_genetic,
          ppc_abundance,cols=1)
dev.off()
```

### Plot for Grand Mean effect size for abundance, genetic diversity and population growth rate

```{r,eval=FALSE}
# Average effect
out_f <- spread_draws(abunmodelDallas, b_Intercept) %>% 
  mutate(study = "Abundance")
# Data frame of summary numbers
out_all_sum <- group_by(out_f, study) %>% 
  mean_qi(b_Intercept)
# Draw plot
Fig1Dallas<-out_f %>%   
  ggplot(aes(b_Intercept, study)) +
  labs(x = "Effect size", y = '')+
  # Zero!
  geom_vline(xintercept = 0, linewidth = .25, lty = 2) +
  stat_halfeye(.width = c(.8, .95), fill = "#FFC34D",alpha = 0.9) +
  # Add text labels
  geom_text(
    data = mutate_if(out_all_sum, is.numeric, round, 2),
    aes(label = str_glue("{b_Intercept} [{.lower}, {.upper}]"), x =1),
    hjust = "inward"
  ) +
  theme
```

```{r}
out_all <- bind_rows(out_Lambda,out_f_abun, out_f_gen) 
out_all_sum <-bind_rows(out_all_sum_lambda, out_all_sum_abun,out_all_sum_gen) 

Figure1<-out_all %>%   
  ggplot(aes(b_Intercept, study)) +
  labs(x = "Effect size", y = '')+
  xlim(-1.8,2)+
  # Zero!
  geom_vline(xintercept = 0, linewidth = .25, lty = 2) +
  stat_halfeye(.width = c(.8, .95), fill = "#FFC34D",alpha = 0.9) +
  # Add text labels
  geom_text(size = 2.8,
    data = mutate_if(out_all_sum, is.numeric, round, 2),
    aes(label = str_glue("{b_Intercept} [{.lower}, {.upper}]"), x =-1.5),
    hjust = "inward"
  ) +
  theme

```

# Detection of publication bias
## Population growth rate 

```{r}
fittedlambda <- fitted(Lambdamodel, dathlambda)
dathlambda$meta_resid <- dathlambda$yi  - fittedlambda[,1]
dathlambda <- dathlambda %>% 
  mutate(p_i = 1/se,
         o_i = meta_resid * p_i)
rm(fittedlambda)
```

egger's regression
```{r, eval =F}
m_egg_lambda <- brm(data = dathlambda, o_i ~ p_i, control = list(adapt_delta = 0.999), cores = 4)
```

```{r}
m_egg_lambda %>% Ftable
```

effect size ~ year
```{r, eval =F}
m_year_lambda <- update(Lambdamodel, .~. + Year, newdata = dathlambda, cores = 4)
```

```{r}
m_year_lambda %>% Ftable
```

```{r}
CorMatrixLambda<-vcv(lambdaphy, corr=TRUE)
Lambdamodelpb <- rma.mv(yi = yi , V = vi, method = "REML", 
                   test="t", dfs="contain",
                   random=list(~1|Author,  ~1|Species), 
             R=list(full_sp=CorMatrixLambda), data = dathlambda)

```

## Population genetic

```{r}
fitted_gen <- fitted(genamodel, datgencp)
datgencp$meta_resid <- datgencp$yi  - fitted_gen[,1]
datgencp <- datgencp %>% 
  mutate(p_i = 1/se,
         o_i = meta_resid * p_i)
```

```{r}
rm(fitted_gen)
```

egger's regression
```{r, eval =F}
m_egg_gen <- brm(data = datgencp, o_i ~ p_i, control = list(adapt_delta = 0.999), cores = 4)
```

```{r}
m_egg_gen %>% Ftable
```

effect size ~ year
```{r, eval =F}
m_year_gen <- update(genamodel, .~. + Year, newdata = datgencp, cores = 4)
```

```{r}
m_year_gen %>% Ftable
```

```{r}
genamodelpb <- rma.mv(yi = yi , V = vi, method = "REML", 
                   test="t", dfs="contain",
                   random=list(~1|Author,  ~1|Species),  data = datgencp)
```

## Population abundance

```{r}
fitted_abun <- fitted(abunmodel, databuncp)
databuncp$meta_resid <- databuncp$yi  - fitted_abun[,1]
databuncp <- databuncp %>% 
  mutate(p_i = 1/se,
         o_i = meta_resid * p_i)
rm(fitted_abun)
```

egger's regression
```{r, eval =F}
m_egg_abun <- brm(data =databuncp, o_i ~ p_i, cores = 4)
```

```{r}
m_egg_abun %>% Ftable
```

effect size ~ year
```{r, eval =F}
m_year_abun <- update(abunmodel, .~. + Year, newdata = databuncp, cores = 4)
```
```{r}
m_year_abun %>% Ftable
```

```{r}
abunmodelpb <- rma.mv(yi = yi , V = vi, method = "REML", 
                   test="t", dfs="contain",
                   random=list(~1|Author,  ~1|Species),  data = databuncp)

```

## Put all the funnels plot together
```{r}
pdf("/Users/jm200/Library/CloudStorage/Dropbox/PhD Project/Chapter 4/Manuscript/Figures/Fig2.pdf", width = 7, height = 4)
par(mfrow=c(1,3))
a <- funnel(Lambdamodelpb, 
       yaxis="seinv", # Inverse of standard error (precision) as the y axis
       level = c(90, 95, 99),  # levels of statistical significance highlighted 
       shade = c("white", "gray55", "gray75"), 
       ylab="Precision (1/SE)",
       xlab= "Population growth rate",
       cex.lab=1.2, 
       col="orange2",
        back="white",
       digits=1, 
       xlim=c(-4,4), 
       ylim=c(0.01,75))

b <- funnel(genamodelpb, 
       yaxis="seinv", # Inverse of standard error (precision) as the y axis
       level = c(90, 95, 99),  # levels of statistical significance highlighted 
       shade = c("white", "gray55", "gray75"), 
       ylab="",
       xlab="Genetic diversity",
       cex.lab=1.2, 
       col="orange2",
       back="white",
       digits=1, 
       xlim=c(-4,4), 
       ylim=c(0.01,75))

c <- funnel(abunmodelpb, 
       yaxis="seinv", # Inverse of standard error (precision) as the y axis
       level = c(90, 95, 99),  # levels of statistical significance highlighted 
       shade = c("white", "gray55", "gray75"), 
       ylab="",
       xlab="Population abundance",
       col="orange2",
       back="white",
       cex.lab=1.2, 
       digits=1, 
       xlim=c(-4,4), 
       ylim=c(0.01,75))

dev.off()
```

# To what extent can heterogeneity among studies be explained by hypothesized moderators that characterize study outcomes?

## Kingdom

```{r}
Kingdomg.model <- brms::brm(
  yi|se(se, sigma = TRUE) ~ (Kingdom-1) + (1|Author) + (1|Species),
  data = datgencp,
  family = gaussian(),
  prior = c(
    prior(normal(0, 1), "b"),
    prior(student_t(3, 0, 5), "sd"),
    prior(student_t(3, 0, 5), "sigma")
  ),
  control = list(adapt_delta = 0.9999, max_treedepth = 15),
                   iter = 5000, cores = 4
)
```

```{r}
summary(Kingdomg.model)
```

```{r}
library(rstan)
Kingdomab.model <- brms::brm(
  yi|se(se, sigma = TRUE) ~ (Kingdom-1) + (1|Author) + (1|Species),
  data = databuncp,
  family = gaussian(),
  prior = c(
    prior(normal(0, 1), "b"),
    prior(student_t(3, 0, 5), "sd"),
    prior(student_t(3, 0, 5), "sigma")
  ),
  control = list(adapt_delta = 0.9999, max_treedepth = 15),
                   iter = 5000, cores = 4
)
```

```{r}
summary(Kingdomab.model)
```

```{r}
Kingdomme.model <- brms::brm(
  yi|se(se, sigma = TRUE) ~ (Method-1) + (1|Author) + (1|Species),
  data = databuncp,
  family = gaussian(),
  prior = c(
    prior(normal(0, 1), "b"),
    prior(student_t(3, 0, 5), "sd"),
    prior(student_t(3, 0, 5), "sigma")
  ),
  control = list(adapt_delta = 0.9999, max_treedepth = 15),
                   iter = 5000, cores = 4
)
```

```{r}
summary(Kingdomme.model)
```

```{r}
themepollination <-  theme(panel.background = element_blank(),
                panel.border = element_rect(fill = NA), 
                strip.background = element_blank(),
                axis.text.x = element_text(size = 8, colour = 'black'),
                axis.text.y = element_text(size = 8, colour = 'black', hjust = 0.5, angle = 90),
                #text        = element_text(size = 8),
                axis.title.y = element_text(size = 8),# size of y lab
                axis.title.x = element_text(size = 8),# size of X lab
                legend.position     = c(0.1,0.8),
                legend.title        = element_text(size = 9),
                legend.text         = element_text(size = 8),
                legend.key.height   = unit(0.8,'line'),
                legend.background   = element_blank(),
                legend.key          = element_rect(colour = NA, fill = NA),
                plot.title  = element_text(hjust = 0.5),
                plot.margin = margin(t = 0.5, r = 0.5, b = 0.5, l = 0.5, unit = "cm"))
```

### Match the effectif size with the pollination data
Pollination and population growth rate
```{r}
Pollinationdata <- read.csv("/Users/jm200/Library/CloudStorage/Dropbox/PhD Project/Chapter 4/Manuscript/Data/Pollinationinfo.csv", header = T)
Pollinationdata$Pollination <- gsub(" ", "",Pollinationdata$Pollination)
datlambdaPl<-left_join(x=datlambda,y=Pollinationdata,by="Species")
datlambdaPl <- datlambdaPl[!apply(datlambdaPl[,c("Pollination", "vi")], 1, anyNA),] 
```

```{r}
Pollination.model <- brms::brm(
  yi|se(se, sigma = TRUE) ~ (Pollination-1) + (1|Author) + (1|gr(Species, cov = A)),
  data = datlambdaPl,
  family = gaussian(),
  data2 = list(A = A),
  prior = c(
    prior(normal(0, 1), "b"),
    prior(student_t(3, 0, 5), "sd"),
    prior(student_t(3, 0, 5), "sigma")
  ),
  control = list(adapt_delta = 0.9999, max_treedepth = 15),
                   iter = 5000, cores = 4
)
  
```

```{r}
plot(Pollination.model)

```

```{r}
summary(Pollination.model)
# fixef(Pollination.model)
```

```{r}
tidybayes::get_variables(Pollination.model)

```

```{r}
out_f_pollambda1 <- spread_draws(Pollination.model, b_Pollinationout) %>% 
  mutate(study = "Crossed\n pollinated") %>% 
  dplyr::rename (Pollination=b_Pollinationout)
# Data frame of summary numbers
out_all_sum_pollambda1 <- group_by(out_f_pollambda1, study) %>% 
  mean_qi(Pollination)



out_f_pollambda2 <- spread_draws(Pollination.model, b_Pollinationself) %>% 
  mutate(study = "Self\n pollinated") %>% 
  dplyr::rename (Pollination=b_Pollinationself)
# Data frame of summary numbers
out_all_sum_pollambda2 <- group_by(out_f_pollambda2, study) %>% 
  mean_qi(Pollination)

out_allpol_lambda <- bind_rows(out_f_pollambda1, out_f_pollambda2) 
out_all_sumpollmbda <-bind_rows(out_all_sum_pollambda1,out_all_sum_pollambda2) 

Figure5a<-out_allpol_lambda %>%   
  ggplot(aes(Pollination, study)) +
  labs(x = "Effect size (population \n growth rate)", y = '')+
  xlim(-3.5,3.5)+
  # Zero!
  geom_vline(xintercept = 0, linewidth = .25, lty = 2) +
  stat_halfeye(.width = c(.8, .95), fill = "#31688e",alpha = 0.9) +
  # Add text labels
  geom_text(size = 2.8,
    data = mutate_if(out_all_sumpollmbda, is.numeric, round, 2),
    aes(label = str_glue("{Pollination} [{.lower}, {.upper}]"), x =-3.4),
    hjust = "inward"
  ) +
  annotate("text", x =-3,y=2.5, label = "a", color = "black",  size = 5) +
  theme(panel.border = element_rect(fill = NA),
             text = element_text(size = 10), # change font sizes
                 legend.position = "none",
                 legend.title = element_text(size = 8),
                 legend.text = element_text(size = 8),
                 legend.background = element_blank(),
                 panel.background = element_rect(fill="transparent"), 
                 axis.text.x = element_text(size = 10, colour = "black"),
                 axis.text.y = element_text(size = 10, colour = "black", hjust = 0.5, angle = 90))

```
Pollination and genetic diversity 
```{r}
datgencp %>% 
  filter(Kingdom == "Plant")->datgenp
gencpPL<-left_join(x =datgenp, y =Pollinationdata ,by="Species")
```


```{r,eval=FALSE}
sp.genp <- gencpPL[, 4]
spnamesp <- as.character(sp.genp)
# Retrieve classifications
resolved_namesp <- classification(spnamesp, db = "ncbi")
resolved_namesp <- unique(resolved_namesp)
# lapply(resolved_namesp, head)
# Building a phylogenetic tree
genptree <- class2tree(resolved_namesp)
#str(gentree)
(genpphy <- genptree$phylo)
# plot(genpphy)
```


```{r}
speciesgen<-data.frame(Species=genpphy$tip.label)
genpdat<- merge(x = speciesgen, y = gencpPL,all.x = TRUE)
genpdat<-genpdat[complete.cases(genpdat), ]
genpcleanTree<-drop.tip(genpphy, genpphy$tip.label[-na.omit(match(genpdat$Species, genpphy$tip.label))])
A<-ape::vcv.phylo(genpcleanTree) # Creates variance-covariance matrix from tree
```

```{r}
set.seed(13)
Pollination.genmodel <- brms::brm(
  yi|se(se, sigma = TRUE) ~ (Pollination-1) + (1|Author) + (1|gr(Species, cov = A)),
  data = genpdat,
  family = gaussian(),
  data2 = list(A = A),
  prior = c(
    prior(normal(0, 1), "b"),
    prior(student_t(3, 0, 3), "sd"),
    prior(student_t(3, 0, 3), "sigma")
  ),
  control = list(adapt_delta = 0.9999, max_treedepth = 15),
                   iter = 5000, cores = 4
)
  
```

```{r}
plot(Pollination.genmodel)
summary(Pollination.genmodel)
fixef(Pollination.genmodel)
```

```{r}
out_f_polgen1 <- spread_draws(Pollination.genmodel, b_Pollinationout) %>% 
  mutate(study = "Crossed\n pollinated") %>% 
  dplyr::rename (Pollination=b_Pollinationout)
# Data frame of summary numbers
out_all_sum_polgen1 <- group_by(out_f_polgen1, study) %>% 
  mean_qi(Pollination)



out_f_polgen2 <- spread_draws(Pollination.genmodel, b_Pollinationself) %>% 
  mutate(study = "Self\n pollinated") %>% 
  dplyr::rename (Pollination=b_Pollinationself)
# Data frame of summary numbers
out_all_sum_polgen2 <- group_by(out_f_polgen2, study) %>% 
  mean_qi(Pollination)

out_allpol_gen <- bind_rows(out_f_polgen1, out_f_polgen2) 
out_all_sumpolgen <-bind_rows(out_all_sum_polgen1,out_all_sum_polgen2) 

Figure5b<-out_allpol_gen %>%   
  ggplot(aes(Pollination, study)) +
  labs(x = "Effect size (Genetic diversity)", y = "")+
   xlim(-5,4)+
  # Zero!
  geom_vline(xintercept = 0, linewidth = .25, lty = 2) +
  stat_halfeye(.width = c(.8, .95), fill = "#31688e",alpha = 0.9) +
  # Add text labels
  geom_text(size = 2.8,
    data = mutate_if(out_all_sumpolgen, is.numeric, round, 2),
    aes(label = str_glue("{Pollination} [{.lower}, {.upper}]"), x =-4.9),
    hjust = "inward"
  ) +
  annotate("text", label = "b", x =-4,y=2.5, size = 5)+
  theme(panel.border = element_rect( fill=NA),
             text = element_text(size = 10), # change font sizes
                 legend.position = "none",
                 legend.title = element_text(size = 8),
                 legend.text = element_text(size = 8),
                 legend.background = element_blank(),
                 panel.background = element_rect(fill="transparent"), 
                 axis.text.x = element_text(size = 10, colour = "black"),
                 axis.text.y = element_text(size = 10, colour = "white", hjust = 0.5, angle = 90))

```
Pollination and plant abundance

```{r}
databuncp %>%
  filter(Kingdom == "Plant")->databp
databuncpPl<-left_join(x=databp,y=Pollinationdata,by="Species")
databuncpPl$Pollination <- gsub("Self", "self",databuncpPl$Pollination)
# unique(databuncpPl$Pollination)
databuncpPl <- databuncpPl[!apply(databuncpPl[,c("Pollination", "vi")], 1, anyNA),] 
traitslamda<-tr8(species_list=databp$Species,download_list=traits) 
```

```{r,eval=FALSE}
sp.abp <- databuncpPl[, 4]
spnamesabp <- as.character(sp.abp)
species2 <- spnamesabp
#length(species2)
# species2 <- gsub("\n", "", species1)
temp <-taxize::gnr_resolve(species2, best_match_only = TRUE, canonical = TRUE)
dat_abp <- merge(data.frame(species2), temp[,c("user_supplied_name", "matched_name2")], 
               by.x = "species2", by.y = "user_supplied_name", all.x = TRUE)
spnamesabp <-dat_abp[,2]
# spnamesabp <- spnamesabp[complete.cases(spnamesabp)]
# length(spnamesabp)
# Retrieve classifications
spnamesabpf <- classification(spnamesabp, db = "ncbi")
resolved_names_abp <- unique(spnamesabpf)
# Building a phylogenetic tree
abptree <- class2tree(resolved_names_abp)
# str(abptree)
(abpphy <- abptree$phylo)
# length(abpphy$edge.length)
```

```{r}
speciesabunp<-data.frame(Species=abpphy$tip.label)
abunpdat<- merge(x = speciesabunp, y = databuncpPl,all.x = TRUE)
abunpdat<-abunpdat[complete.cases(abunpdat), ]
abunpcleanTree<-drop.tip(abpphy, abpphy$tip.label[-na.omit(match(abunpdat$Species, abpphy$tip.label))])
A<-ape::vcv.phylo(abunpcleanTree) # Creates variance-covariance matrix from tree
```

```{r}
Pollination.abunmodel <- brms::brm(
  yi|se(se, sigma = TRUE) ~ (Pollination-1) + (1|Author) + (1|gr(Species, cov = A)),
  data = abunpdat,
  family = gaussian(),
  data2 = list(A = A),
  prior = c(
    prior(normal(0, 1), "b"),
    prior(student_t(3, 0, 1.5), "sd"),
    prior(student_t(3, 0, 1.5), "sigma")
  ), 
  control = list(adapt_delta = 0.9999, max_treedepth = 15),iter = 5000, cores = 4
)
  
```

```{r}
plot(Pollination.abunmodel)
fixef(Pollination.abunmodel)
```

```{r}
out_f_polabun1 <- spread_draws(Pollination.abunmodel, b_Pollinationout) %>% 
  mutate(study = "Crossed\n pollinated") %>% 
  dplyr::rename (Pollination=b_Pollinationout)
# Data frame of summary numbers
out_all_sum_polabund1 <- group_by(out_f_polabun1, study) %>% 
  mean_qi(Pollination)
out_f_polabund2 <- spread_draws(Pollination.abunmodel, b_Pollinationself) %>% 
  mutate(study = "Self\n pollinated") %>% 
  dplyr::rename (Pollination=b_Pollinationself)
# Data frame of summary numbers
out_all_sum_polabun2 <- group_by(out_f_polabund2, study) %>% 
  mean_qi(Pollination)

out_allpol_abun <- bind_rows(out_f_polabun1, out_f_polabund2) 
out_all_sumpolabun <-bind_rows(out_all_sum_polabund1,out_all_sum_polabun2) 

Figure4c<-out_allpol_abun %>%   
  ggplot(aes(Pollination, study)) +
  labs(x = "Effect size ( abundance)", y = '')+
   xlim(-3,3)+
  # Zero!
  geom_vline(xintercept = 0, linewidth = .25, lty = 2) +
  stat_halfeye(.width = c(.8, .95), fill = "#31688e",alpha = 0.9) +
  # Add text labels
  geom_text(size = 2.8,
    data = mutate_if(out_all_sumpolabun, is.numeric, round, 2),
    aes(label = str_glue("{Pollination} [{.lower}, {.upper}]"), x =-2.5),
    hjust = "inward"
  ) +
  annotate("text", label = "c", x =-2.5,y=2.5, size = 5)+
  theme(panel.border = element_rect(fill=NA),
             text = element_text(size = 10), # change font sizes
                 legend.position = "none",
                 legend.title = element_text(size = 8),
                 legend.text = element_text(size = 8),
                 legend.background = element_blank(),
                 panel.background = element_rect(fill="transparent"), 
                 axis.text.x = element_text(size = 10, colour = "black"),
                 axis.text.y = element_text(size = 10, colour = "black", hjust = 0.5, angle = 90))



```

```{r}
pdf("/Users/jm200/Library/CloudStorage/Dropbox/PhD Project/Chapter 4/Manuscript/Figures/Fig5.pdf", width = 8.5, height = 6)
(Fig4 <- ggarrange(Figure5a,Figure5b,Figure5c,
  hjust = -1,
  font.label = list(size = 14, color = "black",
                    face = "plain", family = NULL),
  vjust = 1.5, widths = c(2, 2), heights = 0.5, align = "hv",
  ncol = 2, nrow = 2
))

dev.off()
```


## Cleaning the  species traits  [life span] (https://genomics.senescence.info/species/index.html)

```{r}
datage<-read.csv("/Users/jm200/Library/CloudStorage/Dropbox/PhD Project/Chapter 4/Manuscript/Data/anage_data.csv", header=T)
datage$Species<-paste0(as.character(datage$genus)," ", as.character(datage$species))
datageclean <- datage%>% filter (dataquality %in% c("low","acceptable","high"))
names(datbodymass)
```


```{r}
databuncp %>%
  filter(Kingdom == "Animal")->databa
databuncpLS<-left_join(x=databa,y=datageclean,by="Species")
databuncpLS <- databuncpLS[!apply(databuncpLS[,c("maximumlongevity", "yi")], 1, anyNA),] 
```

```{r,eval=FALSE}
sp.aba <- databuncpLS[, 4]
spnamesaba <- as.character(sp.aba)
# species2 <- unique(spnamesaba)
species2 <- spnamesaba
#length(species2)
# species2 <- gsub("\n", "", species1)
temp <-taxize::gnr_resolve(species2, best_match_only = TRUE, canonical = TRUE)
dat_aba <- merge(data.frame(species2), temp[,c("user_supplied_name", "matched_name2")],
                 by.x = "species2", by.y = "user_supplied_name", all.x = TRUE)
spnamesaba <-dat_aba[,2]
# spnamesaba <- spnamesabp[complete.cases(spnamesaba)]
# length(spnamesaba)
# Retrieve classifications
spnamesabaf <- classification(spnamesaba, db = "ncbi")
resolved_names_aba <- unique(spnamesabaf)
# Building a phylogenetic tree
abatree <- class2tree(resolved_names_aba)
# str(abptree)
(abaphy <- abatree$phylo)
# length(abaphy$edge.length)
```

```{r}
speciesabuna<-data.frame(Species=abaphy$tip.label)
abunadat<- merge(x = speciesabuna, y = databuncpLS,all.x = TRUE)
abunadat<-abunadat[complete.cases(abunadat), ]
abunacleanTree<-drop.tip(abaphy, abaphy$tip.label[-na.omit(match(abunadat$Species, abaphy$tip.label))])
A<-ape::vcv.phylo(abunacleanTree) # Creates variance-covariance matrix from tree
```

```{r}

Lifespan.abunmodel <- brms::brm(
  yi|se(se, sigma = TRUE) ~ scale(maximumlongevity) + (1|Author) + (1|gr(Species, cov = A)),
  data = abunadat,
  family = gaussian(),
  data2 = list(A = A),
  prior = c(
    prior(normal(0, 1), "b"),
    prior(normal(0, 0.5), "Intercept"),
    prior(student_t(3, 0, 0.2), "sd"),
    prior(student_t(3, 0, 0.2), "sigma")
  ),
  iter = 5000, cores = 4
)
  
```

```{r}
plot(Lifespan.abunmodel)
summary(Lifespan.abunmodel)
```

```{r}
theme <-  theme(panel.background = element_blank(),
                panel.border = element_rect(fill = NA), 
                strip.background = element_blank(),
                axis.text.x = element_text(size = 10, colour = 'black'),
                axis.text.y = element_text(size = 10, colour = 'black'),
                #text        = element_text(size = 8),
                axis.title.y = element_text(size = 8, hjust = 0.5, angle = 90),# size of y lab
                axis.title.x = element_text(size = 8),# size of X lab
                legend.position     = "none",
                legend.title        = element_blank(),
                legend.text         = element_blank(),
                legend.background   = element_blank(),
                legend.key          = element_blank(),
                plot.title  = element_text(hjust = 0.5),
                plot.margin = margin(t = 0.5, r = 0.5, b = 0.5, l = 0.5, unit = "cm"))


Figure6 <- ggplot(databuncpLS, aes(
  x = maximumlongevity, y = yi
)) +
  geom_point(aes(colour="orange2"),data =databuncpLS ) +
  # geom_hline(yintercept = 0, linetype="dashed",color = "black", size=0.5)+
  scale_y_continuous(limits = c(-2, 2))+
  annotate("text",size = 2.8,label="-0.00 [-0.01,0.01]", x =60,y=-1) +
  annotate("text", x = 8, y = 1.8, label = "d", color = "black",  size = 5) +
  labs(x = "Longetivity (year)", y = "Effect size") +
  theme(panel.border = element_rect( fill=NA),
             text = element_text(size = 10), # change font sizes
                 legend.position = "none",
                 legend.title = element_text(size = 8),
                 legend.text = element_text(size = 8),
                 legend.background = element_blank(),
                 panel.background = element_rect(fill="transparent"), 
                 axis.text.x = element_text(size = 10, colour = "black"),
                 axis.text.y = element_text(size = 10, colour = "black", hjust = 0.5, angle = 90))

```

