---
title: 'Meta-analysis reveals that species are not most abundant or demographically performant at their range’s centre despite higher genetic diversity'
author: "Jacob Moutouama and Orou Gaoue"
date: "`r Sys.Date()`"
output: 
  rmdformats::readthedown:
    code_folding: hide
    self_contained: true
    number_sections: true
    thumbnails: true
    lightbox: true
    gallery: true
    keep_md: true
    highlight: tango
    df_print: kable 
    toc_depth: 3
    fig_width: 8
    fig_height: 8
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(external = TRUE, echo = F, warning = FALSE, fig.pos = "H", global.device = TRUE)
a4width <- 8.3
a4height <- 11.7
options(knitr.table.format = "latex")
```

This is the code for the paper 'The center-periphery hypothesis revisited: a meta-analysis'.

```{r eval=TRUE,echo=T}
rm(list = ls(all = TRUE))
```

```{r}
# Load the required packages
library(taxize)
library(ape)
library(phytools)
library(ade4)
library(ggplot2)
library(RColorBrewer)
library(dplyr)
library(tidytree)
library(aplot)
library(cowplot)
library(ggpubr)
library(ggsci)
library(R.rsp)
library(flextable)
library(effectsize)
library(metafor)
library(esc)
library(knitr)
library(tidyverse)
library(brms)
library(coda)
library(modelr)
library(gridExtra)
library(pBrackets)
library(performance)
library(kableExtra)
library(tidybayes)
library(formattable)
library(grid)
library(rstan)
library(lme4)
library(lmerTest)
library(BIEN)
library(readr)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(stringr)
library(tidyr)
library(writexl)
library(ggridges)
library(glue)
library(forcats)
library(tidybayes)
library(posterior)
library(bayestestR)
library(rgbif)
library(patchwork)
#library(scater)
rstan_options(auto_write = TRUE)
set.seed(13)
```

Function to export the model summary 
```{r}
Ftable <- function(model, reorder = F, order, percent = F){
  table <- summary(model)$fixed
  table2 <-summary(model, prob = 0.9)$fixed %>% as_tibble()
  names <- rownames(table)
  table <- table %>% 
    as_tibble() %>% 
    dplyr::rename('L95%CL' = `l-95% CI`, 'U95%CL' = `u-95% CI`) %>% 
    select(-Rhat, -Bulk_ESS, -Tail_ESS)
  table$`L90%CL` <- table2$`l-90% CI`
  table$`U90%CL` <- table2$`u-90% CI`
  
  table <-table %>% 
    mutate(Estimate = ifelse(`L95%CL`	* `U95%CL` > 0, 
                             paste(sprintf('%.3f', round(Estimate, 3)),  '*', sep = ''), 
                             ifelse(`L90%CL`	* `U90%CL` > 0, 
                                    paste(sprintf('%.3f', round(Estimate, 3)), '\206', sep = ''), round(Estimate, 3))),
           rowname = names) %>% 
    dplyr::rename('SE    ' = Est.Error)  %>% 
    column_to_rownames(var = 'rowname') 

  if (reorder == T){
    table <- table[order,]
  }
  
  table <- table %>% 
    kable(digits = 3, escape = F, table.attr = "style = \"color: black;\"") %>% 
    kable_styling(position = "left")
  return(table)
}
```

Theme to use for all the graphs

```{r}
my_theme <- theme(
  panel.background = element_blank(),
  panel.border = element_rect(fill = NA), 
  strip.background = element_blank(),
  axis.text.x = element_text(size = 10, colour = 'black'),
  axis.text.y = element_text(size = 10, colour = 'black'),
  axis.title.y = element_text(size = 10),
  axis.title.x = element_text(size = 10),
  legend.position = "none",
  text = element_text(family = "Helvetica", size = 10, face = "plain"),
  legend.title = element_text(size = 9),
  legend.text = element_text(size = 8),
  legend.key.height = unit(0.8, "line"),
  legend.background = element_blank(),
  legend.key = element_rect(colour = NA, fill = NA),
  plot.title = element_text(hjust = 0.5),
  plot.margin = margin(t = 0.5, r = 0.5, b = 0.5, l = 0.5, unit = "cm")
)
```

# Location of empirical studies tesing the CPH
```{r}
# Read your country list 
country_data <- read.csv("/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Data/location_study.csv",
                         header = TRUE, stringsAsFactors = FALSE)

# If multiple countries in one row, split them
countries_long <- country_data %>%
  mutate(Country = str_split(Country, ",")) %>%
  unnest(Country) %>%
  mutate(Country = str_trim(Country)) %>%
  filter(Country != "") %>%
  distinct(Country)

unique_countries <- countries_long$Country

# Get world map
world <- ne_countries(scale = "medium", returnclass = "sf")

# Filter only the countries from your list
world_selected <- world %>%
  filter(admin %in% unique_countries)

# Add centroid points for occurrences
# Reproject to Mollweide (good global projection)
world_selected_proj <- st_transform(world_selected, crs = "+proj=moll")

# Safe centroids inside polygons
centroids_proj <- st_point_on_surface(world_selected_proj)

# Transform back to lon/lat
centroids <- st_transform(centroids_proj, crs = st_crs(world_selected))


# # ---- 4. Create tropical belt polygon (between -23.5 and 23.5 latitude) ----
# tropics <- st_polygon(list(rbind(
#   c(-180, -23.5), c(180, -23.5), c(180, 23.5), c(-180, 23.5), c(-180, -23.5)
# ))) %>%
#   st_sfc(crs = st_crs(world))

tropic_lines <- data.frame(
  long = c(-180, 180, -180, 180),
  lat  = c(-23.5, -23.5, 23.5, 23.5),
  group = c("South", "South", "North", "North")
)
# ---- 5. Plot ----
Fig0 <- ggplot() +
  # Ocean background (light blue rectangle spanning world extent)
  annotate("rect", xmin = -180, xmax = 180, ymin = -90, ymax = 90,
           fill = "lightblue", color = NA) +
  geom_sf(data = world, fill = "grey90", color = "white") +
  geom_line(data = tropic_lines, aes(x = long, y = lat, group = group),
            color = "grey30", size = 0.6, linetype = "dashed") +
  geom_sf(data = world_selected, fill = "grey90", color = "white") + # Selected countries
  geom_sf(data = centroids, color = "red", size = 1) +               # Occurrence points
  #geom_sf_text(data = centroids, aes(label = admin), size = 3, vjust = -1) + # Labels
  coord_sf(expand = FALSE) +
  theme_light() +
  labs(x = "Longitude", y = "Latitude") +
  theme(
    axis.title = element_text(size = 14),  # x and y label size
    axis.text = element_text(size = 12)    # tick label size
  )


# Export as PDF
# ggsave(filename = "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/map_study_countries.pdf",
#        plot = Fig0, width = 10, height = 6)

```

# What is the magnitude of the cumulative effect size across studies for CPH support, and is it significantly different from zero?  

## Population growth rate and CPH

```{r}
datlambda <- read.csv("/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Data/datlambda.csv", header = T)
```

```{r}
#unique(datlambda$Method)
```


# Stastic to get how many of data we have per species. 
```{r}
total_studies <- n_distinct(datlambda$ID)

# Total number of species
# Count unique species overall
total_species <- n_distinct(datlambda$Species)

# Count unique species by kingdom
species_by_kingdom <- datlambda %>%
  group_by(Kingdom) %>%
  summarize(n_species = n_distinct(Species))

# Count repetitions for each species
species_counts <- datlambda %>%
  group_by(Species, Kingdom) %>%
  summarize(n_records = n()) %>%
  arrange(desc(n_records))
```

### Grand Mean effect size and significance 

```{r}
# Define priors
priors_lb <- c(
  prior(normal(0, 1), class = "Intercept"),
  prior(cauchy(0, 0.5), class = "sd")
)

# 6. Fit brms model with phylogeny
Lambdamodel <- brm(
  yi | se(se,sigma=TRUE) ~ 1 + (1|ID) + (1|Species),
  data = datlambda,
  seed = 13,
  family = gaussian(),
  prior = priors_lb,
  control = list(adapt_delta = 0.99, max_treedepth = 15),
  iter = 3000,
  cores = 4
)
```

### Predictive check for population growth rate

```{r}
pp_check(Lambdamodel,ndraws = 1000)+
  xlab("Population growth rate")+ylab("Density")+
  ggtitle("")+
  my_theme->ppc_growthrate
```

### Value for Grand Mean effect size

```{r}
Lambdamodel %>% Ftable(percent = T)
```

### empirical cumulative distribution function

```{r}
#get_variables(Lambdamodel)
post.samples_lambda <- posterior_samples(Lambdamodel, pars="^b_Intercept")
post.samples_lambda <- as_draws_df(Lambdamodel)  # gets all posterior draws as a data.frame
# Select only intercepts (columns starting with "b_Intercept")
post.samples_lambda <- post.samples_lambda[, grepl("^b_Intercept", colnames(post.samples_lambda))]
mean(post.samples_lambda$b_Intercept>0)
smd.ecdf_lambda <- ecdf(post.samples_lambda$b_Intercept)
smd.ecdf_lambda(0)
#We see that with 28.73%, the probability of our pooled effect being smaller than 0.30 is very, very low. 
```

### Forest plot for lambda 
```{r}
# Extract study- and species-level draws
study.draws_lb <- Lambdamodel %>%
  spread_draws(r_ID[ID,], r_Species[Species,], b_Intercept) %>%
  # Combine random effects and intercept
  mutate(b_Intercept = r_ID + r_Species + b_Intercept)

# Create pooled effect draws (just fixed effect intercept)
pooled.effect.draws_lb <- Lambdamodel %>%
  spread_draws(b_Intercept) %>%
  mutate(ID = "Pooled Effect")

# Prepare study-level data
study_data_lb <- study.draws_lb %>%
  mutate(ID = as.character(ID)) %>%
  mutate(ID = str_replace_all(ID, "[.]", " ")) %>%
  # reorder studies by effect size
  mutate(ID = reorder(ID, b_Intercept))

# Bind pooled effect at bottom
forest.data_lb <- bind_rows(study_data_lb, pooled.effect.draws_lb) %>%
  # ensure ID is character so releveling works
  mutate(ID = as.character(ID)) %>%
  # ensure Pooled Effect is last
  mutate(ID = fct_relevel(ID, "Pooled Effect", after = Inf)) %>%
  # add a flag for coloring in the plot
  mutate(effect_type = ifelse(ID == "Pooled Effect", "Pooled", "Study")) %>%
  ungroup()

# Summaries (mean and 95% interval)
forest.data.summary_lb <- forest.data_lb %>%
  group_by(ID) %>%
  mean_qi(b_Intercept) %>%
  # ensure Pooled Effect is last in summary as well
  mutate(ID = fct_relevel(ID, "Pooled Effect", after = Inf)) %>%
  # add the same effect_type column for the summary
  mutate(effect_type = ifelse(ID == "Pooled Effect", "Pooled", "Study"))
```

```{r}
pdf("/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/lambda_cp_forest.pdf", width = 6, height = 6)
ggplot(forest.data_lb, aes(x = b_Intercept, y = ID, fill = effect_type)) +
  # Vertical lines: pooled effect, CI, and zero
  #geom_vline(xintercept = fixef(Lambdamodel)[1, 1], color = "grey", size = 1) +
  geom_vline(xintercept = fixef(Lambdamodel)[1, 3:4],
             color = "grey", linetype = 2) +
  geom_vline(xintercept = 0, color = "black", size = 0.75) +
  
  # Ridge densities
  geom_density_ridges(rel_min_height = 0.01, col = NA, scale = 1, alpha = 0.8) +
  
  # Point-intervals (explicit xmin/xmax)
  geom_pointinterval(
    data = forest.data.summary_lb,
    aes(x = b_Intercept, y = ID, xmin = .lower, xmax = .upper, color = effect_type),
    size = 1,
    orientation = "y"
  ) +
  
  # Labels with estimates and intervals
  geom_text(
    data = mutate_if(forest.data.summary_lb, is.numeric, round, 2),
    aes(label = glue("{b_Intercept} [{.lower}, {.upper}]"), y = ID, x = 2.8),
    hjust = "inward",size = 3
  ) +
  
  # Discrete y-axis scale (reversed so top-down)
  scale_y_discrete(limits = rev(levels(forest.data_lb$ID)),expand = expansion(add = 0.75)) +
  
  # Manual colors for pooled vs studies
  scale_fill_manual(values = c("Study" = "#E69F00", "Pooled" = "#56B4E9")) +
  scale_color_manual(values = c("Study" = "#E69F00", "Pooled" = "#56B4E9")) +
  
  labs(x = "Effect size (lambda)", y = NULL) +
  my_theme
dev.off()

```

### Heterogeneity for population growth rate
```{r}
# Function to calculate h (quantified heterogeneity in percent)
bayes_h <- function(fit, group = c("ID", "Species", "residual")) {
  library(posterior)
  library(parameters)
  
  # Extract posterior draws
  post <- as_draws_df(fit)
  
  # Variance components
  tau2_ID <- if ("sd_ID__Intercept" %in% names(post)) post$`sd_ID__Intercept`^2 else 0
  tau2_Species <- if ("sd_Species__Intercept" %in% names(post)) post$`sd_Species__Intercept`^2 else 0
  sigma2 <- if ("sigma" %in% names(post)) post$sigma^2 else 0
  
  # Total variance
  V_total <- tau2_ID + tau2_Species + sigma2
  
  out <- list()
  
  if ("ID" %in% group) {
    h_ID <- 100 * tau2_ID / V_total
    out$ID <- describe_posterior(h_ID, centrality = "all", ci = 0.95)
  }
  
  if ("Species" %in% group) {
    h_Species <- 100 * tau2_Species / V_total
    out$Species <- describe_posterior(h_Species, centrality = "all", ci = 0.95)
  }
  
  if ("residual" %in% group) {
    h_residual <- 100 * sigma2 / V_total
    out$residual <- describe_posterior(h_residual, centrality = "all", ci = 0.95)
  }
  
  return(out)
}

# Example usage
bayes_h(Lambdamodel)


```

## Genetics and CPH

```{r}
datgencp <- read.csv("/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Data/datgencp.csv", header = T)
#unique(datgencp$Method)
```

```{r}
total_studies_g <- n_distinct(datgencp$ID)
# Count unique species overall
total_species <- n_distinct(datgencp$Species)

# Count unique species by kingdom
species_by_kingdom <- datgencp %>%
  group_by(Kingdom) %>%
  summarize(n_species = n_distinct(Species))

# Count repetitions for each species
species_counts <- datgencp %>%
  group_by(Species, Kingdom) %>%
  summarize(n_records = n()) %>%
  arrange(desc(n_records))

total_species
species_by_kingdom
```

### Grand Mean effect size and significance 

```{r, warning=FALSE}
priors_gn <- c(prior(normal(0,1), class = Intercept),
            prior(cauchy(0,0.5), class = sd))
genamodel <- brm(
  yi | se(se,sigma=TRUE) ~ 1 + (1 | ID) + (1|Species),
  control = list(adapt_delta = 0.9999, max_treedepth = 20),
  prior = priors_gn, iter = 8000, cores = 4,seed = 13,
  data = datgencp
)
```

### Predictive check for genetic diversity

```{r}
pp_check(genamodel,ndraws = 1000)+
  xlab("Genetic diversity")+ylab("Density")+
  ggtitle((""))+my_theme->ppc_genetic
```

### Value for Grand Mean effect size for genetic diversity

```{r}
genamodel %>% Ftable(percent = T)
```

# Heterogeinity for genetic diversity 

```{r}
bayes_h(genamodel)

```

### empirical cumulative distribution function
```{r}
#get_variables(Lambdamodel)
post.samples_gen <- posterior_samples(genamodel, pars="^b_Intercept")
mean(post.samples_gen$b_Intercept>0)
smd.ecdf_gen <- ecdf(post.samples_gen$b_Intercept)
smd.ecdf_gen(0)
# ~98.86% of values are > 0.
```

### Forest plot for genetic 

```{r}
# Extract study- and species-level draws
study.draws_gn <- genamodel %>%
  spread_draws(r_ID[ID,], r_Species[Species,], b_Intercept) %>%
  # Combine random effects and intercept
  mutate(b_Intercept = r_ID + r_Species + b_Intercept)

# Create pooled effect draws (just fixed effect intercept)
pooled.effect.draws_gn <- genamodel %>%
  spread_draws(b_Intercept) %>%
  mutate(ID = "Pooled Effect")

# Prepare study-level data
study_data_gn <- study.draws_gn %>%
  mutate(ID = as.character(ID)) %>%
  mutate(ID = str_replace_all(ID, "[.]", " ")) %>%
  # Reorder studies by effect size
  mutate(ID = reorder(ID, b_Intercept))

# Bind pooled effect at bottom
forest.data_gn <- bind_rows(study_data_gn, pooled.effect.draws_gn) %>%
  mutate(ID = as.character(ID)) %>%
  mutate(ID = fct_relevel(ID, "Pooled Effect", after = Inf)) %>%
  mutate(effect_type = ifelse(ID == "Pooled Effect", "Pooled", "Study")) %>%
  ungroup()

# Summaries (mean and 95% interval)
forest.data.summary_gn <- forest.data_gn %>%
  group_by(ID) %>%
  mean_qi(b_Intercept) %>%
  mutate(ID = fct_relevel(ID, "Pooled Effect", after = Inf)) %>%
  mutate(effect_type = ifelse(ID == "Pooled Effect", "Pooled", "Study"))
```

```{r}
# Plot and save
pdf("/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/gen_cp_forest.pdf", 
    width = 5, height = 10)

ggplot(forest.data_gn, aes(x = b_Intercept, y = ID, fill = effect_type)) +
  
  # Vertical lines: pooled effect CI and zero
  geom_vline(xintercept = fixef(genamodel)[1, 3:4], color = "grey", linetype = 2) +
  geom_vline(xintercept = 0, color = "black", size = 0.75) +
  
  # Ridge densities
  geom_density_ridges(rel_min_height = 0.01, col = NA, scale = 1, alpha = 0.8) +
  
  # Point-intervals (mean + 95% CI)
  geom_pointinterval(
    data = forest.data.summary_gn,
    aes(x = b_Intercept, y = ID, xmin = .lower, xmax = .upper, color = effect_type),
    size = 1, orientation = "y"
  ) +
  # Limitation of x axis
  xlim(-1.5,4.5) +
  
  # Labels with estimates and intervals
  geom_text(
    data = mutate_if(forest.data.summary_gn, is.numeric, round, 2),
    aes(label = glue("{b_Intercept} [{.lower}, {.upper}]"), y = ID, x = 4),
    hjust = "inward", size = 3
  ) +
  
  # Reverse y-axis
  scale_y_discrete(limits = rev(levels(forest.data_gn$ID)), expand = expansion(add = 0.75)) +
  
  # Manual colors
  scale_fill_manual(values = c("Study" = "#E69F00", "Pooled" = "#56B4E9")) +
  scale_color_manual(values = c("Study" = "#E69F00", "Pooled" = "#56B4E9")) +
  
  labs(x = "Effect size (genetic diversity)", y = NULL) +
  my_theme

dev.off()

```


## Population abundance and CPH

```{r}
databuncp <- read.csv("/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Data/databuncp.csv", header = T)
databuncp$Species<-gsub("spp.","sp.",as.character(databuncp$Species))
unique(databuncp$Method)
databuncp <- databuncp %>% filter(vi > 0)
```

```{r}
total_studies_ab <- n_distinct(databuncp$ID)
# Count unique species overall
total_species <- n_distinct(databuncp$Species)

# Count unique species by kingdom
species_by_kingdom <- databuncp %>%
  group_by(Kingdom) %>%
  summarize(n_species = n_distinct(Species))

# Count repetitions for each species
species_counts <- databuncp %>%
  group_by(Species, Kingdom) %>%
  summarize(n_records = n()) %>%
  arrange(desc(n_records))

total_species
species_by_kingdom
```

### Grand Mean effect size and significance 

```{r}
abunmodel <- brm(
  yi | se(se,sigma=TRUE) ~ 1 + (1 | ID) + (1 | Species),
  data = databuncp,seed = 13,
  iter = 4000, warmup = 2000, chains = 4
)
```

### Predictive check for population abundance 

```{r}
ppc_abundance <- pp_check(abunmodel, ndraws = 1000) +
  xlab("Abundance") +
  ylab("Density") +
  ggtitle("") +
  coord_cartesian(xlim = c(-2, 2), ylim = c(0, 3)) +
  my_theme
```

```{r}
abunmodel %>% Ftable(percent = T)
#summary(abunmodel)
```

### Heterogeinity for abundance data

```{r}
bayes_I2(abunmodel, group = c("ID", "Species", "total"))
bayes_h(abunmodel)
```


### empirical cumulative distribution function
```{r}
# Extract intercept posterior samples as numeric
post.samples_ab <- as_draws_df(abunmodel) %>%
  dplyr::pull(b_Intercept)
# Posterior probability Intercept > 0
p_greater <- mean(post.samples_ab > 0)
# Posterior probability Intercept <= 0
p_less_equal <- ecdf(post.samples_ab)(0)
c(P_greater = p_greater, P_less_equal = p_less_equal)

```

### Forest plot for abundance
```{r}
# Extract study- and species-level draws
study.draws_ab <- abunmodel %>%
  spread_draws(r_ID[ID,], r_Species[Species,], b_Intercept, ndraws = 1000) %>%
  # Combine random effects and intercept
  mutate(b_Intercept = r_ID + r_Species + b_Intercept)

# Create pooled effect draws (just fixed effect intercept)
pooled.effect.draws_ab <- abunmodel %>%
  spread_draws(b_Intercept, ndraws = 1000) %>%
  mutate(ID = "Pooled Effect")

# Prepare study-level data
study_data_ab <- study.draws_ab %>%
  mutate(ID = as.character(ID)) %>%
  mutate(ID = str_replace_all(ID, "[.]", " ")) %>%
  # Reorder studies by effect size
  mutate(ID = reorder(ID, b_Intercept))

# Bind pooled effect at bottom
forest.data_ab <- bind_rows(study_data_ab, pooled.effect.draws_ab) %>%
  mutate(ID = as.character(ID)) %>%
  mutate(ID = fct_relevel(ID, "Pooled Effect", after = Inf)) %>%
  mutate(effect_type = ifelse(ID == "Pooled Effect", "Pooled", "Study")) %>%
  ungroup()
# Summaries (mean and 95% interval)
forest.data.summary_ab <- forest.data_ab %>%
  group_by(ID) %>%
  mean_qi(b_Intercept) %>%
  mutate(ID = fct_relevel(ID, "Pooled Effect", after = Inf)) %>%
  mutate(effect_type = ifelse(ID == "Pooled Effect", "Pooled", "Study"))

```

```{r}
# Plot and save
pdf("/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/abun_cp_forest.pdf", 
    width = 5, height =8)

ggplot(forest.data_ab, aes(x = b_Intercept, y = ID, fill = effect_type)) +
  
  # Vertical lines: pooled effect CI and zero
  geom_vline(xintercept = fixef(abunmodel)[1, 3:4], color = "grey", linetype = 2) +
  geom_vline(xintercept = 0, color = "black", size = 0.75) +
  
  # Ridge densities
  geom_density_ridges(rel_min_height = 0.01, col = NA, scale = 1, alpha = 0.8) +
  
  # Point-intervals (mean + 95% CI)
  geom_pointinterval(
    data = forest.data.summary_ab,
    aes(x = b_Intercept, y = ID, xmin = .lower, xmax = .upper, color = effect_type),
    size = 1, orientation = "y"
  ) +
  # Limitation of x axis
  xlim(-2,4) +
  # Labels with estimates and intervals
  geom_text(
    data = mutate_if(forest.data.summary_ab, is.numeric, round, 2),
    aes(label = glue("{b_Intercept} [{.lower}, {.upper}]"), y = ID, x = 3.75),
    hjust = "inward", size = 3
  ) +
  
  # Reverse y-axis
  scale_y_discrete(limits = rev(levels(forest.data_ab$ID)), expand = expansion(add = 0.75)) +
  
  # Manual colors
  scale_fill_manual(values = c("Study" = "#E69F00", "Pooled" = "#56B4E9")) +
  scale_color_manual(values = c("Study" = "#E69F00", "Pooled" = "#56B4E9")) +
  
  labs(x = "Effect size (abundance)", y = NULL) +
  my_theme

dev.off()
```



## Testing if adding data from Dallas et al , 2017  improve the results

```{r}
databuncp %>% 
  filter(!Author %in% c("Dallas et al.", "Santini et al.","Osocrio-Olivera et al"))->Abundcp_SDM
```

### Grand Mean effect size and significance 

```{r, warning=FALSE}
abunmodel_SDM <- brm(
  yi | se(se, sigma = TRUE) ~ 1 + (1 | ID) + (1|Species),
  data = Abundcp_SDM,seed = 13,
  control = list(adapt_delta = 0.95)
)
```

```{r}
abunmodel_SDM %>% Ftable(percent = T)
```

```{r,eval=FALSE}
# Average effect
out_f_SDM <- spread_draws(abunmodel_SDM, b_Intercept) %>% 
  mutate(study = "Abundance")
# Data frame of summary numbers
out_all_sum_SDM <- group_by(out_f_SDM, study) %>% 
  mean_qi(b_Intercept)
```

```{r}
# Draw plot
pdf("/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/Figure_SDM.pdf", width = 6, height = 4)
(Fig1SDM<-out_f_SDM %>%   
  ggplot(aes(b_Intercept, study,fill = after_stat(x < 0))) +
  labs(x = "Effect size", y = '')+
  xlim(-0.5,1.8)+
  # Zero!
  geom_vline(xintercept = 0, linewidth = .25, lty = 2) +
  stat_halfeye(.width = c(.95)) +
  scale_fill_manual(values = c("#FFC34D", "#999999"))+
  # Add text labels
  geom_text(
    data = mutate_if(out_all_sum_SDM, is.numeric, round, 2),
    aes(label = str_glue("{b_Intercept} [{.lower}, {.upper}]"), x =1.5),
    hjust = "inward"
  ) +
  my_theme)
dev.off()
```

### Predictive check for all the models (abundance, population growth rate, population abundance)
```{r}
multiplot <- function(..., plotlist=NULL, cols=1) {
  require(grid)
  plots <- c(list(...), plotlist)
  numPlots = length(plots)
  
  layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                   ncol = cols, nrow = ceiling(numPlots/cols))
  
  if (numPlots==1) {
    print(plots[[1]])
  } else {
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    for (i in 1:numPlots) {
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

```

```{r}
# print figure
pdf("/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/PPC.pdf",useDingbats = F,height=6,width=6)
multiplot(ppc_growthrate,ppc_genetic,
          ppc_abundance,cols=2)
dev.off()
```

### Plot for Grand Mean effect size for abundance, genetic diversity and population growth rate
```{r}
# Average effect
out_Lambda <- spread_draws(Lambdamodel, b_Intercept) %>% 
  mutate(study = "Population\ngrowth rate")
# Data frame of summary numbers
out_all_sum_lambda <- group_by(out_Lambda, study) %>% 
  mean_qi(b_Intercept)
# Draw plot
Fig1a <- out_Lambda %>%
  ggplot(aes(x = b_Intercept, y = study, fill = after_stat(x < 0))) +
  labs(x = "", y = "") +
  xlim(-3, 2) +
  geom_vline(xintercept = 0, linewidth = .25, lty = 2) +
  stat_halfeye(.width = c(.95)) +
  scale_fill_manual(values = c("#FFC34D", "#999999")) +
  geom_text(
    data = mutate_if(out_all_sum_lambda, is.numeric, round, 2),
    aes(label = str_glue("{b_Intercept} [{.lower}, {.upper}]"),
        x = -2., y = 1.05),
    hjust = "inward"
  ) +
  my_theme

```

```{r}
# Average effect
out_f_gen <- spread_draws(genamodel, b_Intercept) %>% 
  mutate(study = "Genetic \n diversity")
# Data frame of summary numbers
out_all_sum_gen <- group_by(out_f_gen, study) %>% 
  mean_qi(b_Intercept)
# Draw plot
Fig1b<-out_f_gen %>%   
  ggplot(aes(x=b_Intercept, y=study,fill = after_stat(x < 0))) +
  labs(x = "", y = '')+
  xlim(-2,2)+
  # Zero!
  geom_vline(xintercept = 0, linewidth = .25, lty = 2) +
  stat_halfeye(.width = c(.95)) +
  scale_fill_manual(values = c("#FFC34D", "#999999"))+
  # Add text labels
  geom_text(
    data = mutate_if(out_all_sum_gen, is.numeric, round, 2),
    aes(label = str_glue("{b_Intercept} [{.lower}, {.upper}]"), x =-1.6),
    hjust = "inward"
  ) +
  my_theme
  
```


```{r}
# Average effect
out_f_abun <- spread_draws(abunmodel, b_Intercept) %>% 
  mutate(study = "Abundance")
# Data frame of summary numbers
out_all_sum_abun <- group_by(out_f_abun, study) %>% 
  mean_qi(b_Intercept)
# Draw plot
Fig1c<-out_f_abun %>%   
  ggplot(aes(x=b_Intercept, y=study,fill = after_stat(x < 0))) +
  labs(x = "", y = '')+
  xlim(-2,2)+
  # Zero!
  geom_vline(xintercept = 0, linewidth = .25, lty = 2) +
  stat_halfeye(.width = c(.95)) +
  scale_fill_manual(values = c("#FFC34D", "#999999"))+
  # Add text labels
  geom_text(
    data = mutate_if(out_all_sum_abun, is.numeric, round, 2),
    aes(label = str_glue("{b_Intercept} [{.lower}, {.upper}]"), x =-1.6),
    hjust = "inward"
  ) +
  my_theme
```

```{r}
out_all <- bind_rows(out_Lambda,out_f_abun, out_f_gen) 
out_all_sum <-bind_rows(out_all_sum_lambda, out_all_sum_abun,out_all_sum_gen) 
pdf("/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/Figure1.pdf", width = 6, height = 5)
(Figure1<-out_all %>%   
  ggplot(aes(b_Intercept, study,fill = after_stat(x < 0))) +
  labs(x = "Effect size", y = '')+
  xlim(-1.8,1.8)+
  # Zero!
  geom_vline(xintercept = 0, linewidth = .25, lty = 2) +
  stat_halfeye(.width = c(.95)) +
  scale_fill_manual(values = c("#FFC34D", "#999999"))+
  # Add text labels
  geom_text(size = 3,
    data = mutate_if(out_all_sum, is.numeric, round, 2),
    aes(label = str_glue("{b_Intercept} [{.lower}, {.upper}]"), x =-1.5),
    hjust = "inward"
  )+
  my_theme)
dev.off()
```

# Detection of publication bias
## Population growth rate 

```{r}
fittedlambda <- fitted(
  Lambdamodel,
  newdata = datlambda,
  allow_new_levels = TRUE
)
datlambda$meta_resid <- datlambda$yi  - fittedlambda[,1]
datlambda <- datlambda %>% 
  mutate(p_i = 1/se,
         o_i = meta_resid * p_i)
# rm(fittedlambda)
```

egger's regression
```{r, eval =F}
m_egg_lambda <- brm(data = datlambda, o_i ~ p_i, control = list(adapt_delta = 0.999), cores = 4)
```

```{r}
summary(m_egg_lambda)
```

effect size ~ year
```{r, eval =F}
m_year_lambda <- update(Lambdamodel, .~. + Year, newdata = datlambda, cores = 4)
```

```{r}
summary(m_year_lambda)
```

```{r}
Lambdamodel_rm <- rma.mv(
  yi = yi,
  V = vi,
  method = "REML",
  test = "t",
  dfs = "contain",
  random = list(~1 | ID, ~1 | Species),
  data = datlambda
)
```

## Population genetic

```{r}
fitted_gen <- fitted(genamodel, datgencp)
datgencp$meta_resid <- datgencp$yi  - fitted_gen[,1]
datgencp <- datgencp %>% 
  mutate(p_i = 1/se,
         o_i = meta_resid * p_i)
```

egger's regression
```{r, eval =F}
m_egg_gen <- brm(data = datgencp, o_i ~ p_i, control = list(adapt_delta = 0.999), cores = 4)
```

```{r}
summary(m_egg_gen)
```

effect size ~ year
```{r, eval =F}
m_year_gen <- update(genamodel, .~. + Year, newdata = datgencp, cores = 4)
```

```{r}
summary(m_year_gen)
```

```{r}
genamodelpb <- rma.mv(yi = yi , V = vi, method = "REML", 
                   test="t", dfs="contain",
                   random=list(~1|ID,  ~1|Species),  data = datgencp)
```

## Population abundance

```{r}
fitted_abun <- fitted(abunmodel, databuncp)
databuncp$meta_resid <- databuncp$yi  - fitted_abun[,1]
databuncp <- databuncp %>% 
  mutate(p_i = 1/se,
         o_i = meta_resid * p_i)
#rm(fitted_abun)
```

egger's regression
```{r, eval =F}
m_egg_abun <- brm(data =databuncp, o_i ~ p_i, cores = 4)
```

```{r}
summary(m_egg_abun)
```

effect size ~ year
```{r, eval =F}
m_year_abun <- update(abunmodel, .~. + Year, newdata = databuncp, cores = 4)
```

```{r}
summary(m_year_abun)
```

```{r}
abunmodelpb <- rma.mv(yi = yi , V = vi, method = "REML", 
                   test="t", dfs="contain",
                   random=list(~1|ID,  ~1|Species),  data = databuncp)

```

```{r}
# Population growth rate
fitted_pop <- fitted(m_year_lambda, newdata = datlambda, re_formula = NA)
datlambda$pred <- fitted_pop[, "Estimate"]
datlambda$pred_lwr <- fitted_pop[, "Q2.5"]
datlambda$pred_upr <- fitted_pop[, "Q97.5"]

# Genetic diversity
fitted_gen <- fitted(m_year_gen, newdata = datgencp, re_formula = NA)
datgencp$pred <- fitted_gen[, "Estimate"]
datgencp$pred_lwr <- fitted_gen[, "Q2.5"]
datgencp$pred_upr <- fitted_gen[, "Q97.5"]

# Population abundance
fitted_abun <- fitted(m_year_abun, newdata = databuncp, re_formula = NA)
databuncp$pred <- fitted_abun[, "Estimate"]
databuncp$pred_lwr <- fitted_abun[, "Q2.5"]
databuncp$pred_upr <- fitted_abun[, "Q97.5"]

```

## Put all the funnels plot together

```{r}
# Layout: 3 top trend panels (larger), 3 bottom funnel plots
pdf("/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/Figure22.pdf", width = 9, height = 6)
layout_matrix <- matrix(c(1,2,3,
                          4,5,6), nrow=2, byrow=TRUE)
layout(layout_matrix, heights=c(1,1))  # top row taller

par(mar=c(5,4,2,2),family="Helvetica")
# Panel 1: Genetic diversity trend
datgencp <- datgencp[order(datgencp$Year), ]
ylim_range <- range(c(datgencp$yi, datgencp$pred_lwr, datgencp$pred_upr))
plot(datgencp$Year, datgencp$yi, 
     pch=19, col=adjustcolor("orange2", alpha.f=0.4),
     xlab="Year", ylab="Effect size (genetic diversity)",
     main="", xlim=c(min(datgencp$Year), max(datgencp$Year)), ylim=c(-5,5))
lines(datgencp$Year, datgencp$pred, col="orange2", lwd=3)
polygon(c(datgencp$Year, rev(datgencp$Year)),
        c(datgencp$pred_lwr, rev(datgencp$pred_upr)),
        col=adjustcolor("orange2", alpha.f=0.2), border=NA)
mtext("a", side=3, line=0.5, at=min(datgencp$Year), cex=1.2, font=1)

# Panel 2: Population abundance trend
databuncp <- databuncp[order(databuncp$Year), ]
#ylim_range <- range(c(databuncp$yi, databuncp$pred_lwr, databuncp$pred_upr))
plot(databuncp$Year, databuncp$yi, 
     pch=19, col=adjustcolor("orange2", alpha.f=0.4),
     xlab="Year", ylab="Effect size (abundance)",
     main="", xlim=c(min(databuncp$Year), max(databuncp$Year)), ylim=c(-2,2))
# lines(databuncp$Year, databuncp$pred, col="orange2", lwd=3)
# polygon(c(databuncp$Year, rev(databuncp$Year)),
#         c(databuncp$pred_lwr, rev(databuncp$pred_upr)),
#         col=adjustcolor("orange2", alpha.f=0.2), border=NA)
mtext("b", side=3, line=0.5, at=min(databuncp$Year), cex=1.2, font=1)
# Panel 3: Lambda trend
datlambda <- datlambda[order(datlambda$Year), ]
ylim_range <- range(c(datlambda$yi, datlambda$pred_lwr, datlambda$pred_upr))
plot(datlambda$Year, datlambda$yi, 
     pch=19, col=adjustcolor("orange2", alpha.f=0.4),
     xlab="Year", ylab="Effect size (population growth rate)",
     main="", xlim=c(min(datlambda$Year), max(datlambda$Year)), ylim=c(-5,5))
# lines(dathlambda$Year, dathlambda$pred, col="orange2", lwd=3)
# polygon(c(dathlambda$Year, rev(dathlambda$Year)),
#         c(dathlambda$pred_lwr, rev(dathlambda$pred_upr)),
#         col=adjustcolor("orange2", alpha.f=0.4), border=NA)
mtext("c", side=3, line=0.5, at=min(datlambda$Year), cex=1.2, font=1)
# Bottom row: funnel plots
funnel(Lambdamodel_rm, yaxis="seinv", level=c(90,95,99),
       shade=c("white","gray55","gray75"), ylab="Precision (1/SE)",
       xlab="Effect size (population growth rate)", cex.lab=1.2, col="orange2",
       back="white", digits=1, xlim=c(-4,4), ylim=c(0.01,75))
mtext("d", side=3, line=0.5, at=-4, cex=1.2, font=1)
box()

funnel(genamodelpb, yaxis="seinv", level=c(90,95,99),
       shade=c("white","gray55","gray75"), ylab="", xlab="Effect size (genetic diversity)",
       cex.lab=1.2, col=adjustcolor("orange2", alpha.f=0.4), back="white", digits=1, xlim=c(-4,4), ylim=c(0.01,75))
mtext("e", side=3, line=0.5, at=-4, cex=1.2, font=1)
box()

funnel(abunmodelpb, yaxis="seinv", level=c(90,95,99),
       shade=c("white","gray55","gray75"), ylab="", xlab="Effect size (abundance)",
       cex.lab=1.2, col=adjustcolor("orange2", alpha.f=0.4), back="white", digits=1, xlim=c(-4,4), ylim=c(0.01,75))
mtext("f", side=3, line=0.5, at=-4, cex=1.2, font=1)
box()
dev.off()
```

# To what extent can heterogeneity among studies be explained by hypothesized moderators that characterize study outcomes?

## Kingdom
```{r}
datgencp %>%
  group_by(Kingdom) %>%
  summarize(n_species = n_distinct(Species))
```

### Genetic diversity

```{r}
Kingdomg.model <- brms::brm(
  yi|se(se, sigma = TRUE) ~ (Kingdom-1) + (1|ID) + (1|Species),
  data = datgencp,
  family = gaussian(),
  prior = c(
    prior(normal(0, 1), "b"),
    prior(student_t(3, 0, 5), "sd"),
    prior(student_t(3, 0, 5), "sigma")
  ),
  control = list(adapt_delta = 0.9999, max_treedepth = 15),
                   iter = 5000, cores = 4
)
```

```{r}
summary(Kingdomg.model)
```
#### empirical cumulative distribution function
```{r}
get_variables(Kingdomg.model)
post.samples_kg <- posterior_samples(Kingdomg.model, pars=c("b_KingdomAnimal","b_KingdomPlant"))
mean(post.samples_kg$b_KingdomAnimal>0)
mean(post.samples_kg$b_KingdomPlant>0)
smd.ecdf_kg_al <- ecdf(post.samples_kg$b_KingdomAnimal)
smd.ecdf_kg_pl <- ecdf(post.samples_kg$b_KingdomPlant)
smd.ecdf_kg_al(0)
smd.ecdf_kg_pl(0)
# We see that with 0.0877, the probability of our pooled effect being smaller than 0 for animal 
# We see that with 0.0264, the probability of our pooled effect being smaller than 0 for plant is is very, very low

```

### Population abundance
```{r}
# Total number of species
total_species <- databuncp %>% 
  summarize(n_species = n_distinct(Species))

# Number of species per kingdom
species_by_kingdom <- databuncp %>%
  group_by(Kingdom) %>%
  summarize(n_species = n_distinct(Species))

total_species
species_by_kingdom

```

```{r}
Kingdomab.model <- brms::brm(
  yi|se(se, sigma = TRUE) ~ (Kingdom-1) + (1|ID) + (1|Species),
  data = databuncp,
  family = gaussian(),
  prior = c(
    prior(normal(0, 1), "b"),
    prior(student_t(3, 0, 5), "sd"),
    prior(student_t(3, 0, 5), "sigma")
  ),
  control = list(adapt_delta = 0.9999, max_treedepth = 15),
                   iter = 5000, cores = 4
)
```

```{r}
summary(Kingdomab.model)
```

#### empirical cumulative distribution function
```{r}
get_variables(Kingdomab.model)
post.samples_k_ab <- posterior_samples(Kingdomab.model, pars=c("b_KingdomAnimal","b_KingdomPlant"))
mean(post.samples_k_ab$b_KingdomAnimal>0)
mean(post.samples_k_ab$b_KingdomPlant>0)
smd.ecdf_kab_al <- ecdf(post.samples_k_ab$b_KingdomAnimal)
smd.ecdf_kab_pl <- ecdf(post.samples_k_ab$b_KingdomPlant)
smd.ecdf_kab_al(0)
smd.ecdf_kab_pl(0)
# We see that with 0.1968, the probability of our pooled effect being smaller than 0 for animal 
# We see that with 0.1442, the probability of our pooled effect being smaller than 0 for plant 

```

### Plot Kingdom vs effect size
```{r}
out_f_kg <- spread_draws(Kingdomg.model, b_KingdomAnimal) %>% 
  mutate(study = "Animal") %>% 
  dplyr::rename (Kingdom=b_KingdomAnimal)
# Data frame of summary numbers
out_all_sum_kg <- group_by(out_f_kg, study) %>% 
  mean_qi(Kingdom)

out_f_ka <- spread_draws(Kingdomg.model, b_KingdomPlant) %>% 
  mutate(study = "Plant") %>% 
  dplyr::rename (Kingdom=b_KingdomPlant)
# Data frame of summary numbers
out_all_sum_ka <- group_by(out_f_ka, study) %>% 
  mean_qi(Kingdom)

out_allk <- bind_rows(out_f_kg, out_f_ka) 
out_all_sumk <-bind_rows(out_all_sum_kg,out_all_sum_ka) 

Figure3a<-out_allk %>%   
  ggplot(aes(Kingdom, study,fill = after_stat(x < 0))) +
  labs(x = "Effect size (genetic diversity)", y = '')+
  xlim(-1.75,1.75)+
  # Zero!
  geom_vline(xintercept = 0, linewidth = .25, lty = 2) +
  stat_halfeye(.width = c(0.95, 0.95)) +
   scale_fill_manual(values = c("#F0E442", "#999999"))+
  # Add text labels
  geom_text(size = 2.8,
    data = mutate_if(out_all_sumk, is.numeric, round, 2),
    aes(label = str_glue("{Kingdom} [{.lower}, {.upper}]"), x =-1.5),
    hjust = "inward"
  ) +
  #annotate("text", x =-1.5,y=2.5, label = "a", color = "black",  size = 5) +
  # annotate(geom="text", x=1.7, y=1, label="n=11",size = 3,
  #             color="black")+
  # annotate(geom="text", x=1.7, y=2, label="n=41",size = 3,
  #             color="black")+
  theme(panel.border = element_rect(fill = NA),
             text = element_text(size = 12), # change font sizes
                 legend.position = "none",
                 legend.title = element_text(size = 8),
                 legend.text = element_text(size = 8),
                 legend.background = element_blank(),
                 panel.background = element_rect(fill="transparent"), 
                 axis.text.x = element_text(size = 12, colour = "black"),
                 axis.text.y = element_text(size = 12, colour = "black", hjust = 0.5, angle = 0))
```

```{r}
out_f_kaa <- spread_draws(Kingdomab.model, b_KingdomAnimal) %>% 
  mutate(study = "Animal") %>% 
  dplyr::rename (Kingdom=b_KingdomAnimal)
# Data frame of summary numbers
out_all_sum_kaa <- group_by(out_f_kaa, study) %>% 
  mean_qi(Kingdom)

out_f_kap <- spread_draws(Kingdomab.model, b_KingdomPlant) %>% 
  mutate(study = "Plant") %>% 
  dplyr::rename (Kingdom=b_KingdomPlant)
# Data frame of summary numbers
out_all_sum_kap <- group_by(out_f_kap, study) %>% 
  mean_qi(Kingdom)

out_allkab <- bind_rows(out_f_kaa, out_f_kap) 
out_all_sumkab <-bind_rows(out_all_sum_kaa,out_all_sum_kap) 

Figure3b<-out_allkab %>%   
  ggplot(aes(Kingdom, study,fill = after_stat(x < 0))) +
  labs(x = "Effect size (abundance)", y = "")+
  xlim(-1.2,1.2)+
  # Zero!
  geom_vline(xintercept = 0, linewidth = .25, lty = 2) +
  stat_halfeye(.width = c(0.95, 0.95)) +
   scale_fill_manual(values = c("#F0E442", "#999999"))+
  #Add text labels
  geom_text(size = 2.8,
    data = mutate_if(out_all_sumkab, is.numeric, round, 2),
    aes(label = str_glue("{Kingdom} [{.lower}, {.upper}]"), x =-1),
    hjust = "inward"
  ) +
  #annotate("text", x =-1,y=2.5, label = "b", color = "black",  size = 5) +
  # annotate(geom="text", x=1.1, y=1, label="n=337",size = 3,
  #             color="black")+
  # annotate(geom="text", x=1.1, y=2, label="n=3475",size = 3,
  #             color="black")+
  theme(panel.border = element_rect( fill=NA),
             text = element_text(size = 12), # change font sizes
                 legend.position = "none",
                 legend.title = element_text(size = 8),
                 legend.text = element_text(size = 8),
                 legend.background = element_blank(),
                 panel.background = element_rect(fill="transparent"), 
                 axis.text.x = element_text(size = 12, colour = "black"),
                 axis.text.y = element_text(size = 12, colour = "black"))
```

```{r}
pdf("/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/Figure3.pdf",
    width = 9, height = 4)

# Assign ggarrange to Figure3
Figure3 <- ggarrange(Figure3a, Figure3b,
                     ncol = 2, nrow = 1,
                     heights = c(1,1),        # equal height for vertical layout
                     align = "v",             # vertical alignment
                     labels = c("a","b"),     # add panel labels
                     label.x = 0.09,             # shift labels outside left
                     label.y = 1.02,          # above panel
                     font.label = list(size = 16, face = "plain"))  # face inside font.label

# Print the figure
Figure3

dev.off()
```

## Methods
```{r}
table(databuncp$Kingdom, databuncp$Method)
```

```{r}
Kingdomme.model <- brms::brm(
  yi|se(se, sigma = TRUE) ~ (Method-1) + (1|ID) + (1|Species),
  data = databuncp,
  family = gaussian(),
  prior = c(
    prior(normal(0, 1), "b"),
    prior(student_t(3, 0, 5), "sd"),
    prior(student_t(3, 0, 5), "sigma")
  ),
  control = list(adapt_delta = 0.9999, max_treedepth = 15),
                   iter = 5000, cores = 4
)
```

```{r}

summary(Kingdomme.model)
```

### empirical cumulative distribution function
```{r}
get_variables(Kingdomme.model)
post.samples_mt <- posterior_samples(Kingdomme.model, pars=c("b_MethodED","b_MethodMD","b_MethodGD"))
mean(post.samples_mt$b_KingdomAnimal>0)
mean(post.samples_mt$b_KingdomPlant>0)
smd.ecdf_mt_ed <- ecdf(post.samples_mt$b_MethodED)
smd.ecdf_mt_md <- ecdf(post.samples_mt$b_MethodMD)
smd.ecdf_mt_gd <- ecdf(post.samples_mt$b_MethodGD)
smd.ecdf_mt_ed(0)
smd.ecdf_mt_md(0)
smd.ecdf_mt_gd(0)
```

### Testing the difference between methods

```{r}
methohyp1<-hypothesis(Kingdomme.model,"MethodED-MethodGD=0")
```

```{r}
methohyp2<-hypothesis(Kingdomme.model,"MethodMD-MethodGD=0")
```

```{r}
methohyp2<-hypothesis(Kingdomme.model,"MethodMD-MethodED=0")
```

### Plot method and effect size 

```{r}
out_ed <- spread_draws(Kingdomme.model, b_MethodED) %>% 
  mutate(study = "Euclidean\n distance") %>% 
  dplyr::rename (Method=b_MethodED)
# Data frame of summary numbers
out_all_ed <- group_by(out_ed, study) %>% 
  mean_qi(Method)

out_md <- spread_draws(Kingdomme.model, b_MethodMD) %>% 
  mutate(study = "Mahalalobis\n distance") %>% 
  dplyr::rename (Method=b_MethodMD)
# Data frame of summary numbers
out_all_md <- group_by(out_md, study) %>% 
  mean_qi(Method)

out_gd <- spread_draws(Kingdomme.model, b_MethodGD) %>% 
  mutate(study = "Geographic\n distance") %>% 
  dplyr::rename (Method=b_MethodGD)
# Data frame of summary numbers
out_all_gd <- group_by(out_gd, study) %>% 
  mean_qi(Method)

out_all_method <- bind_rows(out_gd,out_ed, out_md) 
out_all_sum_method <-bind_rows(out_all_gd,out_all_ed,out_all_md) 
```

```{r}
pdf("/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/Figure6.pdf", width = 6, height = 4)
(Figure6<-out_all_method %>%   
  ggplot(aes(Method, study,fill = after_stat(x < 0))) +
  labs(x = "Effect size", y = '')+
  xlim(-1,1)+
  # Zero!
  geom_vline(xintercept = 0, linewidth = .25, lty = 2) +
  stat_halfeye(.width = c(0.95, 0.95)) +
  scale_fill_manual(values = c("#F0E442", "#999999"))+
  # Add text labels
  geom_text(size = 2.8,
    data = mutate_if(out_all_sum_method, is.numeric, round, 2),
    aes(label = str_glue("{Method} [{.lower}, {.upper}]"), x =-0.85),
    hjust = "inward"
  ) +
  # annotate("text", x =0.8,y=1, label = "n=1822", color = "black",  size = 3) +
  # annotate("text", x =0.8,y=2, label = "n=1484", color = "black",  size = 3) +
  # annotate("text", x =0.8,y=3, label = "n=506", color = "black",  size = 3) +
  theme(panel.border = element_rect(fill = NA),
             text = element_text(size = 10), # change font sizes
                 legend.position = "none",
                 legend.title = element_text(size = 8),
                 legend.text = element_text(size = 8),
                 legend.background = element_blank(),
                 panel.background = element_rect(fill="transparent"), 
                 axis.text.x = element_text(size = 10, colour = "black"),
                 axis.text.y = element_text(size = 10, colour = "black", hjust = 0.5, angle = 0)))
dev.off()
```

## Lifespan
### Cleaning the  species traits  [life span] (https://genomics.senescence.info/species/index.html)

```{r}
databuncp %>% 
  filter(Kingdom=="Animal")->databa

# Import ANAge
datage <- read.table(
  "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Data/anage_data.txt",
  header = TRUE,
  sep = "\t",
  stringsAsFactors = FALSE
)

datage_subset <- datage %>%
  mutate(Species = paste0(as.character(Genus), " ", as.character(Species)),
         Order = Order) %>%
  select(Species, Order, Maximum.longevity..yrs., Body.mass..g., Data.quality, Specimen.origin)

# Import PanTHERIA WR05
pan05 <- read.delim(
  "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Data/PanTHERIA_1-0_WR05_Aug2008.txt",
  header = TRUE,
  stringsAsFactors = FALSE
)

pan05_subset <- pan05 %>%
  mutate(
    Species = MSW05_Binomial,
    Order = MSW05_Order,
    Maximum.longevity..yrs. = ifelse(X17.1_MaxLongevity_m > 0, X17.1_MaxLongevity_m / 12, NA),
    Body.mass..g. = ifelse(X5.1_AdultBodyMass_g > 0, X5.1_AdultBodyMass_g, NA),
    Data.quality = NA,
    Specimen.origin = NA
  ) %>%
  select(Species, Order, Maximum.longevity..yrs., Body.mass..g., Data.quality, Specimen.origin)
# Import PanTHERIA WR93
pan93 <- read.delim(
  "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Data/PanTHERIA_1-0_WR93_Aug2008.txt",
  header = TRUE,
  stringsAsFactors = FALSE
)

pan93_subset <- pan93 %>%
  mutate(
    Species = MSW93_Binomial,
    Order = MSW93_Order,
    Maximum.longevity..yrs. = ifelse(X17.1_MaxLongevity_m > 0, X17.1_MaxLongevity_m / 12, NA),
    Body.mass..g. = ifelse(X5.1_AdultBodyMass_g > 0, X5.1_AdultBodyMass_g, NA),
    Data.quality = NA,
    Specimen.origin = NA
  ) %>%
  select(Species, Order, Maximum.longevity..yrs., Body.mass..g., Data.quality, Specimen.origin)

# Import Amniote-Life-History-Database
amniote <- read.csv(
  "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Data/Amniote-Life-History-Database.csv",
  header = TRUE,
  stringsAsFactors = FALSE
)

amniote_clean <- amniote %>%
  mutate(
    adult_body_mass_g = as.numeric(gsub("Quantity\\[|\\]|\"Grams\"|,.*", "", adult_body_mass_g)),
    adult_body_mass_g = ifelse(grepl("Missing", adult_body_mass_g), NA, adult_body_mass_g),
    maximum_longevity_y = as.numeric(gsub("Quantity\\[|\\]|\"Years\"|,.*", "", maximum_longevity_y)),
    maximum_longevity_y = ifelse(grepl("Missing", maximum_longevity_y), NA, maximum_longevity_y)
  )

amniote_subset <- amniote_clean %>%
  mutate(
    Species = paste(genus, species),
    Order = order,
    Maximum.longevity..yrs. = ifelse(maximum_longevity_y > 0, maximum_longevity_y, NA),
    Body.mass..g. = ifelse(adult_body_mass_g > 0, adult_body_mass_g, NA),
    Data.quality = NA,
    Specimen.origin = NA
  ) %>%
  select(Species, Order, Maximum.longevity..yrs., Body.mass..g., Data.quality, Specimen.origin)

# Import bodymass and longevity
bodymass_data_ds <- read.csv(
  "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Data/Bodymass.csv",
  header = TRUE,
  stringsAsFactors = FALSE
)
bodymass_data_ds<-bodymass_data_ds %>% 
  rename(Body.mass..g.=BodyMass,
         Maximum.longevity..yrs.=Longevity)
# Merge all datasets
all_traits <- bind_rows(datage_subset, pan05_subset, pan93_subset, amniote_subset,bodymass_data_ds)

# Getting the species list 
specie_bm_ab <-unique(all_traits$Species)

# Run all names at once
res_specie_bm_ab <- name_backbone_checklist(name = specie_bm_ab)

# Extract orders
sp_orders_ab <- as.data.frame(res_specie_bm_ab[,c("canonicalName","order")])

# Match the order with the species in the original dataset. 

all_traits_ab<-left_join(all_traits,sp_orders_ab,by=c("Species"="canonicalName")) %>% 
  dplyr::select(-Order)

# Average by species
all_traits_avg <- all_traits_ab %>%
  # Count how many times each species appears
  group_by(Species) %>%
  mutate(species_count = n()) %>%
  ungroup() %>%
  # For species with duplicates, compute mean per order
  group_by(order, Species) %>%
  summarise(
    Maximum.longevity..yrs. = if (first(species_count) > 1) {
      mean(Maximum.longevity..yrs., na.rm = TRUE)
    } else {
      first(Maximum.longevity..yrs.)
    },
    Body.mass..g. = if (first(species_count) > 1) {
      mean(Body.mass..g., na.rm = TRUE)
    } else {
      first(Body.mass..g.)
    },
    .groups = "drop"
  ) %>%
  # Replace NaN (all NAs) with NA
  mutate(
    Maximum.longevity..yrs. = ifelse(is.nan(Maximum.longevity..yrs.), NA, Maximum.longevity..yrs.),
    Body.mass..g. = ifelse(is.nan(Body.mass..g.), NA, Body.mass..g.)
  )

```

```{r}
# Merge with main dataset (databa)
databuncpLS <- left_join(databa, all_traits_avg, by = "Species",
  relationship = "many-to-many")

# 8. Count species with traits
trait_counts <- databuncpLS %>%
  distinct(Species, .keep_all = TRUE) %>%
  summarise(
    n_species = n(),
    n_longevity = sum(!is.na(Maximum.longevity..yrs.)),
    n_bodymass = sum(!is.na(Body.mass..g.)),
    n_both = sum(!is.na(Maximum.longevity..yrs.) & !is.na(Body.mass..g.))
  )
trait_counts
```

```{r}
# Prepare data
databuncpLS_clean <- databuncpLS %>%
  filter(!is.na(Maximum.longevity..yrs.), !is.na(Body.mass..g.)) %>%
  distinct(Species, .keep_all = TRUE) %>%
  mutate(Body.mass_kg = Body.mass..g. / 1000)  # convert g to kg

# Fit mixed-effects model
fit_mixed_bm_ls <- lmer(log10(Maximum.longevity..yrs.) ~ log10(Body.mass_kg) + (1 | order), data = databuncpLS_clean)
summary(fit_mixed_bm_ls)
# Extract fixed-effect slope and intercept
fixef_coef <- fixef(fit_mixed_bm_ls)
intercept_fixed <- fixef_coef[1]
slope_fixed <- fixef_coef[2]

# Predict overall fixed effect line with CI
pred_df <- data.frame(
  Body.mass_kg = seq(min(databuncpLS_clean$Body.mass_kg), max(databuncpLS_clean$Body.mass_kg), length.out = 100)
)
pred_df$log10Longevity <- predict(fit_mixed_bm_ls, newdata = pred_df, re.form = NA)
pred_df$log10Longevity_upper <- pred_df$log10Longevity + 1.96 * summary(fit_mixed_bm_ls)$sigma
pred_df$log10Longevity_lower <- pred_df$log10Longevity - 1.96 * summary(fit_mixed_bm_ls)$sigma

# Get random-effect intercepts per Order
ranef_df <- ranef(fit_mixed_bm_ls)$order %>%
  tibble::rownames_to_column(var = "order") %>%
  rename(rand_intercept = `(Intercept)`)

# Generate lines for each Order (random intercept)
order_lines <- ranef_df %>%
  mutate(
    slope = slope_fixed
  ) %>%
  tidyr::crossing(Body.mass_kg = pred_df$Body.mass_kg) %>%
  mutate(log10Longevity = slope * log10(Body.mass_kg) + intercept_fixed + rand_intercept)


# # Plot
# # Plot
# Fig_bodymass_longevity <- ggplot(dat_clean, aes(x = Body.mass_kg, y = Maximum.longevity..yrs.)) +
#   geom_point(alpha = 0.6, color = "steelblue") +
#   
#   # Random-intercept lines for each Order
#   geom_line(data = order_lines, aes(x = Body.mass_kg, y = 10^log10Longevity, group = order),
#             color = "gray70", alpha = 0.6) +
#   
#   # Overall fixed-effect line
#   geom_line(data = pred_df, aes(x = Body.mass_kg, y = 10^log10Longevity),
#             color = "darkred", size = 1.2) +
#   
#   # Fixed-effect CI (note inherit.aes = FALSE)
#   geom_ribbon(data = pred_df, aes(x = Body.mass_kg, ymin = 10^log10Longevity_lower, ymax = 10^log10Longevity_upper),
#               fill = "red", alpha = 0.2, inherit.aes = FALSE) +
#   
#   scale_x_log10() +
#   scale_y_log10() +
#   labs(
#     x = "Body mass (kg, log scale)",
#     y = "Maximum longevity (years, log scale)"
#   ) +
#   theme_light()

# Export PDF
#ggsave("bodymass_longevity_mixed_effect_CI.pdf", plot = p, width = 8, height = 6)


```


```{r}
# Fit simple linear regression (log-log)
fit <- lm(log10(Maximum.longevity..yrs.) ~ log10(Body.mass_kg), data =databuncpLS_clean )
summary_fit <- summary(fit)

# Extract slope and p-value
slope <- round(summary_fit$coefficients[2, 1], 3)
pval <- signif(summary_fit$coefficients[2, 4], 3)

# Scatter plot with regression line + 95% CI
p_simple <- ggplot(dat_clean, aes(x = Body.mass_kg, y = Maximum.longevity..yrs.)) +
  geom_point(alpha = 0.6, color = "steelblue") +
  geom_smooth(method = "lm", se = TRUE, color = "darkred") +
  scale_x_log10() +
  scale_y_log10() +
  labs(
    x = "Body mass (kg, log scale)",
    y = "Maximum longevity (years, log scale)"
  ) +
  annotate("text", x = 1, y = 1, label = paste0("Slope = ", slope, "\nP = ", pval), hjust = 0, size = 4) +
  theme_light()

# Export PDF
# ggsave("/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/bodymass_longevity.pdf", plot = p_simple, width = 5, height = 5)

```

```{r}
# Filter data and resolve species names
abunadat <- databuncpLS_clean %>%
  filter(!is.na(Body.mass..g.) & !is.na(yi)) %>%
  mutate(species2 = Species)

# Resolve names via NCBI
Species_bm<-unique(abunadat$species2)

# Function to safely resolve a single species
safe_verify <- function(sp) {
  res <- tryCatch(
    gna_verifier(sp, best_match_only = TRUE, canonical = TRUE),
    error = function(e) return(data.frame(submitted_name = sp, matched_name = NA))
  )
  # If multiple rows returned, take the first (best match)
  if(nrow(res) > 1) res <- res[1, , drop = FALSE]
  return(res)
}

# Apply to all unique species
resolved_list <- lapply(Species_bm, safe_verify)

# Combine results into one data.frame
resolved_ab <- do.call(rbind, resolved_list)

# Inspect results
head(resolved_ab)

abunadat <- abunadat %>%
  left_join(resolved_ab[, c("submittedName", "matchedCanonicalSimple")],
            by = c("species2" = "submittedName")) %>%
  rename(Species_resolved = matchedCanonicalSimple)

# Retrieve taxonomy and build phylogenetic tree
sp_classif_a <- classification(unique(abunadat$Species_resolved), db = "ncbi")
abaphy <- class2tree(sp_classif_a)$phylo
# write.nexus(abaphy, file = "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Data/Phylogeny/abaphy.nex")

```

```{r}
# Clean tree and align with data
abunadat <- abunadat %>%
  filter(Species %in% abaphy$tip.label)
abunacleanTree <- drop.tip(abaphy, abaphy$tip.label[!abaphy$tip.label %in% abunadat$Species])
A <- ape::vcv.phylo(abunacleanTree)

# Fit phylogenetic mixed model using log10(Body.mass..g.)
bodymass_model <- brm(
  yi | se(se, sigma = TRUE) ~ scale(log10(Body.mass..g.)) + (1|ID) + (1|gr(Species, cov = A)),
  data = abunadat,
  family = gaussian(),
  data2 = list(A = A),
  prior = c(
    prior(normal(0, 1), class = "b"),
    prior(normal(0, 0.5), class = "Intercept"),
    prior(student_t(3, 0, 0.2), class = "sd"),
    prior(student_t(3, 0, 0.2), class = "sigma")
  ),
  iter = 5000,
  cores = 4,
  control = list(adapt_delta = 0.95)  # increase from default 0.8
)

summary(bodymass_model)

# 5. Plot relationship with regression line and CI
abunadat$logBodyMass <- log10(abunadat$Body.mass..g.)

preds <- fitted(bodymass_model, newdata = abunadat, re_formula = NA, summary = TRUE)
abunadat$fit <- preds[, "Estimate"]
abunadat$fit_lower <- preds[, "Q2.5"]
abunadat$fit_upper <- preds[, "Q97.5"]

pdf("/Users/jm200/Library/CloudStorage/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/Figure4_bodymass.pdf", width = 5, height = 4)
ggplot(abunadat, aes(x = logBodyMass, y = yi)) +
  geom_point(color = "orange2", size = 2.5, alpha = 0.6) +
  #geom_ribbon(aes(ymin = fit_lower, ymax = fit_upper), fill = "gray80", alpha = 0.4) +
  #geom_line(aes(y = fit), color = "darkred", size = 1) +
  #geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  labs(x = "Log10 Body mass (g)", y = "Effect size") +
  theme_light(base_size = 10) +
  theme(panel.border = element_rect(fill = NA),
        axis.text.y = element_text(angle = 90, hjust = 0.5))
dev.off()

```


## Pollination mode
### Pollination and population growth rate
```{r}
Pollinationdata <- read.csv("/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Data/Pollinationinfo.csv", header = T)
Pollinationdata$Pollination <- gsub(" ", "",Pollinationdata$Pollination)
datlambdaPl<-left_join(x=datlambda,y=Pollinationdata,by="Species")
datlambdaPl <- datlambdaPl[!apply(datlambdaPl[,c("Pollination", "vi")], 1, anyNA),] 
```

```{r}
sp.lambdaplant <- datlambdaPl[, 5]
spnamesplant <- as.character(sp.lambdaplant)
# Retrieve classifications
resolved_namesp <- classification(spnamesplant, db = "ncbi")
resolved_namesp <- unique(resolved_namesp)
# lapply(resolved_namesp, head)
# Building a phylogenetic tree
lambtree <- class2tree(resolved_namesp)
#str(gentree)
(lambtreephy <- lambtree$phylo)
```

```{r}
# 1. Create a data frame of tip labels
spnamesplant <- data.frame(Species = lambtreephy$tip.label)

# 2. Merge with your lambda data (keeping all tips from the tree)
podypdat <- merge(x = spnamesplant, y = datlambdaPl, by = "Species", all.x = TRUE)

# 3. Keep only complete cases
podypdat <- podypdat[complete.cases(podypdat), ]

# 4. Drop tips from the tree that are not in the cleaned dataset
tips_to_drop <- setdiff(lambtreephy$tip.label, podypdat$Species)
lambdcleanTree <- drop.tip(lambtreephy, tips_to_drop)

# 5. Compute the phylogenetic variance-covariance matrix
A <- ape::vcv.phylo(lambdcleanTree)

```

```{r}
datlambdaPl$Species <- factor(datlambdaPl$Species)
# 2. Reorder A to match the levels of Species
# Keep only species present in both the data and the phylo matrix
species_in_both <- intersect(datlambdaPl$Species, rownames(A))
datlambdaPl <- datlambdaPl %>% filter(Species %in% species_in_both)

# Reorder the factor levels to match the covariance matrix
datlambdaPl$Species <- factor(datlambdaPl$Species, levels = species_in_both)

# Reorder A to match the data
A <- A[species_in_both, species_in_both]

Pollination.lambmodel <- brms::brm(
  yi|se(se, sigma = TRUE) ~ (Pollination-1) + (1|ID) + (1|gr(Species, cov = A)),
  data = datlambdaPl,
  family = gaussian(),
  data2 = list(A = A),
  prior = c(
    prior(normal(0, 1), "b"),
    prior(student_t(3, 0, 5), "sd"),
    prior(student_t(3, 0, 5), "sigma")
  ),
  control = list(adapt_delta = 0.9999, max_treedepth = 15),
                   iter = 5000, cores = 4
)
  
```

```{r}
summary(Pollination.lambmodel)
```

#### empirical cumulative distribution function
```{r}
get_variables(Pollination.lambmodel)
post.samples_p_lambda <- posterior_samples(Pollination.lambmodel, pars=c("b_Pollinationout","b_Pollinationself"))
mean(post.samples_p_lambda$b_Pollinationout>0)
mean(post.samples_p_lambda$b_Pollinationself>0)
smd.ecdf_p_lambda_out <- ecdf(post.samples_p_lambda$b_Pollinationout)
smd.ecdf_p_lambda_self <- ecdf(post.samples_p_lambda$b_Pollinationself)
smd.ecdf_p_lambda_out(0)
smd.ecdf_p_lambda_self(0)
```

### Pollination and abundance
```{r}
databuncp %>%
  filter(Kingdom == "Plant")->databp
databuncpPl<-left_join(x=databp,y=Pollinationdata,by="Species")
databuncpPl$Pollination <- gsub("Self", "self",databuncpPl$Pollination)
# unique(databuncpPl$Pollination)
databuncpPl <- databuncpPl[!apply(databuncpPl[,c("Pollination", "vi")], 1, anyNA),]
databuncpPl %>%
  group_by(Pollination) %>%
  summarize(n_species = n_distinct(Species))
# unique(databuncpPl$Species[is.na(databuncpPl$Pollination)])
```

```{r,eval=FALSE}
sp.abp <- databuncpPl[, 5]
spnamesabp <- as.character(sp.abp)
species2 <- unique(spnamesabp)
#length(species2)
species2 <- trimws(gsub("\n", "", species2))  # clean up
batch_size <- 50  # keep well under the 100 limit
species_batches <- split(species2, ceiling(seq_along(species2) / batch_size))

# Query each batch
results <- lapply(species_batches, function(batch) {
  taxize::gna_verifier(batch, best_match_only = TRUE, canonical = TRUE)
})

# Combine all results
temp <- bind_rows(results)

dat_abp <- merge(data.frame(species2), temp[,c("submittedName", "matchedCanonicalSimple")], 
               by.x = "species2", by.y = "submittedName", all.x = TRUE)
spnamesabp <-dat_abp[,2]
# spnamesabp <- spnamesabp[complete.cases(spnamesabp)]
# length(spnamesabp)
# Retrieve classifications
spnamesabpf <- classification(spnamesabp, db = "ncbi")
resolved_names_abp <- unique(spnamesabpf)
# Building a phylogenetic tree
abptree <- class2tree(resolved_names_abp)
# str(abptree)
(abpphy <- abptree$phylo)
# length(abpphy$edge.length)
# Save to NEXUS file
# write.nexus(abpphy, file = "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Data/Phylogeny/abpphy.nex")
```

```{r}
# Create a data frame of tip labels
speciesabunp <- data.frame(Species = abpphy$tip.label)

# Merge with your abundance data (keeping all tips from the tree)
abunpdat <- merge(x = speciesabunp, y = databuncpPl, by = "Species", all.x = TRUE)

# Keep only complete cases
abunpdat <- abunpdat[complete.cases(abunpdat), ]

# Drop tips from the tree that are not in the cleaned dataset
tips_to_drop <- setdiff(abpphy$tip.label, abunpdat$Species)
abunpcleanTree <- drop.tip(abpphy, tips_to_drop)

# Compute the phylogenetic variance-covariance matrix
A <- ape::vcv.phylo(abunpcleanTree)

```

```{r}
Pollination.abunmodel <- brms::brm(
  yi|se(se, sigma = TRUE) ~ (Pollination-1) + (1|ID) + (1|gr(Species, cov = A)),
  data = abunpdat,
  family = gaussian(),
  data2 = list(A = A),
  prior = c(
    prior(normal(0, 1), "b"),
    prior(student_t(3, 0, 1.5), "sd"),
    prior(student_t(3, 0, 1.5), "sigma")
  ), 
  control = list(adapt_delta = 0.9999, max_treedepth = 15),iter = 5000, cores = 4
)
```

```{r}
summary(Pollination.abunmodel)
```

#### empirical cumulative distribution function
```{r}
get_variables(Pollination.abunmodel)
post.samples_p_ab <- posterior_samples(Pollination.abunmodel, pars=c("b_Pollinationout","b_Pollinationself"))
mean(post.samples_p_ab$b_Pollinationout>0)
mean(post.samples_p_ab$b_Pollinationself>0)
smd.ecdf_p_ab_out <- ecdf(post.samples_p_ab$b_Pollinationout)
smd.ecdf_p_ab_self <- ecdf(post.samples_p_ab$b_Pollinationself)
smd.ecdf_p_ab_out(0)
smd.ecdf_p_ab_self(0)
```

### Pollination and genetic diversity 
```{r}
datgencp %>% 
  filter(Kingdom == "Plant")->datgenp
gencpPL<-left_join(x =datgenp, y =Pollinationdata ,by="Species")
# unique(gencpPL$Species[is.na(gencpPL$Pollination)])
```

```{r}
# lookup table for pollination systems
pollination_gen <- data.frame(
  Species = c("Zostera noltii",
              "Narcissus longispathus",
              "Vaccinium stamineum",
              "Caryocar brasiliense",
              "Arabidopsis thaliana",
              "Clarkia xantiana xantiana",
              "Cirsium canescens"),
  Pollination = c("out",     # Zostera noltii
                  "out",          # Narcissus longispathus
                  "out",     # Vaccinium stamineum
                  "out",          # Caryocar brasiliense
                  "self",           # Arabidopsis thaliana
                  "out",     # Clarkia xantiana xantiana
                  "out")     # Cirsium canescens (likely outcrossed)
)

# merge into your main dataset (replace df with your object name)
gencpPL <- gencpPL %>%
  left_join(pollination_gen, by = "Species") %>%
  mutate(Pollination = coalesce(Pollination.y, Pollination.x)) %>%
  select(-Pollination.x, -Pollination.y)

```


```{r,eval=FALSE}
sp.genp <- gencpPL[, 5]
spnamesp <- as.character(sp.genp)
# Retrieve classifications
resolved_namesp <- classification(spnamesp, db = "ncbi")
resolved_namesp <- unique(resolved_namesp)
# lapply(resolved_namesp, head)
# Building a phylogenetic tree
genptree <- class2tree(resolved_namesp)
#str(gentree)
(genpphy <- genptree$phylo)
# plot(genpphy)
# write.nexus(genpphy, file = "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Data/Phylogeny/genpphy.nex")
```

```{r}
# Create a data frame of tip labels from the phylogenetic tree
speciesgen <- data.frame(Species = genpphy$tip.label)

# Merge with your trait data, keeping all species from the tree
genpdat <- merge(x = speciesgen, y = gencpPL, by = "Species", all.x = TRUE)

# Keep only complete cases
genpdat <- genpdat[complete.cases(genpdat), ]

# Drop tips from the tree that are not present in the cleaned dataset
tips_to_drop <- setdiff(genpphy$tip.label, genpdat$Species)
genpcleanTree <- drop.tip(genpphy, tips_to_drop)

# Compute the phylogenetic variance-covariance matrix
A <- ape::vcv.phylo(genpcleanTree)

# Reorder A to match the species order in the data (important for brms)
genpdat$Species <- factor(genpdat$Species, levels = rownames(A))
A <- A[levels(genpdat$Species), levels(genpdat$Species)]

```

```{r}
set.seed(13)
Pollination.genmodel <- brms::brm(
  yi|se(se, sigma = TRUE) ~ (Pollination-1) + (1|ID) + (1|gr(Species, cov = A)),
  data = genpdat,
  family = gaussian(),
  data2 = list(A = A),
  prior = c(
    prior(normal(0, 1), "b"),
    prior(student_t(3, 0, 3), "sd"),
    prior(student_t(3, 0, 3), "sigma")
  ),
  control = list(adapt_delta = 0.9999, max_treedepth = 15),
                   iter = 5000, cores = 4
)
  
```

```{r}
summary(Pollination.genmodel)
```

#### empirical cumulative distribution function
```{r}
get_variables(Pollination.genmodel)
post.samples_p_gen <- posterior_samples(Pollination.genmodel, pars=c("b_Pollinationout","b_Pollinationself"))
mean(post.samples_p_gen$b_Pollinationout>0)
mean(post.samples_p_gen$b_Pollinationself>0)
smd.ecdf_p_gen_out <- ecdf(post.samples_p_gen$b_Pollinationout)
smd.ecdf_p_gen_self <- ecdf(post.samples_p_gen$b_Pollinationself)
smd.ecdf_p_gen_out(0)
smd.ecdf_p_gen_self(0)
# We see that with 0.1791, the probability of our pooled effect being smaller than 0 for out 
# We see that with 0.1979, the probability of our pooled effect being smaller than 0 for self 
```

### Plot Pollination and population growth rate, population genetic and population abundance

```{r}
out_f_pollambda1 <- spread_draws(Pollination.abunmodel, b_Pollinationout) %>% 
  mutate(study = "Cross\n pollinated") %>% 
  dplyr::rename (Pollination=b_Pollinationout)
# Data frame of summary numbers
out_all_sum_pollambda1 <- group_by(out_f_pollambda1, study) %>% 
  mean_qi(Pollination)



out_f_pollambda2 <- spread_draws(Pollination.abunmodel, b_Pollinationself) %>% 
  mutate(study = "Self\n pollinated") %>% 
  dplyr::rename (Pollination=b_Pollinationself)
# Data frame of summary numbers
out_all_sum_pollambda2 <- group_by(out_f_pollambda2, study) %>% 
  mean_qi(Pollination)

out_allpol_lambda <- bind_rows(out_f_pollambda1, out_f_pollambda2) 
out_all_sumpollmbda <-bind_rows(out_all_sum_pollambda1,out_all_sum_pollambda2) 

Figure5a<-out_allpol_lambda %>%   
  ggplot(aes(Pollination, study,fill = after_stat(x < 0))) +
  labs(x = "Effect size (population growth rate)", y = '')+
  xlim(-2,2)+
  # Zero!
  geom_vline(xintercept = 0, linewidth = .25, lty = 2) +
  stat_halfeye(.width = c(0.95, 0.95)) +
  scale_fill_manual(values = c("#F0E442", "#999999"))+
  # Add text labels
  geom_text(size = 3,
    data = mutate_if(out_all_sumpollmbda, is.numeric, round, 2),
    aes(label = str_glue("{Pollination} [{.lower}, {.upper}]"), x =-1.8),
    hjust = "inward"
  ) +
  # annotate("text", x =-3,y=2.5, label = "a", color = "black",  size = 5) +
  # annotate("text", x = 3,y=1, label = "n=12", color = "black",  size = 3) +
  # annotate("text", x =3,y=2, label = "n=5", color = "black",  size = 3) +
  theme(panel.border = element_rect(fill = NA),
             text = element_text(size = 12), # change font sizes
                 legend.position = "none",
                 legend.title = element_text(size = 8),
                 legend.text = element_text(size = 8),
                 legend.background = element_blank(),
                 panel.background = element_rect(fill="transparent"), 
                 axis.text.x = element_text(size = 12, colour = "black"),
                 axis.text.y = element_text(size = 12, colour = "black", hjust = 0.5, angle = 0))
```


```{r}
out_f_polgen1 <- spread_draws(Pollination.genmodel, b_Pollinationout) %>% 
  mutate(study = "Cross\n pollinated") %>% 
  dplyr::rename (Pollination=b_Pollinationout)
# Data frame of summary numbers
out_all_sum_polgen1 <- group_by(out_f_polgen1, study) %>% 
  mean_qi(Pollination)

out_f_polgen2 <- spread_draws(Pollination.genmodel, b_Pollinationself) %>% 
  mutate(study = "Self\n pollinated") %>% 
  dplyr::rename (Pollination=b_Pollinationself)
# Data frame of summary numbers
out_all_sum_polgen2 <- group_by(out_f_polgen2, study) %>% 
  mean_qi(Pollination)

out_allpol_gen <- bind_rows(out_f_polgen1, out_f_polgen2) 
out_all_sumpolgen <-bind_rows(out_all_sum_polgen1,out_all_sum_polgen2) 

Figure5b<-out_allpol_gen %>%   
  ggplot(aes(Pollination, study,fill = after_stat(x < 0))) +
  labs(x = "Effect size (Genetic diversity)", y = "")+
   xlim(-3.5,3)+
  # Zero!
  geom_vline(xintercept = 0, linewidth = .25, lty = 2) +
  stat_halfeye(.width = c(0.95, 0.95)) +
  scale_fill_manual(values = c("#F0E442", "#999999"))+
  # Add text labels
  geom_text(size = 3,
    data = mutate_if(out_all_sumpolgen, is.numeric, round, 2),
    aes(label = str_glue("{Pollination} [{.lower}, {.upper}]"), x =-3.45),
    hjust = "inward"
  ) +
  # annotate("text", label = "b", x =-4,y=2.5, size = 5)+
  # annotate("text", x = 3,y=1, label = "n=40", color = "black",  size = 3) +
  # annotate("text", x =3,y=2, label = "n=7", color = "black",  size = 3) +
  theme(panel.border = element_rect( fill=NA),
             text = element_text(size = 12), # change font sizes
                 legend.position = "none",
                 legend.title = element_text(size = 8),
                 legend.text = element_text(size = 8),
                 legend.background = element_blank(),
                 panel.background = element_rect(fill="transparent"), 
                 axis.text.x = element_text(size = 12, colour = "black"),
                 axis.text.y = element_text(size = 12, colour = "black", hjust = 0.5, angle = 0))
```


```{r}
out_f_polabun1 <- spread_draws(Pollination.abunmodel, b_Pollinationout) %>% 
  mutate(study = "Cross\n pollinated") %>% 
  dplyr::rename (Pollination=b_Pollinationout)
# Data frame of summary numbers
out_all_sum_polabund1 <- group_by(out_f_polabun1, study) %>% 
  mean_qi(Pollination)

out_f_polabund2 <- spread_draws(Pollination.abunmodel, b_Pollinationself) %>% 
  mutate(study = "Self\n pollinated") %>% 
  dplyr::rename (Pollination=b_Pollinationself)
# Data frame of summary numbers
out_all_sum_polabun2 <- group_by(out_f_polabund2, study) %>% 
  mean_qi(Pollination)

out_allpol_abun <- bind_rows(out_f_polabun1, out_f_polabund2) 
out_all_sumpolabun <-bind_rows(out_all_sum_polabund1,out_all_sum_polabun2) 

Figure5c<-out_allpol_abun %>%   
  ggplot(aes(Pollination, study,fill = after_stat(x < 0))) +
  labs(x = "Effect size (abundance)", y = '')+
   xlim(-2,2)+
  # Zero!
  geom_vline(xintercept = 0, linewidth = .25, lty = 2) +
  stat_halfeye(.width = c(0.95, 0.95)) +
   scale_fill_manual(values = c("#F0E442", "#999999"))+
  # Add text labels
  geom_text(size = 3,
    data = mutate_if(out_all_sumpolabun, is.numeric, round, 2),
    aes(label = str_glue("{Pollination} [{.lower}, {.upper}]"), x =-1.8),
    hjust = "inward"
  ) +
  # annotate("text", label = "c", x =-2.5,y=2.5, size = 5)+
  # annotate("text", x = 2,y=1, label = "n=324", color = "black",  size = 3) +
  # annotate("text", x =2,y=2, label = "n=13", color = "black",  size = 3) +
  theme(panel.border = element_rect(fill=NA),
             text = element_text(size = 12), # change font sizes
                 legend.position = "none",
                 legend.title = element_text(size = 8),
                 legend.text = element_text(size = 8),
                 legend.background = element_blank(),
                 panel.background = element_rect(fill="transparent"), 
                 axis.text.x = element_text(size = 12, colour = "black"),
                 axis.text.y = element_text(size = 12, colour = "black", hjust = 0.5, angle = 0))

```

```{r}

# Arrange plots
pdf("/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/Figure5.pdf", width = 14, height = 5)
Figure5 <- (Figure5a + Figure5b + Figure5c) +
  plot_annotation(
    tag_levels = 'a'  # automatically assigns 'a', 'b', 'c'
  ) &
  theme(
    plot.tag = element_text(size = 16, face = "plain"),
    plot.tag.position = c(0.1, 1.0)  # top-left corner
  )

Figure5
dev.off()
```

```{r}
# Load packages

# Get unique species list
species_list <- as.character(unique(datlambda$Species))

# Retrieve all traits available for these species
all_traits <- BIEN_trait_species(species_list)

library(dplyr)
library(tidyr)
library(stringr)

# Function to calculate mode
get_mode <- function(x) {
  ux <- unique(x[!is.na(x)])
  if(length(ux) == 0) return(NA)
  ux[which.max(tabulate(match(x, ux)))]
}

# Clean trait values
all_traits_clean <- all_traits %>%
  mutate(trait_value_clean = str_trim(trait_value))

# Detect numeric traits
numeric_traits <- all_traits_clean %>%
  mutate(is_numeric = !is.na(as.numeric(trait_value_clean))) %>%
  group_by(trait_name) %>%
  summarise(n_numeric = sum(is_numeric, na.rm = TRUE),
            n_total = n(),
            .groups = "drop") %>%
  filter(n_numeric / n_total > 0.5) %>%   # if >50% numeric, treat as numeric
  pull(trait_name)

# Split numeric and categorical traits
traits_numeric <- all_traits_clean %>%
  filter(trait_name %in% numeric_traits) %>%
  mutate(trait_value = as.numeric(trait_value_clean)) %>%
  group_by(scrubbed_species_binomial, trait_name) %>%
  summarise(trait_mean = mean(trait_value, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = trait_name, values_from = trait_mean)

traits_categorical <- all_traits_clean %>%
  filter(!trait_name %in% numeric_traits) %>%
  group_by(scrubbed_species_binomial, trait_name) %>%
  summarise(trait_mode = get_mode(trait_value_clean), .groups = "drop") %>%
  pivot_wider(names_from = trait_name, values_from = trait_mode)
traits_categorical$`whole plant growth form`
# Combine numeric and categorical traits
traits_combined <- traits_numeric %>%
  left_join(traits_categorical, by = "scrubbed_species_binomial")
view(traits_combined)
# Join with abundance data
datlambda_with_traits <- datlambda %>%
  left_join(traits_combined, by = c("Species" = "scrubbed_species_binomial"))

names(all_traits)
# Pivot to wide format
traits_wide <- all_traits %>%
  select(scrubbed_species_binomial, trait_name, trait_value) %>%   # select relevant columns
  pivot_wider(names_from = trait_name, values_from = trait_value)

# Inspect the wide dataframe
head(traits_wide)
view(traits_wide)


# Compute mean of each trait per species
traits_mean <- all_traits %>%
  select(scrubbed_species_binomial, trait_name, trait_value) %>%
  # Convert trait_value to numeric if possible
  mutate(trait_value = as.numeric(trait_value)) %>%
  group_by(scrubbed_species_binomial, trait_name) %>%
  summarise(trait_mean = mean(trait_value, na.rm = TRUE), .groups = "drop") %>%
  # Pivot to wide format
  pivot_wider(names_from = trait_name, values_from = trait_mean)

view(traits_mean)
```

```{r}
lambda_lifeform<-left_join(datlambda,traits_categorical, by=c("Species"="scrubbed_species_binomial"))
lambda_lifeform %>% 
  dplyr::select(ID,Species,`whole plant growth form`,yi,vi,se)->lambda_lifeform_meta
# Create a lookup table for species with NA growth forms
growth_forms <- c(
  "Erythranthe cardinalis" = "Herb",
  "Erythranthe lewisii"    = "Herb",
  "Cypripedium calcicola"  = "Herb",   # terrestrial orchid
  "Ascophyllum nodosum"    = "Alga",
  "Fucus serratus"         = "Alga",
  "Penanthes roanensis"    = "Herb",   # likely Prenanthes roanensis
  "Ranunculus austro-oreganus" = "Herb",
  "Thunbergia atacorensis "="Herb"
)

# Replace NA values in whole.plant.growth.form
lambda_lifeform_meta$`whole plant growth form` <- ifelse(
  is.na(lambda_lifeform_meta$`whole plant growth form`),
  growth_forms[lambda_lifeform_meta$Species],
  lambda_lifeform_meta$`whole plant growth form`
)

lambda_lifeform_meta_clean <- lambda_lifeform_meta %>%
  # rename column
  rename(growth_form = `whole plant growth form`) %>%
  # standardize capitalization
  mutate(growth_form = str_to_title(growth_form)) %>%
  # remove rows with Alga
  filter(growth_form != "Alga")

```

```{r}
# Ensure variables are set correctly
lambda_lifeform_meta_clean <- lambda_lifeform_meta_clean %>%
  mutate(
    Species = factor(Species),
    growth_form = factor(growth_form)  # categorical predictor
  )

# Keep only species present in both the data and the phylo matrix
species_in_both <- intersect(lambda_lifeform_meta_clean$Species, rownames(A))
lambda_lifeform_meta_clean <- lambda_lifeform_meta_clean %>%
  filter(Species %in% species_in_both)

# Reorder Species factor levels to match covariance matrix
lambda_lifeform_meta_clean$Species <- factor(
  lambda_lifeform_meta_clean$Species,
  levels = species_in_both
)

# Reorder phylogenetic covariance matrix
A <- A[species_in_both, species_in_both]

# Run the phylogenetic meta-regression model with growth_form
GrowthForm.lambmodel <- brms::brm(
  yi | se(se, sigma = TRUE) ~ (growth_form-1) + (1|gr(Species, cov = A)),
  data   = lambda_lifeform_meta_clean,
  family = gaussian(),
  data2  = list(A = A),
  prior  = c(
    prior(normal(0, 1), "b"),          # fixed effects for growth_form
    prior(student_t(3, 0, 5), "sd"),   # random effects
    prior(student_t(3, 0, 5), "sigma") # residual
  ),
  control = list(adapt_delta = 0.9999, max_treedepth = 15),
  iter = 5000, cores = 4
)

```

```{r}
summary(GrowthForm.lambmodel)
```

#### empirical cumulative distribution function
```{r}
get_variables(GrowthForm.lambmodel)
post.samples_p_form <- posterior_samples(GrowthForm.lambmodel, pars=c("b_growth_formHerb","b_growth_formShrub"))
mean(post.samples_p_form$b_growth_formHerb>0)
mean(post.samples_p_form$b_growth_formShrub>0)
smd.ecdf_p_lm_herb <- ecdf(post.samples_p_form$b_growth_formHerb)
smd.ecdf_p_lm_shurb <- ecdf(post.samples_p_form$b_growth_formShrub)
smd.ecdf_p_lm_herb(0)
smd.ecdf_p_lm_shurb(0)
# We see that with 0.4105, the probability of our pooled effect being smaller than 0 for out 
# We see that with 0.5283, the probability of our pooled effect being smaller than 0 for self 
```

### Plot growth form and population growth rate, population genetic and population abundance

```{r}
herb_lambda <- spread_draws(GrowthForm.lambmodel, b_growth_formHerb) %>% 
  mutate(study = "Herb") %>% 
  dplyr::rename (Form=b_growth_formHerb)
# Data frame of summary numbers
all_sum_lam_herb <- group_by(herb_lambda, study) %>% 
  mean_qi(Form)

shurb_ambda <- spread_draws(GrowthForm.lambmodel, b_growth_formShrub) %>% 
  mutate(study = "Shrub") %>% 
  dplyr::rename (Form=b_growth_formShrub)
# Data frame of summary numbers
all_sum_pollam_shurb <- group_by(shurb_ambda, study) %>% 
  mean_qi(Form)


tree_ambda <- spread_draws(GrowthForm.lambmodel, b_growth_formTree) %>% 
  mutate(study = "Tree") %>% 
  dplyr::rename (Form=b_growth_formTree)
# Data frame of summary numbers
all_sum_pollam_tree <- group_by(tree_ambda, study) %>% 
  mean_qi(Form)

all_lambda_form <- bind_rows(herb_lambda, shurb_ambda,tree_ambda) 
all_sumlambda_form <-bind_rows(all_sum_lam_herb,all_sum_pollam_shurb,all_sum_pollam_tree) 

Figure7a<-all_lambda_form %>%   
  ggplot(aes(Form, study,fill = after_stat(x < 0))) +
  labs(x = "Effect size (population growth rate)", y = '')+
  xlim(-4.5,3.5)+
  # Zero!
  geom_vline(xintercept = 0, linewidth = .25, lty = 2) +
  stat_halfeye(.width = c(0.95, 0.95)) +
  scale_fill_manual(values = c("#F0E442", "#999999"))+
  # Add text labels
  geom_text(size = 2.8,
    data = mutate_if(all_sumlambda_form, is.numeric, round, 2),
    aes(label = str_glue("{Form} [{.lower}, {.upper}]"), x =-4.5),
    hjust = "inward"
  ) +
  # annotate("text", x =-3,y=2.5, label = "a", color = "black",  size = 5) +
  # annotate("text", x = 3,y=1, label = "n=12", color = "black",  size = 3) +
  # annotate("text", x =3,y=2, label = "n=5", color = "black",  size = 3) +
  theme(panel.border = element_rect(fill = NA),
             text = element_text(size = 10), # change font sizes
                 legend.position = "none",
                 legend.title = element_text(size = 8),
                 legend.text = element_text(size = 8),
                 legend.background = element_blank(),
                 panel.background = element_rect(fill="transparent"), 
                 axis.text.x = element_text(size = 10, colour = "black"),
                 axis.text.y = element_text(size = 10, colour = "black", hjust = 0.5, angle = 0))
```



```{r}
lambda_seed <- datlambda %>%
  left_join(traits_mean, by = c("Species" = "scrubbed_species_binomial")) %>%
  # Rename seed_mass if needed
  rename(seed_mass = `seed mass`) %>%  # only if the column has a space
  # Remove rows with NA in seed_mass
  filter(!is.na(seed_mass))

```

```{r}
seed_model_phylo <- brm(
  yi | se(se, sigma = TRUE) ~ log10(seed_mass) + (1|gr(Species, cov = A)),
  data = lambda_seed,
  data2 = list(A = A),
  family = gaussian(),
  prior = c(
    prior(normal(0, 1), class = "b"),
    prior(normal(0, 0.5), class = "Intercept"),
    prior(student_t(3, 0, 0.2), class = "sigma")
  ),
  iter = 5000,
  cores = 4,
  control = list(adapt_delta = 0.99)
)
```

```{r}
summary(seed_model)
```

```{r}
# Fit the seed_mass model
seed_model <- brm(
  yi | se(se, sigma = TRUE) ~ log10(seed_mass),
  data = lambda_seed,
  family = gaussian(),
  prior = c(
    prior(normal(0, 1), class = "b"),
    prior(normal(0, 0.5), class = "Intercept"),
    prior(student_t(3, 0, 0.2), class = "sigma")
  ),
  iter = 5000,
  cores = 4,
  control = list(adapt_delta = 0.99)
)

summary(seed_model)

# 2. Add log10-transformed predictor for plotting
lambda_seed$logSeedMass <- log10(lambda_seed$seed_mass)

# 3. Get fitted values with credible intervals
seed_preds <- fitted(seed_model, newdata = lambda_seed, re_formula = NA, summary = TRUE)
lambda_seed <- lambda_seed %>%
  mutate(fit = seed_preds[, "Estimate"],
         fit_lower = seed_preds[, "Q2.5"],
         fit_upper = seed_preds[, "Q97.5"])


label_lamd_seed <- lambda_seed %>%
  summarize(
    minX = min(log10(seed_mass)),
    maxX = max(log10(seed_mass)),
    lower = min(fit_lower),
    upper = max(fit_upper)
  ) %>%
  mutate(label = str_glue("CI [{round(lower,2)}, {round(upper,2)}]"))

# Plot regression line with CI
pdf("/Users/jm200/Library/CloudStorage/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/Figure4_seedmass.pdf",
    width = 5, height = 4)

Fig7d<-ggplot(lambda_seed, aes(x = logSeedMass, y = yi)) +
  geom_point(color = "orange2", size = 2.5, alpha = 0.6) +
  # geom_ribbon(aes(ymin = fit_lower, ymax = fit_upper), fill = "gray80", alpha = 0.4) +
  # geom_line(aes(y = fit), color = "blue4", size = 1) +
  labs(x = "Log10 Seed mass", y = "Effect size (population growth rate)") +
  my_theme

dev.off()

```

## Genetic 
```{r}
species_list_gp <- unique(as.character(datgenp$Species))
all_traits_gp <- BIEN_trait_species(species_list_gp)
all_traits_gp_clean <- all_traits_gp %>%
  mutate(trait_value_clean = str_trim(trait_value))
```

```{r}
get_mode <- function(x) {
  ux <- unique(x[!is.na(x)])
  if(length(ux) == 0) return(NA)
  ux[which.max(tabulate(match(x, ux)))]
}

numeric_gp_traits <- all_traits_gp_clean %>%
  mutate(is_numeric = !is.na(as.numeric(trait_value_clean))) %>%
  group_by(trait_name) %>%
  summarise(n_numeric = sum(is_numeric, na.rm = TRUE),
            n_total = n(),
            .groups = "drop") %>%
  filter(n_numeric / n_total > 0.5) %>%
  pull(trait_name)

```

```{r}
# Split numeric and categorical traits
traits_numeric_g <- all_traits_gp_clean %>%
  filter(trait_name %in% numeric_traits) %>%
  mutate(trait_value = as.numeric(trait_value_clean)) %>%
  group_by(scrubbed_species_binomial, trait_name) %>%
  summarise(trait_mean = mean(trait_value, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = trait_name, values_from = trait_mean)
#view(traits_numeric_g)
```

```{r}
traits_gp_categorical <- all_traits_gp_clean %>%
  filter(!trait_name %in% numeric_gp_traits) %>%
  group_by(scrubbed_species_binomial, trait_name) %>%
  summarise(trait_mode = get_mode(trait_value_clean), .groups = "drop") %>%
  pivot_wider(names_from = trait_name, values_from = trait_mode)
# view(traits_gp_categorical)
```

```{r}
gen_lifeform <- left_join(datgenp, traits_gp_categorical, by = c("Species" = "scrubbed_species_binomial"))
gen_lifeform %>% 
  dplyr::select(ID,Species,`whole plant growth form`,yi,vi,se)->gen_lifeform_meta
# view(gen_lifeform_meta)
unique(gen_lifeform_meta$`whole plant growth form`)
```

```{r}
growth_forms <- c(
  "Cymodocea nodosa"          = "Herb",   # aquatic plant
  "Saccorhiza polyschides"    = "Alga",   # if needed, can exclude later
  "Campanulastrum americanum" = "Herb",
  "Cypripedium"               = "Herb",
  "Genista legionensis"       = "Shrub",
  "Saponaria bellidifolia"    = "Herb",
  "Neotinea maculata"         = "Herb",
  "Viola elatior"             = "Herb",
  "Viola pumila"              = "Herb",
  "Viola stagnina"            = "Herb",
  "Pinus contorta"            = "Tree",
  "Grevillea repens"          = "Shrub",
  "Anthyllis montana"         = "Shrub",
  "Carex magellanica"         = "Herb",
  "Epidendrum fulgens"        = "Herb"
)
```

```{r}
# Replace NA values in whole.plant.growth.form
gen_lifeform$`whole plant growth form` <- ifelse(
  is.na(gen_lifeform$`whole plant growth form`),
  growth_forms[gen_lifeform$Species],
  gen_lifeform$`whole plant growth form`
)

unique(gen_lifeform$`whole plant growth form`)
gen_lifeform_clean <- gen_lifeform %>%
  # rename column
  rename(growth_form = `whole plant growth form`) %>%
  # remove rows with unwanted growth forms and NAs
  filter(!is.na(growth_form) & !tolower(growth_form) %in% c("alga", "aquatic*","aquatic","forb")) %>%
  # standardize capitalization for values
  mutate(
    growth_form = case_when(
      tolower(growth_form) %in% c("tree") ~ "Tree",
      tolower(growth_form) %in% c("shrub") ~ "Shrub",
      tolower(growth_form) %in% c("herb") ~ "Herb",
      TRUE ~ growth_form
    ),
    # convert to factor
    growth_form = factor(growth_form)
  )

unique(gen_lifeform_clean$growth_form)


```

```{r}
# 1. Remove rows with missing growth_form
gen_lifeform_clean <- gen_lifeform_clean %>% filter(!is.na(growth_form))

# 2. Only keep species present in A
species_in_matrix <- intersect(gen_lifeform_clean$Species, rownames(A))
gen_lifeform_clean <- gen_lifeform_clean %>% filter(Species %in% species_in_matrix)

# 3. Reorder Species factor to match covariance matrix
gen_lifeform_clean$Species <- factor(gen_lifeform_clean$Species, levels = species_in_matrix)

# 4. Reorder phylogenetic covariance matrix
A <- A[species_in_matrix, species_in_matrix]

# 5. Fit the phylogenetic meta-regression
growthform_model_gp <- brm(
  yi | se(se, sigma = TRUE) ~ 0 + growth_form + (1|ID) + (1 | gr(Species, cov = A)),
  data = gen_lifeform_clean,
  family = gaussian(),
  data2 = list(A = A),
  prior = c(
    prior(normal(0, 1), class = "b"),
    prior(student_t(3, 0, 5), class = "sd"),
    prior(student_t(3, 0, 5), class = "sigma")
  ),
  control = list(adapt_delta = 0.9999, max_treedepth = 15),
  iter = 5000, cores = 4
)


```

```{r}
summary(growthform_model_gp)
```

```{r}
# Variables
get_variables(growthform_model_gp)

# Posterior samples for growth forms
post.samples_gp_form <- posterior_samples(
  growthform_model_gp,
  pars = c("b_growth_formHerb","b_growth_formShrub","b_growth_formTree")
)

# Posterior probabilities > 0
mean(post.samples_gp_form$b_growth_formHerb > 0)
mean(post.samples_gp_form$b_growth_formShrub > 0)
mean(post.samples_gp_form$b_growth_formTree > 0)

# ECDF for probabilities below 0
smd.ecdf_gp_herb  <- ecdf(post.samples_gp_form$b_growth_formHerb)
smd.ecdf_gp_shrub <- ecdf(post.samples_gp_form$b_growth_formShrub)
smd.ecdf_gp_tree  <- ecdf(post.samples_gp_form$b_growth_formTree)

smd.ecdf_gp_herb(0)
smd.ecdf_gp_shrub(0)
smd.ecdf_gp_tree(0)

# Plot for growth form effects

# Herb
herb_gp <- spread_draws(growthform_model_gp, b_growth_formHerb) %>%
  mutate(study = "Herb") %>%
  dplyr::rename(Form = b_growth_formHerb)
all_sum_gp_herb <- group_by(herb_gp, study) %>% mean_qi(Form)

# Shrub
shrub_gp <- spread_draws(growthform_model_gp, b_growth_formShrub) %>%
  mutate(study = "Shrub") %>%
  dplyr::rename(Form = b_growth_formShrub)
all_sum_gp_shrub <- group_by(shrub_gp, study) %>% mean_qi(Form)

# Tree
tree_gp <- spread_draws(growthform_model_gp, b_growth_formTree) %>%
  mutate(study = "Tree") %>%
  dplyr::rename(Form = b_growth_formTree)
all_sum_gp_tree <- group_by(tree_gp, study) %>% mean_qi(Form)

# Combine
all_gp_form <- bind_rows(herb_gp, shrub_gp, tree_gp)
all_sum_gp_form <- bind_rows(all_sum_gp_herb, all_sum_gp_shrub, all_sum_gp_tree)

# Plot
Figure7b <- all_gp_form %>%
  ggplot(aes(Form, study, fill = after_stat(x < 0))) +
  labs(x = "Effect size (genetic diversity)", y = '') +
  xlim(-4.5, 3) +
  geom_vline(xintercept = 0, linewidth = .25, lty = 2) +
  stat_halfeye(.width = c(0.95, 0.95)) +
  scale_fill_manual(values = c("#F0E442", "#999999")) +
  geom_text(
    size = 2.8,
    data = mutate_if(all_sum_gp_form, is.numeric, round, 2),
    aes(label = str_glue("{Form} [{.lower}, {.upper}]"), x = -4.5),
    hjust = "inward"
  ) +
  theme(
    panel.border     = element_rect(fill = NA),
    text             = element_text(size = 10),
    legend.position  = "none",
    legend.title     = element_text(size = 8),
    legend.text      = element_text(size = 8),
    legend.background= element_blank(),
    panel.background = element_rect(fill="transparent"),
    axis.text.x      = element_text(size = 10, colour = "black"),
    axis.text.y      = element_text(size = 10, colour = "black", hjust = 0.5, angle = 0)
  )

```

```{r}
traits_numeric_g$`seed mass`
# Compute mean of each trait per species
traits_g <- traits_numeric_g %>%
  select(scrubbed_species_binomial, `seed mass`,) %>%
  # Convert trait_value to numeric if possible
  mutate(seed_mass = as.numeric(`seed mass`)) %>% 
  na.omit()

```


```{r}
gen_seed <- gencpPL %>%
  left_join(traits_g, by = c("Species" = "scrubbed_species_binomial")) %>%
  # Remove rows with NA in seed_mass
  filter(!is.na(seed_mass))

```

```{r}
# Fit the seed_mass model
seed_model_g <- brm(
  yi | se(se, sigma = TRUE) ~ log10(seed_mass),
  data = gen_seed,
  family = gaussian(),
  prior = c(
    prior(normal(0, 1), class = "b"),
    prior(normal(0, 0.5), class = "Intercept"),
    prior(student_t(3, 0, 0.2), class = "sigma")
  ),
  iter = 5000,
  cores = 4,
  control = list(adapt_delta = 0.99)
)

summary(seed_model_g)

# 2. Add log10-transformed predictor for plotting
gen_seed$logSeedMass <- log10(gen_seed$seed_mass)

# 3. Get fitted values with credible intervals
seed_preds_g <- fitted(seed_model_g, newdata = gen_seed, re_formula = NA, summary = TRUE)
gen_seed <- gen_seed %>%
  mutate(fit = seed_preds_g[, "Estimate"],
         fit_lower = seed_preds_g[, "Q2.5"],
         fit_upper = seed_preds_g[, "Q97.5"])


# Plot regression line with CI
pdf("/Users/jm200/Library/CloudStorage/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/Figure4_seedmass.pdf",
    width = 5, height = 4)

Fig7f<-ggplot(gen_seed, aes(x = logSeedMass, y = yi)) +
  geom_point(color = "orange2", size = 2.5, alpha = 0.6) +
  geom_ribbon(aes(ymin = fit_lower, ymax = fit_upper), fill = "gray80", alpha = 0.4) +
  geom_line(aes(y = fit), color = "blue4", size = 1) +
  labs(x = "Log10 Seed mass", y = "Effect size (population growth rate)") +
  my_theme

dev.off()

```


```{r}
# Prepare trait data
species_list_ab <- unique(as.character(abunpdat$Species))

# Retrieve traits
all_traits_ab <- BIEN_trait_species(species_list_ab)

# Mode function for categorical traits
compute_mode <- function(x) {
  ux <- unique(x[!is.na(x)])
  if(length(ux) == 0) return(NA)
  ux[which.max(tabulate(match(x, ux)))]
}

# Clean trait values
all_traits_clean_ab <- all_traits_ab %>%
  mutate(trait_value_clean = str_trim(trait_value))

# Detect numeric traits
numeric_traits_ab <- all_traits_clean_ab %>%
  mutate(is_numeric = !is.na(as.numeric(trait_value_clean))) %>%
  group_by(trait_name) %>%
  summarise(n_numeric = sum(is_numeric, na.rm = TRUE),
            n_total = n(),
            .groups = "drop") %>%
  filter(n_numeric / n_total > 0.5) %>%
  pull(trait_name)

# Split numeric and categorical
traits_numeric_ab <- all_traits_clean_ab %>%
  filter(trait_name %in% numeric_traits_ab) %>%
  mutate(trait_value = as.numeric(trait_value_clean)) %>%
  group_by(scrubbed_species_binomial, trait_name) %>%
  summarise(trait_mean = mean(trait_value, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = trait_name, values_from = trait_mean)

traits_categorical_ab <- all_traits_clean_ab %>%
  filter(!trait_name %in% numeric_traits_ab) %>%
  group_by(scrubbed_species_binomial, trait_name) %>%
  summarise(trait_mode = compute_mode(trait_value_clean), .groups = "drop") %>%
  pivot_wider(names_from = trait_name, values_from = trait_mode)

# Combine numeric + categorical traits
traits_all_ab <- traits_numeric_ab %>%
  left_join(traits_categorical_ab, by = "scrubbed_species_binomial")

# Merge with abundance data
abund_traits_ab <- abunpdat %>%
  left_join(traits_all_ab, by = c("Species" = "scrubbed_species_binomial")) %>%
  rename(seed_mass_g = `seed mass`) %>%  # rename for clarity
  filter(!is.na(seed_mass_g))

# Fit seed mass meta-regression with Species random effect

```

```{r}
seed_mass_model_ab <- brm(
  yi | se(se, sigma = TRUE) ~ log10(seed_mass_g) + (1 | Species),
  data = abund_traits_ab,
  family = gaussian(),
  prior = c(
    prior(normal(0, 1), class = "b"),          # fixed effect
    prior(normal(0, 0.5), class = "Intercept"), 
    prior(student_t(3, 0, 0.2), class = "sd"), # random effect
    prior(student_t(3, 0, 0.2), class = "sigma") # residual
  ),
  iter = 5000,
  cores = 4,
  control = list(adapt_delta = 0.99)
)

summary(seed_mass_model_ab)

# Prepare fitted values for plotting
abund_traits_ab <- abund_traits_ab %>%
  mutate(logSeedMass = log10(seed_mass_g))

fitted_vals_ab <- fitted(seed_mass_model_ab, newdata = abund_traits_ab, re_formula = NA, summary = TRUE)

abund_traits_ab <- abund_traits_ab %>%
  mutate(fit = fitted_vals_ab[, "Estimate"],
         fit_lower = fitted_vals_ab[, "Q2.5"],
         fit_upper = fitted_vals_ab[, "Q97.5"])

# 5. Plot

pdf("/Users/jm200/Library/CloudStorage/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/Figure4_seedmass_ab_SpeciesRE.pdf",
    width = 5, height = 4)

Fig7e<-ggplot(abund_traits_ab, aes(x = logSeedMass, y = yi)) +
  geom_point(color = "orange2", size = 2.5, alpha = 0.6) +
  # geom_ribbon(aes(ymin = fit_lower, ymax = fit_upper), fill = "gray80", alpha = 0.4) +
  # geom_line(aes(y = fit), color = "navy", size = 1) +
  labs(x = "Log10 Seed mass (g)", y = "Effect size ( abundance)") +
  my_theme

dev.off()
```

```{r}
# 1. Merge abundance data with categorical traits
lifeform_ab <- left_join(abunpdat, traits_categorical_ab, 
                         by = c("Species" = "scrubbed_species_binomial"))

# Define growth forms for species with NA
growth_forms <- c(
  "Carya ovalis"         = "Tree",
  "Castanea pumila"      = "Tree",
  "Chrysolepis chrysophylla" = "Tree",
  "Crataegus crus-galli" = "Shrub",   # sometimes small tree
  "Ehretia anacua"       = "Tree",
  "Fagus sylvatica"      = "Tree",
  "Fraxinus caroliniana" = "Tree",
  "Fraxinus velutina"    = "Tree",
  "Gleditsia aquatica"   = "Tree",
  "Gordonia lasianthus"  = "Tree",
  "Halesia diptera"      = "Tree",
  "Juniperus ashei"      = "Tree",
  "Juniperus pinchotii"  = "Tree",
  "Pinus remota"         = "Tree",
  "Planera aquatica"     = "Tree",
  "Prunus americana"     = "Tree",
  "Prunus emarginata"    = "Tree",
  "Quercus agrifolia"    = "Tree",
  "Quercus grisea"       = "Tree",
  "Quercus hypoleucoides"= "Tree",
  "Quercus incana"       = "Tree",
  "Quercus marilandica"  = "Tree",
  "Quercus pagoda"       = "Tree",
  "Quercus shumardii"    = "Tree",
  "Quercus similis"      = "Tree",
  "Quercus sinuata"      = "Tree",
  "Sorbus americana"     = "Tree",
  "Sorbus decora"        = "Tree",
  "Torreya californica"  = "Tree",
  "Vernicia fordii"      = "Tree"
)

# Replace NA values in whole plant growth form
lifeform_ab$`whole plant growth form` <- ifelse(
  is.na(lifeform_ab$`whole plant growth form`),
  growth_forms[lifeform_ab$Species],
  lifeform_ab$`whole plant growth form`
)

lifeform_ab_meta <- lifeform_ab %>%
  dplyr::select(ID, Species, `whole plant growth form`, yi, vi, se)

lifeform_ab_meta_clean <- lifeform_ab_meta %>%
  rename(growth_form = `whole plant growth form`) %>%
  mutate(
    growth_form = str_to_title(growth_form),               # standardize capitalization
    growth_form = ifelse(growth_form %in% c("Shrub*", "Shrub"), "Shrub", growth_form)  # combine Shrub* and Shrub
  ) %>%
  filter(
    !is.na(growth_form),                                   # remove NAs
    growth_form != "Alga",                                 # remove Alga
    !grepl("Sclerophyll", growth_form)                     # remove Sclerophyll*
  ) %>%
  mutate(
    growth_form = factor(growth_form)                      # convert back to factor
  )

# Check result

unique(lifeform_ab_meta_clean$growth_form)
species_in_matrix <- intersect(lifeform_ab_meta_clean$Species, rownames(A))
lifeform_ab_meta_clean <- lifeform_ab_meta_clean %>%
  filter(Species %in% species_in_matrix)

# Reorder Species factor to match covariance matrix
lifeform_ab_meta_clean$Species <- factor(
  lifeform_ab_meta_clean$Species,
  levels = species_in_matrix
)

# Reorder phylogenetic covariance matrix
A <- A[species_in_matrix, species_in_matrix]


# Fit phylogenetic meta-regression with growth form
growthform_model_ab <- brm(
  yi | se(se, sigma = TRUE) ~ (growth_form - 1) + (1|ID) + (1 | gr(Species, cov = A)),
  data = lifeform_ab_meta_clean,
  family = gaussian(),
  data2 = list(A = A),
  prior = c(
    prior(normal(0, 1), class = "b"),          # fixed effects
    prior(student_t(3, 0, 5), class = "sd"),   # random effects
    prior(student_t(3, 0, 5), class = "sigma") # residual
  ),
  control = list(adapt_delta = 0.9999, max_treedepth = 15),
  iter = 5000, cores = 4
)
```

```{r}
summary(growthform_model_ab)
get_variables(growthform_model_ab)
```

```{r}
# Variables
get_variables(growthform_model_ab)

# Posterior samples for growth forms
post.samples_ab_form <- posterior_samples(
  growthform_model_ab,
  pars = c("b_growth_formHerb","b_growth_formShrub","b_growth_formTree")
)

# Posterior probabilities > 0
mean(post.samples_ab_form$b_growth_formHerb > 0)
mean(post.samples_ab_form$b_growth_formShrub > 0)
mean(post.samples_ab_form$b_growth_formTree > 0)

# ECDF for probabilities below 0
smd.ecdf_ab_herb  <- ecdf(post.samples_ab_form$b_growth_formHerb)
smd.ecdf_ab_shrub <- ecdf(post.samples_ab_form$b_growth_formShrub)
smd.ecdf_ab_tree  <- ecdf(post.samples_ab_form$b_growth_formTree)

smd.ecdf_ab_herb(0)
smd.ecdf_ab_shrub(0)
smd.ecdf_ab_tree(0)

# Plot for growth form effects

# Herb
herb_ab <- spread_draws(growthform_model_ab, b_growth_formHerb) %>%
  mutate(study = "Herb") %>%
  dplyr::rename(Form = b_growth_formHerb)
all_sum_ab_herb <- group_by(herb_ab, study) %>% mean_qi(Form)

# Shrub
shrub_ab <- spread_draws(growthform_model_ab, b_growth_formShrub) %>%
  mutate(study = "Shrub") %>%
  dplyr::rename(Form = b_growth_formShrub)
all_sum_ab_shrub <- group_by(shrub_ab, study) %>% mean_qi(Form)

# Tree
tree_ab <- spread_draws(growthform_model_ab, b_growth_formTree) %>%
  mutate(study = "Tree") %>%
  dplyr::rename(Form = b_growth_formTree)
all_sum_ab_tree <- group_by(tree_ab, study) %>% mean_qi(Form)

# Combine
all_ab_form <- bind_rows(herb_ab, shrub_ab, tree_ab)
all_sum_ab_form <- bind_rows(all_sum_ab_herb, all_sum_ab_shrub, all_sum_ab_tree)

# Plot
Figure7c <- all_ab_form %>%
  ggplot(aes(Form, study, fill = after_stat(x < 0))) +
  labs(x = "Effect size (abundance)", y = '') +
  xlim(-2.5, 2) +
  geom_vline(xintercept = 0, linewidth = .25, lty = 2) +
  stat_halfeye(.width = c(0.95, 0.95)) +
  scale_fill_manual(values = c("#F0E442", "#999999")) +
  geom_text(
    size = 2.8,
    data = mutate_if(all_sum_ab_form, is.numeric, round, 2),
    aes(label = str_glue("{Form} [{.lower}, {.upper}]"), x = -2.3),
    hjust = "inward"
  ) +
  theme(
    panel.border     = element_rect(fill = NA),
    text             = element_text(size = 10),
    legend.position  = "none",
    legend.title     = element_text(size = 8),
    legend.text      = element_text(size = 8),
    legend.background= element_blank(),
    panel.background = element_rect(fill="transparent"),
    axis.text.x      = element_text(size = 10, colour = "black"),
    axis.text.y      = element_text(size = 10, colour = "black", hjust = 0.5, angle = 0)
  )


```


```{r}
# Arrange plots
pdf("/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/Figure7.pdf", width = 8, height = 6)
Figure7 <- (Figure7a + Figure7b) / (Figure7c  + Fig7e)+
  plot_annotation(
    tag_levels = 'a'  # automatically assigns 'a', 'b', 'c'
  ) &
  theme(
    plot.tag = element_text(size = 14, face = "plain"),
    plot.tag.position = c(0, 1.0)  # top-left corner
  )

Figure7
dev.off()
```

# Population abundance 
```{r}
databuncp %>%
  filter(Kingdom == "Plant")->databp
species_list_ab<-databp$Species
missing_pollination_ab <- check_missing_species(species_list_ab, pollination_data, "Pollination_syndrome")
missing_fruit_mass_ab <- check_missing_species(species_list_ab, fruit_mass_data, "Fruit_dry_mass")
missing_growth_form_ab <- check_missing_species(species_list_ab, growth_form_data, "Plant_growth_form")

# Combine results
missing_all_ab <- bind_rows(missing_pollination_ab, missing_fruit_mass_ab, missing_growth_form_ab)
missing_all

```

```{r}
seedmass<-left_join(databp,fruit_mass_data,by=c("Species"="SpeciesName"))
na_per_column <- colSums(is.na(seedmass))
na_per_column
seedmass %>% 
  dplyr::select(ID,Species,yi,vi,se,OrigValueStr) %>% 
  na.omit()
```


