---
title: "Meta-analysis reveals that species are not most abundant or demographically performant at their range’s centre despite higher genetic diversity"
author: "Jacob Moutouama and Orou Gaoue"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    code_folding: hide
    self_contained: true
    number_sections: true
    thumbnails: true
    lightbox: true
    gallery: true
    keep_md: true
    highlight: tango
    df_print: kable
    toc_depth: 3
    fig_width: 8
    fig_height: 8
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(
  external = TRUE,
  echo = FALSE,
  warning = FALSE,
  fig.pos = "H",
  global.device = TRUE
)

# Plot size (A4 reference)
a4width <- 8.3
a4height <- 11.7

options(knitr.table.format = "latex")
set.seed(13)

rstan_options(auto_write = TRUE)

```

```{r}
rm(list = ls(all = TRUE))
```

```{r libraries}
## Packages
# Core data manipulation + plotting
library(tidyverse)
library(ggplot2)
library(ggridges)
library(glue)
library(forcats)

# Bayesian modeling
library(brms)
library(tidybayes)
library(posterior)
library(bayestestR)

# Meta-analysis tools
library(metafor)
library(effectsize)

# Spatial mapping
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)

# Tables
library(kableExtra)
library(flextable)

# Misc
library(rstan)

# Names
library(taxize)
library(rgbif)

# Phylogenie
library(ape)

# Species traits
library(BIEN)
```

```{r}
set.seed(13)
```

# Utility Functions
```{r functions}
Ftable <- function(model, reorder = FALSE, order = NULL, percent = FALSE) {
  fixed_table <- summary(model)$fixed
  fixed_90 <- summary(model, prob = 0.9)$fixed %>% as_tibble()

  # Clean and rename columns
  fixed_table <- fixed_table %>%
    as_tibble() %>%
    rename(
      L95CL = `l-95% CI`,
      U95CL = `u-95% CI`,
      SE = Est.Error
    ) %>%
    select(-Rhat, -Bulk_ESS, -Tail_ESS)

  # Add 90% credible intervals
  fixed_table <- fixed_table %>%
    mutate(
      L90CL = fixed_90$`l-90% CI`,
      U90CL = fixed_90$`u-90% CI`
    )

  # Add significance marker
  fixed_table <- fixed_table %>%
    mutate(
      Estimate = ifelse(
        L95CL * U95CL > 0,
        paste(sprintf("%.3f", round(Estimate, 3)), "*", sep = ""),
        ifelse(
          L90CL * U90CL > 0,
          paste(sprintf("%.3f", round(Estimate, 3)), "\u00B0", sep = ""),
          sprintf("%.3f", round(Estimate, 3))
        )
      ),
      rowname = rownames(summary(model)$fixed)
    ) %>%
    column_to_rownames(var = "rowname")

  if (reorder) {
    fixed_table <- fixed_table[order, ]
  }

  fixed_table %>%
    kable(digits = 3, escape = FALSE) %>%
    kable_styling(position = "left")
}

```

# Plot theme
```{r theme}
my_theme <- theme(
  panel.background = element_blank(),
  panel.border = element_rect(fill = NA),
  strip.background = element_blank(),
  axis.text = element_text(size = 10, color = "black"),
  axis.title = element_text(size = 10),
  legend.position = "none",
  text = element_text(family = "Helvetica", size = 10),
  plot.title = element_text(hjust = 0.5),
  plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm")
)
```

# Study Location Map
```{r map_data}
country_data <- read_csv(
  "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Data/location_study.csv"
)
# ---- 1: Load and clean country names ----
countries_long <- country_data %>%
  mutate(Country = str_split(Country, ",")) %>%
  unnest(Country) %>%
  mutate(Country = str_trim(Country)) %>%
  filter(Country != "") %>%
  mutate(
    Country = case_when(
      Country == "USA" ~ "United States of America",
      Country %in% c("England", "Scotland", "Great Britain") ~ "United Kingdom",
      Country == "Germain" ~ "Germany",
      Country == "Belgique" ~ "Belgium",
      Country == "Swiss" ~ "Switzerland",
      Country == "Portugal and France" ~ "France",
      TRUE ~ Country
    )
  ) %>%
  distinct(Country)

# ---- 2: World polygons ----
world_sf <- ne_countries(scale = "medium", returnclass = "sf")

# Select only your countries
world_selected <- world_sf %>%
  filter(admin %in% countries_long$Country)

# ---- 3: Compute centroids for plotting ----
centroids_sf <- world_selected %>%
  st_centroid() %>%
  st_transform(crs = "+proj=robin +lon_0=0 +datum=WGS84 +units=m +no_defs")

world_selected <- world_selected %>%
  st_transform(crs = "+proj=robin +lon_0=0 +datum=WGS84 +units=m +no_defs")

world_sf <- world_sf %>%
  st_transform(crs = "+proj=robin +lon_0=0 +datum=WGS84 +units=m +no_defs")

# ---- 4: Plot ----
# Robinson projection string
robinson <- "+proj=robin +lon_0=0 +datum=WGS84 +units=m +no_defs"

# Create graticule (every 30° lat, 60° lon)
graticule <- st_graticule(
  lat = seq(-60, 90, by = 30),
  lon = seq(-180, 180, by = 60)
) %>%
  st_transform(crs = robinson)

# Transform world and centroids
world_sf_robin <- st_transform(world_sf, crs = robinson)
world_selected_robin <- st_transform(world_selected, crs = robinson)
centroids_sf_robin <- st_transform(centroids_sf, crs = robinson)

# ---- Plot Fig0 with graticule ----
Fig0 <- ggplot() +
  # Base map
  geom_sf(data = world_sf_robin, fill = "grey90", color = "white", size = 0.1) +
  # Graticule grid
  geom_sf(data = graticule, color = "grey70", size = 0.12, linetype = "solid") +
  # Highlight selected countries
  geom_sf(data = world_selected_robin, fill = "grey60", color = "white", size = 0.2) +
  # Centroid markers
  geom_sf(data = centroids_sf_robin, color = "red", size = 2) +
  coord_sf(crs = robinson) +
  labs(x = NULL, y = NULL) +
  theme_void() +
  theme(
    panel.background = element_rect(fill = "white", color = NA)
  )

# Display the figure
Fig0
```

```{r}
ggsave(
  filename  = "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/map_study_countries.pdf",
  plot = Fig0,
  device = cairo_pdf,     # best for vector graphics
  width = 8,              # inches
  height = 6,             # inches
  units = "in"
)
```


# Abundance and CPH
## Load dataset
```{r data_load}
datlambda <- read_csv(
  "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Data/datlambda.csv"
)
datlambda_unique <- datlambda %>%
  distinct(ID, .keep_all = TRUE)
```

## GD vs CD Subsets
```{r subsets}
datlambda_gd <- datlambda %>% filter(Method == "GD")
datlambda_cd <- datlambda %>% filter(Method %in% c("ED", "MD"))
```


##  Bayesian Meta-analysis (GD)
### Model
```{r model_gd}
priors_lb <- c(
  prior(normal(0, 1), class = "Intercept"),
  prior(cauchy(0, 0.5), class = "sd")
)

Lambdamodel_gd <- brm(
  yi | se(se, sigma = TRUE) ~ 1 + (1 | ID) + (1 | Species),
  data = datlambda_gd,
  family = gaussian(),
  prior = priors_lb,
  seed = 13,
  iter = 3000,
  cores = 4,
  control = list(adapt_delta = 0.99, max_treedepth = 15)
)
```

### Posterior predictive check
```{r ppc_gd}
ppc_growthrate_gd <- pp_check(Lambdamodel_gd, ndraws = 1000) +
  xlab("Population growth rate") +
  ylab("Density") +
  my_theme

ppc_growthrate_gd
```

### Grand mean effect table
```{r table_gd}
Lambdamodel_gd %>% Ftable(percent = TRUE)
```

### Posterior probability of positive effect
```{r ecdf_gd}
post_samples_lambda_gd <- as_draws_df(Lambdamodel_gd, variable = "b_Intercept")
mean(post_samples_lambda_gd$b_Intercept > 0)

smd_ecdf_lambda_gd <- ecdf(post_samples_lambda_gd$b_Intercept)
smd_ecdf_lambda_gd(0)

```

### Forest plot (GD)
```{r forest_gd}
# Extract study and species random effects
study.draws_lb_gd <- Lambdamodel_gd %>%
  spread_draws(r_ID[ID, ], r_Species[Species, ], b_Intercept) %>%
  mutate(b_Intercept = r_ID + r_Species + b_Intercept)

# Pooled effect draws
pooled.effect.draws_lb_gd <- Lambdamodel_gd %>%
  spread_draws(b_Intercept) %>%
  mutate(ID = "Pooled Effect")

# Combine and prepare for plotting
# Prepare study-level data
study_data_lb_gd <- study.draws_lb_gd %>%
  mutate(ID = as.character(ID)) %>%
  mutate(ID = str_replace_all(ID, "[.]", " ")) %>%
  # reorder studies by effect size
  mutate(ID = reorder(ID, b_Intercept))

# Bind pooled effect at bottom
forest.data_lb_gd <- bind_rows(study_data_lb_gd, pooled.effect.draws_lb_gd) %>%
  # ensure ID is character so releveling works
  mutate(ID = as.character(ID)) %>%
  # ensure Pooled Effect is last
  mutate(ID = fct_relevel(ID, "Pooled Effect", after = Inf)) %>%
  # add a flag for coloring in the plot
  mutate(effect_type = ifelse(ID == "Pooled Effect", "Pooled", "Study")) %>%
  ungroup()

# forest.data_lb_gd <- forest.data_lb_gd %>%
#   arrange(b_Intercept) %>%
#   mutate(ID = factor(ID, levels = ID))

# Summaries (mean and 95% interval)
forest.data.summary_lb_gd <- forest.data_lb_gd %>%
  group_by(ID) %>%
  mean_qi(b_Intercept) %>%
  # ensure Pooled Effect is last in summary as well
  mutate(ID = fct_relevel(ID, "Pooled Effect", after = Inf)) %>%
  # add the same effect_type column for the summary
  mutate(effect_type = ifelse(ID == "Pooled Effect", "Pooled", "Study"))
```

```{r}
# Plot
forest.plot_lb_gd<-ggplot(forest.data_lb_gd, aes(x = b_Intercept, y = ID, fill = effect_type)) +
  geom_vline(xintercept = fixef(Lambdamodel_gd)[1, 3:4], color = "grey", linetype = 2) +
  geom_vline(xintercept = 0, color = "black", size = 0.75) +
  geom_density_ridges(rel_min_height = 0.01, col = NA, scale = 1, alpha = 0.8) +
  geom_pointinterval(
    data = forest.data.summary_lb_gd,
    aes(x = b_Intercept, y = ID, xmin = .lower, xmax = .upper, color = effect_type),
    size = 1,
    orientation = "y"
  ) +
  geom_text(
    data = mutate_if(forest.data.summary_lb_gd, is.numeric, round, 2),
    aes(label = glue("{b_Intercept} [{.lower}, {.upper}]"), y = ID, x = 2.8),
    hjust = "inward",
    size = 3
  ) +
  scale_y_discrete(limits = rev(levels(forest.data_lb_gd$ID)), labels = NULL, breaks = NULL, expand = expansion(add = 0.75)) +
  scale_fill_manual(values = c("Study" = "#E69F00", "Pooled" = "#56B4E9")) +
  scale_color_manual(values = c("Study" = "#E69F00", "Pooled" = "#56B4E9")) +
  labs(x = "Effect size (lambda)", y = NULL) +
  my_theme
forest.plot_lb_gd
```


```{r}
# Extract study and species random effects
forest.data_lb_gd <- Lambdamodel_gd %>%
  spread_draws(r_ID[ID, ], r_Species[Species, ], b_Intercept) %>%
  mutate(b_Intercept = r_ID + r_Species + b_Intercept) %>%
  mutate(ID = as.character(ID)) %>%
  mutate(ID = str_replace_all(ID, "[.]", " ")) %>%
  bind_rows(
    Lambdamodel_gd %>%
      spread_draws(b_Intercept) %>%
      mutate(ID = "Pooled Effect")
  ) %>%
  mutate(effect_type = ifelse(ID == "Pooled Effect", "Pooled", "Study")) %>%
  ungroup()

# Summaries
forest.data.summary_lb_gd <- forest.data_lb_gd %>%
  group_by(ID) %>%
  mean_qi(b_Intercept) %>%
  mutate(effect_type = ifelse(ID == "Pooled Effect", "Pooled", "Study"))

# Create Study 1..n labels ranked by ascending effect size
study_ranking <- forest.data.summary_lb_gd %>%
  filter(ID != "Pooled Effect") %>%
  arrange(b_Intercept) %>%
  mutate(new_label = paste0("Study ", row_number()))  # Study 1 = lowest effect

# Create lookup vector: old ID -> new label
label_lookup <- setNames(study_ranking$new_label, study_ranking$ID)

# Apply new labels to both data frames
forest.data_lb_gd <- forest.data_lb_gd %>%
  mutate(ID = ifelse(ID == "Pooled Effect", "Pooled Effect", label_lookup[ID]))

forest.data.summary_lb_gd <- forest.data.summary_lb_gd %>%
  mutate(ID = ifelse(ID == "Pooled Effect", "Pooled Effect", label_lookup[ID]))

# Build id_order: Study 1 to Study n, then Pooled Effect
id_order <- c(paste0("Study ", seq_len(nrow(study_ranking))), "Pooled Effect")

# Apply factor levels
forest.data_lb_gd         <- forest.data_lb_gd         %>% mutate(ID = factor(ID, levels = id_order))
forest.data.summary_lb_gd <- forest.data.summary_lb_gd %>% mutate(ID = factor(ID, levels = id_order))

# Plot
forest.plot_lb_gd <- ggplot(
    forest.data_lb_gd,
    aes(x = b_Intercept, y = ID, fill = effect_type)
  ) +
  geom_vline(xintercept = fixef(Lambdamodel_gd)[1, 3:4], color = "grey", linetype = 2) +
  geom_vline(xintercept = 0, color = "black", size = 0.75) +
  geom_density_ridges(rel_min_height = 0.01, col = NA, scale = 1, alpha = 0.8) +
  geom_pointinterval(
    data = forest.data.summary_lb_gd,
    aes(x = b_Intercept, xmin = .lower, xmax = .upper, color = effect_type),
    size = 1, orientation = "y"
  ) +
  geom_text(
    data = mutate_if(forest.data.summary_lb_gd, is.numeric, round, 2),
    aes(label = glue("{b_Intercept} [{.lower}, {.upper}]"), x = 2.8),
    hjust = "inward", size = 3
  ) +
  scale_y_discrete(
    limits = rev(id_order),
    expand = expansion(add = 0.75)
  ) +
  scale_fill_manual(values  = c("Study" = "#E69F00", "Pooled" = "#56B4E9")) +
  scale_color_manual(values = c("Study" = "#E69F00", "Pooled" = "#56B4E9")) +
  labs(x = "Effect size (lambda)", y = NULL) +
  my_theme

forest.plot_lb_gd
```

```{r}
ggsave(
  filename  = "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/lambda_gd_forest.pdf",
  plot = forest.plot_lb_gd,
  device = cairo_pdf,     # best for vector graphics
  width = 6,              # inches
  height = 6,             # inches
  units = "in"
)
```

```{r}
I2_gd <- as_draws_df(Lambdamodel_gd) %>%
  transmute(
    tau2_ID = (sd_ID__Intercept)^2,
    tau2_Species = (sd_Species__Intercept)^2,
    sigma2 = (sigma)^2
  ) %>%
  mutate(
    vbar = mean(datlambda_gd$se^2),
    var_total = tau2_ID + tau2_Species + sigma2,
    I2_total = var_total / (var_total + vbar)
  ) %>%
  summarise(
    I2_mean = mean(I2_total),
    I2_median = median(I2_total),
    I2_CI_lower = quantile(I2_total, 0.025),
    I2_CI_upper = quantile(I2_total, 0.975)
  )

I2_gd
```
## Bayesian Meta-analysis (CD)
### Model
```{r}
Lambdamodel_cd <- brm(
  yi | se(se, sigma = TRUE) ~ 1 + (1 | ID) + (1 | Species),
  data = datlambda_cd,
  family = gaussian(),
  prior = priors_lb,
  seed = 13,
  iter = 3000,
  cores = 4,
  control = list(adapt_delta = 0.99, max_treedepth = 15)
)
```

### Posterior predictive check
```{r}
ppc_growthrate_cd <- pp_check(Lambdamodel_cd, ndraws = 1000) +
  xlab("Population growth rate") +
  ylab("Density") +
  my_theme

ppc_growthrate_cd

```

### Grand mean effect table
```{r}
Lambdamodel_cd %>% Ftable(percent = TRUE)
```

### Posterior probability of positive effect
```{r}
post_samples_lambda_cd <- as_draws_df(Lambdamodel_cd, variable = "b_Intercept")
mean(post_samples_lambda_cd$b_Intercept > 0)

smd_ecdf_lambda_cd <- ecdf(post_samples_lambda_cd$b_Intercept)
smd_ecdf_lambda_cd(0)
```

### Forest plot (CD)
```{r}
# Extract study and species random effects
study.draws_lb_cd <- Lambdamodel_cd %>%
  spread_draws(r_ID[ID, ], r_Species[Species, ], b_Intercept) %>%
  mutate(b_Intercept = r_ID + r_Species + b_Intercept)

# Pooled effect draws
pooled.effect.draws_lb_cd <- Lambdamodel_cd %>%
  spread_draws(b_Intercept) %>%
  mutate(ID = "Pooled Effect")

# Combine and prepare for plotting
study_data_lb_cd <- study.draws_lb_cd %>%
  mutate(ID = as.character(ID)) %>%
  mutate(ID = str_replace_all(ID, "[.]", " ")) %>%
  mutate(ID = reorder(ID, b_Intercept))

forest.data_lb_cd <- bind_rows(study_data_lb_cd, pooled.effect.draws_lb_cd) %>%
  mutate(ID = as.character(ID)) %>%
  mutate(ID = fct_relevel(ID, "Pooled Effect", after = Inf)) %>%
  mutate(effect_type = ifelse(ID == "Pooled Effect", "Pooled", "Study")) %>%
  ungroup()

# Summaries (mean and 95% interval)
forest.data.summary_lb_cd <- forest.data_lb_cd %>%
  group_by(ID) %>%
  mean_qi(b_Intercept) %>%
  mutate(ID = fct_relevel(ID, "Pooled Effect", after = Inf)) %>%
  mutate(effect_type = ifelse(ID == "Pooled Effect", "Pooled", "Study"))

```

```{r}
forest.plot_lb_cd <- ggplot(forest.data_lb_cd, aes(x = b_Intercept, y = ID, fill = effect_type)) +
  geom_vline(xintercept = fixef(Lambdamodel_cd)[1, 3:4], color = "grey", linetype = 2) +
  geom_vline(xintercept = 0, color = "black", size = 0.75) +
  geom_density_ridges(rel_min_height = 0.01, col = NA, scale = 1, alpha = 0.8) +
  geom_pointinterval(
    data = forest.data.summary_lb_cd,
    aes(x = b_Intercept, y = ID, xmin = .lower, xmax = .upper, color = effect_type),
    size = 1,
    orientation = "y"
  ) +
  geom_text(
    data = mutate_if(forest.data.summary_lb_cd, is.numeric, round, 2),
    aes(label = glue("{b_Intercept} [{.lower}, {.upper}]"), y = ID, x = 4.5),
    hjust = "inward",
    size = 3
  ) +
  scale_y_discrete(limits = rev(levels(forest.data_lb_cd$ID)), labels = NULL, breaks = NULL, expand = expansion(add = 0.75)) +
  scale_fill_manual(values = c("Study" = "#E69F00", "Pooled" = "#56B4E9")) +
  scale_color_manual(values = c("Study" = "#E69F00", "Pooled" = "#56B4E9")) +
  labs(x = "Effect size (lambda)", y = NULL) +
  my_theme

forest.plot_lb_cd
```

```{r}
ggsave(
  filename  = "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/lambda_cd_forest.pdf",
  plot = forest.plot_lb_cd,
  device = cairo_pdf,     # best for vector graphics
  width = 6,              # inches
  height = 6,             # inches
  units = "in"
)
```

### I² Calculation (CD)
```{r}
I2_cd <- as_draws_df(Lambdamodel_cd) %>%
  transmute(
    tau2_ID = (sd_ID__Intercept)^2,
    tau2_Species = (sd_Species__Intercept)^2,
    sigma2 = (sigma)^2
  ) %>%
  mutate(
    vbar = mean(datlambda_cd$se^2),
    var_total = tau2_ID + tau2_Species + sigma2,
    I2_total = var_total / (var_total + vbar)
  ) %>%
  summarise(
    I2_mean = mean(I2_total),
    I2_median = median(I2_total),
    I2_CI_lower = quantile(I2_total, 0.025),
    I2_CI_upper = quantile(I2_total, 0.975)
  )

I2_cd
```

# Genetics and CPH
```{r}
datgencp <- read.csv(
  "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Data/datgencp.csv",
  header = TRUE
)
```

## Basic data summaries
```{r}
total_studies_g <- n_distinct(datgencp$ID)
total_species    <- n_distinct(datgencp$Species)

species_by_kingdom <- datgencp %>%
  group_by(Kingdom) %>%
  summarize(n_species = n_distinct(Species))

species_counts <- datgencp %>%
  group_by(Species, Kingdom) %>%
  summarize(n_records = n()) %>%
  arrange(desc(n_records))

total_species
species_by_kingdom

```

## Subset data by Method
```{r}
datgencp_gd <- datgencp %>% filter(Method == "GD")
datgencp_cd <- datgencp %>% filter(Method == "ED")
```

## Model: Genetic diversity (GD)
### Fit the model
```{r}
priors_gn <- c(
  prior(normal(0, 1), class = Intercept),
  prior(cauchy(0, 0.5), class = sd)
)

genamodel_gd <- brm(
  yi | se(se, sigma = TRUE) ~ 1 + (1 | ID) + (1 | Species),
  data = datgencp_gd,
  prior = priors_gn,
  iter = 8000,
  cores = 4,
  seed = 13,
  control = list(adapt_delta = 0.9999, max_treedepth = 15)
)

```

### Posterior predictive check
```{r}
pp_check(genamodel_gd, ndraws = 1000) +
  xlab("Genetic diversity") +
  ylab("Density") +
  ggtitle("") +
  my_theme -> ppc_genetic_gd
```

### Grand mean effect
```{r}
genamodel_gd %>% Ftable(percent = TRUE)
```

### Probability that effect > 0
```{r}
post.samples_gen_gd <- posterior_samples(genamodel_gd, pars = "^b_Intercept")
mean(post.samples_gen_gd$b_Intercept > 0)
```

### Heterogeneity (I²) for GD
```{r}
I2_gd <- as_draws_df(genamodel_gd) %>%
  transmute(
    tau2_ID      = (sd_ID__Intercept)^2,
    tau2_Species = (sd_Species__Intercept)^2,
    sigma2       = (sigma)^2
  ) %>%
  mutate(
    vbar      = mean(datgencp_gd$se^2),
    var_total = tau2_ID + tau2_Species + sigma2,
    I2_total  = var_total / (var_total + vbar)
  ) %>%
  summarise(
    I2_mean     = mean(I2_total),
    I2_median   = median(I2_total),
    I2_CI_lower = quantile(I2_total, 0.025),
    I2_CI_upper = quantile(I2_total, 0.975)
  )

I2_gd

```

### Forest plot for GD
```{r}
study.draws_gd <- genamodel_gd %>%
  spread_draws(r_ID[ID, ], r_Species[Species, ], b_Intercept) %>%
  mutate(b_Intercept = r_ID + r_Species + b_Intercept)

pooled.effect.draws_gd <- genamodel_gd %>%
  spread_draws(b_Intercept) %>%
  mutate(ID = "Pooled Effect")

study_data_gd <- study.draws_gd %>%
  mutate(ID = as.character(ID)) %>%
  mutate(ID = str_replace_all(ID, "[.]", " ")) %>%
  mutate(ID = reorder(ID, b_Intercept))

forest.data_gd <- bind_rows(study_data_gd, pooled.effect.draws_gd) %>%
  mutate(ID = as.character(ID)) %>%
  mutate(ID = fct_relevel(ID, "Pooled Effect", after = Inf)) %>%
  mutate(effect_type = ifelse(ID == "Pooled Effect", "Pooled", "Study")) %>%
  ungroup()

forest.data.summary_gd <- forest.data_gd %>%
  group_by(ID) %>%
  mean_qi(b_Intercept) %>%
  mutate(ID = fct_relevel(ID, "Pooled Effect", after = Inf)) %>%
  mutate(effect_type = ifelse(ID == "Pooled Effect", "Pooled", "Study"))

```

```{r}
forest.plot_gd<-ggplot(forest.data_gd, aes(x = b_Intercept, y = ID, fill = effect_type)) +
  geom_vline(xintercept = fixef(genamodel_gd)[1, 3:4], color = "grey", linetype = 2) +
  geom_vline(xintercept = 0, color = "black", size = 0.75) +
  geom_density_ridges(rel_min_height = 0.01, col = NA, scale = 1, alpha = 0.8) +
  geom_pointinterval(
    data = forest.data.summary_gd,
    aes(x = b_Intercept, y = ID, xmin = .lower, xmax = .upper, color = effect_type),
    size = 1, orientation = "y"
  ) +
  xlim(-1.5, 5) +
  geom_text(
    data = mutate_if(forest.data.summary_gd, is.numeric, round, 2),
    aes(label = glue("{b_Intercept} [{.lower}, {.upper}]"), y = ID, x = 4.5),
    hjust = "inward", size = 3
  ) +
  scale_y_discrete(limits = rev(levels(forest.data_gd$ID)), labels = NULL, breaks = NULL, expand = expansion(add = 0.75)) +
  scale_fill_manual(values = c("Study" = "#E69F00", "Pooled" = "#56B4E9")) +
  scale_color_manual(values = c("Study" = "#E69F00", "Pooled" = "#56B4E9")) +
  labs(x = "Effect size (genetic diversity)", y = NULL) +
  my_theme

forest.plot_gd
```


```{r}
# Extract study and species random effects
forest.data_gd <- genamodel_gd %>%
  spread_draws(r_ID[ID, ], r_Species[Species, ], b_Intercept) %>%
  mutate(b_Intercept = r_ID + r_Species + b_Intercept) %>%
  mutate(ID = as.character(ID)) %>%
  mutate(ID = str_replace_all(ID, "[.]", " ")) %>%
  bind_rows(
    genamodel_gd %>%
      spread_draws(b_Intercept) %>%
      mutate(ID = "Pooled Effect")
  ) %>%
  mutate(effect_type = ifelse(ID == "Pooled Effect", "Pooled", "Study")) %>%
  ungroup()

# Summaries
forest.data.summary_gd <- forest.data_gd %>%
  group_by(ID) %>%
  mean_qi(b_Intercept) %>%
  mutate(effect_type = ifelse(ID == "Pooled Effect", "Pooled", "Study"))

# Create Study 1..n labels ranked by ascending effect size
study_ranking_gd <- forest.data.summary_gd %>%
  filter(ID != "Pooled Effect") %>%
  arrange(b_Intercept) %>%
  mutate(new_label = paste0("Study ", row_number()))

# Create lookup vector: old ID -> new label
label_lookup_gd <- setNames(study_ranking_gd$new_label, study_ranking_gd$ID)

# Apply new labels to both data frames
forest.data_gd <- forest.data_gd %>%
  mutate(ID = ifelse(ID == "Pooled Effect", "Pooled Effect", label_lookup_gd[ID]))

forest.data.summary_gd <- forest.data.summary_gd %>%
  mutate(ID = ifelse(ID == "Pooled Effect", "Pooled Effect", label_lookup_gd[ID]))

# Build id_order: Study 1 to Study n, then Pooled Effect
id_order_gd <- c(paste0("Study ", seq_len(nrow(study_ranking_gd))), "Pooled Effect")

# Apply factor levels
forest.data_gd         <- forest.data_gd         %>% mutate(ID = factor(ID, levels = id_order_gd))
forest.data.summary_gd <- forest.data.summary_gd %>% mutate(ID = factor(ID, levels = id_order_gd))

# Plot
forest.plot_gd <- ggplot(
    forest.data_gd,
    aes(x = b_Intercept, y = ID, fill = effect_type)
  ) +
  geom_vline(xintercept = fixef(genamodel_gd)[1, 3:4], color = "grey", linetype = 2) +
  geom_vline(xintercept = 0, color = "black", size = 0.75) +
  geom_density_ridges(rel_min_height = 0.01, col = NA, scale = 1, alpha = 0.8) +
  geom_pointinterval(
    data = forest.data.summary_gd,
    aes(x = b_Intercept, xmin = .lower, xmax = .upper, color = effect_type),
    size = 1, orientation = "y"
  ) +
  xlim(-1.5, 5) +
  geom_text(
    data = mutate_if(forest.data.summary_gd, is.numeric, round, 2),
    aes(label = glue("{b_Intercept} [{.lower}, {.upper}]"), x = 4.5),
    hjust = "inward", size = 3
  ) +
  scale_y_discrete(
    limits = rev(id_order_gd),
    expand = expansion(add = 0.75)
  ) +
  scale_fill_manual(values  = c("Study" = "#E69F00", "Pooled" = "#56B4E9")) +
  scale_color_manual(values = c("Study" = "#E69F00", "Pooled" = "#56B4E9")) +
  labs(x = "Effect size (genetic diversity)", y = NULL) +
  my_theme

forest.plot_gd
```

```{r}
ggsave(
  filename  = "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/gen_forest_gd.pdf",
  plot = forest.plot_gd,
  device = cairo_pdf,     # best for vector graphics
  width = 6,              # inches
  height = 10,             # inches
  units = "in"
)
```
## Model: Genetic diversity (CD)
```{r}
genamodel <- brm(
  yi | se(se, sigma = TRUE) ~ 1 + (1 | ID) + (1 | Species),
  data = datgencp,
  prior = priors_gn,
  iter = 8000,
  cores = 4,
  seed = 13,
  control = list(adapt_delta = 0.9999, max_treedepth = 15)
)

```

## Posterior predictive check
```{r}
pp_check(genamodel, ndraws = 1000) +
  xlab("Genetic diversity") +
  ylab("Density") +
  ggtitle("") +
  my_theme -> ppc_genetic
```

## Grand mean effect
```{r}
genamodel %>% Ftable(percent = TRUE)
```

## Probability that effect > 0
```{r}
post.samples_gen<- posterior_samples(genamodel, pars = "^b_Intercept")
mean(post.samples_gen$b_Intercept > 0)
```

## Heterogeneity for CD
```{r}
I2 <- as_draws_df(genamodel) %>%
  transmute(
    tau2_ID      = (sd_ID__Intercept)^2,
    tau2_Species = (sd_Species__Intercept)^2,
    sigma2       = (sigma)^2
  ) %>%
  mutate(
    vbar      = mean(datgencp_cd$se^2),
    var_total = tau2_ID + tau2_Species + sigma2,
    I2_total  = var_total / (var_total + vbar)
  ) %>%
  summarise(
    I2_mean     = mean(I2_total),
    I2_median   = median(I2_total),
    I2_CI_lower = quantile(I2_total, 0.025),
    I2_CI_upper = quantile(I2_total, 0.975)
  )

I2
```
## Forest plot for CD
```{r}
study.draws <- genamodel %>%
  spread_draws(r_ID[ID, ], r_Species[Species, ], b_Intercept) %>%
  mutate(b_Intercept = r_ID + r_Species + b_Intercept)

pooled.effect.draws <- genamodel %>%
  spread_draws(b_Intercept) %>%
  mutate(ID = "Pooled Effect")

study_data <- study.draws %>%
  mutate(ID = as.character(ID)) %>%
  mutate(ID = str_replace_all(ID, "[.]", " ")) %>%
  mutate(ID = reorder(ID, b_Intercept))

forest.data <- bind_rows(study_data, pooled.effect.draws) %>%
  mutate(ID = as.character(ID)) %>%
  mutate(ID = fct_relevel(ID, "Pooled Effect", after = Inf)) %>%
  mutate(effect_type = ifelse(ID == "Pooled Effect", "Pooled", "Study")) %>%
  ungroup()

forest.data.summary <- forest.data %>%
  group_by(ID) %>%
  mean_qi(b_Intercept) %>%
  mutate(ID = fct_relevel(ID, "Pooled Effect", after = Inf)) %>%
  mutate(effect_type = ifelse(ID == "Pooled Effect", "Pooled", "Study"))

```

```{r}
gen.forest.plot <- ggplot(forest.data, aes(x = b_Intercept, y = ID, fill = effect_type)) +
  geom_vline(xintercept = fixef(genamodel)[1, 3:4], color = "grey", linetype = 2) +
  geom_vline(xintercept = 0, color = "black", size = 0.75) +
  geom_density_ridges(rel_min_height = 0.01, col = NA, scale = 1, alpha = 0.8) +
  geom_pointinterval(
    data = forest.data.summary,
    aes(x = b_Intercept, y = ID, xmin = .lower, xmax = .upper, color = effect_type),
    size = 1, orientation = "y"
  ) +
  xlim(-1.5, 5) +
  geom_text(
    data = mutate_if(forest.data.summary, is.numeric, round, 2),
    aes(label = glue("{b_Intercept} [{.lower}, {.upper}]"), y = ID, x = 4.5),
    hjust = "inward", size = 3
  ) +
  scale_y_discrete(
    limits = rev(levels(forest.data$ID)),
    labels = NULL,
    breaks = NULL,
    expand = expansion(add = 0.75)
  ) +
  scale_fill_manual(values = c("Study" = "#E69F00", "Pooled" = "#56B4E9")) +
  scale_color_manual(values = c("Study" = "#E69F00", "Pooled" = "#56B4E9")) +
  labs(x = "Effect size (genetic diversity)", y = NULL) +
  my_theme

gen.forest.plot

```

```{r}
ggsave(
  filename = "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/gen_forest.pdf",
  plot = gen.forest.plot,
  device = cairo_pdf,
  width = 6,
  height = 11,
  units = "in"
)

```

## Population abundance and CPH
```{r}
databuncp <- read.csv("/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Data/databuncp.csv",
                      header = TRUE)
databuncp$Species <- gsub("spp\\.", "sp.", as.character(databuncp$Species))
databuncp <- databuncp %>% filter(vi > 0)
#unique(databuncp$Method)
```

```{r}
total_studies_ab <- n_distinct(databuncp$ID)
total_species     <- n_distinct(databuncp$Species)

species_by_kingdom <- databuncp %>%
  group_by(Kingdom) %>%
  summarize(n_species = n_distinct(Species))

species_counts <- databuncp %>%
  group_by(Species, Kingdom) %>%
  summarize(n_records = n()) %>%
  arrange(desc(n_records))

total_species
species_by_kingdom

```

## Abundance by Geographic Distance (GD)
```{r}
databuncp_gd <- databuncp %>% filter(Method == "GD")
```

### Model
```{r}
priors_ab_gd <- c(
  prior(normal(0, 2), class = "Intercept"),
 prior(student_t(3, 0, 1), class = "sd")
)

abunmodel_gd <- brm(
  yi | se(se, sigma = TRUE) ~ 1 + (1 | ID) + (1 | Species),
  data = databuncp_gd,
  seed = 13,
  iter = 8000,
  warmup = 2000,
  chains = 4,
  control = list(adapt_delta = 0.99, max_treedepth = 15),
  prior = priors_ab_gd
)
```

### Posterior predictive check
```{r}

ppc_abundance_gd <- pp_check(abunmodel_gd, ndraws = 1000) +
  scale_fill_manual(
    values = c("grey70", "steelblue"),
    labels = c("Observed data", "Posterior predictive"),
    name = ""
  ) +
  theme_test() +
  theme(legend.position = c(0.85, 0.85))


ppc_abundance_gd <- pp_check(abunmodel_gd, ndraws = 1000) +
  xlab("Abundance") +
  ylab("Density") +
  ggtitle("") +
  coord_cartesian(xlim = c(-2, 2), ylim = c(0, 3)) +
   scale_colour_manual(
    name = "",
    values = c("Observed" = "black", "Predicted" = "skyblue"),
    labels = c("Observed data", "Posterior predictive")
  ) +
   theme(legend.position = c(0.85,0.85)) +
  theme_test()

```

### Grand mean effect
```{r}
abunmodel_gd %>% Ftable(percent = TRUE)
```

### Heterogeneity (I²)
```{r}
I2_ab <- as_draws_df(abunmodel_gd) %>%
  transmute(
    tau2_ID      = (sd_ID__Intercept)^2,
    tau2_Species = (sd_Species__Intercept)^2,
    sigma2       = (sigma)^2
  ) %>%
  mutate(
    vbar      = mean(databuncp_gd$se^2),
    var_total = tau2_ID + tau2_Species + sigma2,
    I2_total  = var_total / (var_total + vbar)
  ) %>%
  summarise(
    I2_mean     = mean(I2_total),
    I2_median   = median(I2_total),
    I2_CI_lower = quantile(I2_total, 0.025),
    I2_CI_upper = quantile(I2_total, 0.975)
  )

I2_ab
```

### Posterior probability intercept > 0
```{r}
post.samples_ab_gd <- as_draws_df(abunmodel_gd) %>%
  pull(b_Intercept)

mean(post.samples_ab_gd > 0)

```

# Forest plot
```{r}
study.draws_ab_gd <- abunmodel_gd %>%
  spread_draws(r_ID[ID, ], r_Species[Species, ], b_Intercept, ndraws = 1000) %>%
  mutate(b_Intercept = r_ID + r_Species + b_Intercept)

pooled.effect.draws_ab_gd <- abunmodel_gd %>%
  spread_draws(b_Intercept, ndraws = 1000) %>%
  mutate(ID = "Pooled Effect")

study_data_ab_gd <- study.draws_ab_gd %>%
  mutate(ID = as.character(ID)) %>%
  mutate(ID = str_replace_all(ID, "[.]", " ")) %>%
  mutate(ID = reorder(ID, b_Intercept))

forest.data_ab_gd <- bind_rows(study_data_ab_gd, pooled.effect.draws_ab_gd) %>%
  mutate(ID = as.character(ID)) %>%
  mutate(ID = fct_relevel(ID, "Pooled Effect", after = Inf)) %>%
  mutate(effect_type = ifelse(ID == "Pooled Effect", "Pooled", "Study")) %>%
  ungroup()

forest.data.summary_ab_gd <- forest.data_ab_gd %>%
  group_by(ID) %>%
  mean_qi(b_Intercept) %>%
  mutate(ID = fct_relevel(ID, "Pooled Effect", after = Inf)) %>%
  mutate(effect_type = ifelse(ID == "Pooled Effect", "Pooled", "Study"))

```

```{r}
ab.forest.plot_gd<-ggplot(forest.data_ab_gd, aes(x = b_Intercept, y = ID, fill = effect_type)) +
  geom_vline(xintercept = fixef(abunmodel_gd)[1, 3:4], color = "grey", linetype = 2) +
  geom_vline(xintercept = 0, color = "black", size = 0.75) +
  geom_density_ridges(rel_min_height = 0.01, col = NA, scale = 1, alpha = 0.8) +
  geom_pointinterval(
    data = forest.data.summary_ab_gd,
    aes(x = b_Intercept, y = ID, xmin = .lower, xmax = .upper, color = effect_type),
    size = 1, orientation = "y"
  ) +
  xlim(-3, 4) +
  geom_text(
    data = mutate_if(forest.data.summary_ab_gd, is.numeric, round, 2),
    aes(label = glue("{b_Intercept} [{.lower}, {.upper}]"), y = ID, x = 3.75),
    hjust = "inward", size = 3
  ) +
  scale_y_discrete(
    limits = rev(levels(forest.data_ab_gd$ID)),
    labels = NULL,
    breaks = NULL,
    expand = expansion(add = 0.75)
  ) +
  #scale_y_discrete(limits = rev(levels(forest.data_ab_gd$ID)), expand = expansion(add = 0.75)) +
  scale_fill_manual(values = c("Study" = "#E69F00", "Pooled" = "#56B4E9")) +
  scale_color_manual(values = c("Study" = "#E69F00", "Pooled" = "#56B4E9")) +
  labs(x = "Effect size (abundance)", y = NULL) +
  my_theme

ab.forest.plot_gd
```

```{r}
# Extract study and species random effects
forest.data_ab_gd <- abunmodel_gd %>%
  spread_draws(r_ID[ID, ], r_Species[Species, ], b_Intercept, ndraws = 1000) %>%
  mutate(b_Intercept = r_ID + r_Species + b_Intercept) %>%
  mutate(ID = as.character(ID)) %>%
  mutate(ID = str_replace_all(ID, "[.]", " ")) %>%
  bind_rows(
    abunmodel_gd %>%
      spread_draws(b_Intercept, ndraws = 1000) %>%
      mutate(ID = "Pooled Effect")
  ) %>%
  mutate(effect_type = ifelse(ID == "Pooled Effect", "Pooled", "Study")) %>%
  ungroup()

# Summaries
forest.data.summary_ab_gd <- forest.data_ab_gd %>%
  group_by(ID) %>%
  mean_qi(b_Intercept) %>%
  mutate(effect_type = ifelse(ID == "Pooled Effect", "Pooled", "Study"))

# Create Study 1..n labels ranked by ascending effect size
study_ranking_ab_gd <- forest.data.summary_ab_gd %>%
  filter(ID != "Pooled Effect") %>%
  arrange(b_Intercept) %>%
  mutate(new_label = paste0("Study ", row_number()))

# Create lookup vector: old ID -> new label
label_lookup_ab_gd <- setNames(study_ranking_ab_gd$new_label, study_ranking_ab_gd$ID)

# Apply new labels to both data frames
forest.data_ab_gd <- forest.data_ab_gd %>%
  mutate(ID = ifelse(ID == "Pooled Effect", "Pooled Effect", label_lookup_ab_gd[ID]))

forest.data.summary_ab_gd <- forest.data.summary_ab_gd %>%
  mutate(ID = ifelse(ID == "Pooled Effect", "Pooled Effect", label_lookup_ab_gd[ID]))

# Build id_order: Study 1 to Study n, then Pooled Effect
id_order_ab_gd <- c(paste0("Study ", seq_len(nrow(study_ranking_ab_gd))), "Pooled Effect")

# Apply factor levels
forest.data_ab_gd         <- forest.data_ab_gd         %>% mutate(ID = factor(ID, levels = id_order_ab_gd))
forest.data.summary_ab_gd <- forest.data.summary_ab_gd %>% mutate(ID = factor(ID, levels = id_order_ab_gd))

# Plot
ab.forest.plot_gd <- ggplot(
    forest.data_ab_gd,
    aes(x = b_Intercept, y = ID, fill = effect_type)
  ) +
  geom_vline(xintercept = fixef(abunmodel_gd)[1, 3:4], color = "grey", linetype = 2) +
  geom_vline(xintercept = 0, color = "black", size = 0.75) +
  geom_density_ridges(rel_min_height = 0.01, col = NA, scale = 1, alpha = 0.8) +
  geom_pointinterval(
    data = forest.data.summary_ab_gd,
    aes(x = b_Intercept, xmin = .lower, xmax = .upper, color = effect_type),
    size = 1, orientation = "y"
  ) +
  xlim(-3, 4) +
  geom_text(
    data = mutate_if(forest.data.summary_ab_gd, is.numeric, round, 2),
    aes(label = glue("{b_Intercept} [{.lower}, {.upper}]"), x = 3.75),
    hjust = "inward", size = 3
  ) +
  scale_y_discrete(
    limits = rev(id_order_ab_gd),
    expand = expansion(add = 0.75)
  ) +
  scale_fill_manual(values  = c("Study" = "#E69F00", "Pooled" = "#56B4E9")) +
  scale_color_manual(values = c("Study" = "#E69F00", "Pooled" = "#56B4E9")) +
  labs(x = "Effect size (abundance)", y = NULL) +
  my_theme

ab.forest.plot_gd
```


```{r}
ggsave(
  filename = "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/abun_cp_forest_gd.pdf",
  plot = ab.forest.plot_gd,
  device = cairo_pdf,
  width = 5,
  height = 8,
  units = "in"
)

```

## Abundance by Environmental Distance (ED)
```{r}
databuncp_ed <- databuncp %>% filter(Method == "ED")
```

### Model
```{r}
priors_ab_ed <- c(
  prior(normal(0, 2), class = "Intercept"),
  prior(student_t(3, 0, 1), class = "sd")
)

abunmodel_ed <- brm(
  yi | se(se, sigma = TRUE) ~ 1 + (1 | ID) + (1 | Species),
  data = databuncp_ed,
  seed = 13,
  iter = 8000,
  warmup = 2000,
  chains = 4,
  control = list(adapt_delta = 0.99, max_treedepth = 15),
  prior = priors_ab_ed
)

```

### Posterior predictive check
```{r}
ppc_abundance_ed <- pp_check(abunmodel_ed, ndraws = 1000) +
  xlab("Abundance") +
  ylab("Density") +
  ggtitle("") +
  coord_cartesian(xlim = c(-2, 2), ylim = c(0, 3)) +
  my_theme
```

### Grand mean effect
```{r}
abunmodel_ed %>% Ftable(percent = TRUE)
```
### Heterogeneity (I²)
```{r}
I2_ab_ed <- as_draws_df(abunmodel_ed) %>%
  transmute(
    tau2_ID      = (sd_ID__Intercept)^2,
    tau2_Species = (sd_Species__Intercept)^2,
    sigma2       = (sigma)^2
  ) %>%
  mutate(
    vbar      = mean(databuncp_ed$se^2),
    var_total = tau2_ID + tau2_Species + sigma2,
    I2_total  = var_total / (var_total + vbar)
  ) %>%
  summarise(
    I2_mean     = mean(I2_total),
    I2_median   = median(I2_total),
    I2_CI_lower = quantile(I2_total, 0.025),
    I2_CI_upper = quantile(I2_total, 0.975)
  )

I2_ab_ed

```

### Posterior probability intercept > 0
```{r}
post.samples_ab_ed <- as_draws_df(abunmodel_ed) %>%
  pull(b_Intercept)

mean(post.samples_ab_ed > 0)

```

### Forest plot
```{r}
study.draws_ab_ed <- abunmodel_ed %>%
  spread_draws(r_ID[ID, ], r_Species[Species, ], b_Intercept, ndraws = 1000) %>%
  mutate(b_Intercept = r_ID + r_Species + b_Intercept)

pooled.effect.draws_ab_ed <- abunmodel_ed %>%
  spread_draws(b_Intercept, ndraws = 1000) %>%
  mutate(ID = "Pooled Effect")

study_data_ab_ed <- study.draws_ab_ed %>%
  mutate(ID = as.character(ID)) %>%
  mutate(ID = str_replace_all(ID, "[.]", " ")) %>%
  mutate(ID = reorder(ID, b_Intercept))

forest.data_ab_ed <- bind_rows(study_data_ab_ed, pooled.effect.draws_ab_ed) %>%
  mutate(ID = as.character(ID)) %>%
  mutate(ID = fct_relevel(ID, "Pooled Effect", after = Inf)) %>%
  mutate(effect_type = ifelse(ID == "Pooled Effect", "Pooled", "Study")) %>%
  ungroup()

forest.data.summary_ab_ed <- forest.data_ab_ed %>%
  group_by(ID) %>%
  mean_qi(b_Intercept) %>%
  mutate(ID = fct_relevel(ID, "Pooled Effect", after = Inf)) %>%
  mutate(effect_type = ifelse(ID == "Pooled Effect", "Pooled", "Study"))

```

```{r}
ab.forest.plot_ed <- ggplot(forest.data_ab_ed, aes(x = b_Intercept, y = ID, fill = effect_type)) +
  geom_vline(xintercept = fixef(abunmodel_ed)[1, 3:4], color = "grey", linetype = 2) +
  geom_vline(xintercept = 0, color = "black", size = 0.75) +
  geom_density_ridges(rel_min_height = 0.01, col = NA, scale = 1, alpha = 0.8) +
  geom_pointinterval(
    data = forest.data.summary_ab_ed,
    aes(x = b_Intercept, y = ID, xmin = .lower, xmax = .upper, color = effect_type),
    size = 1, orientation = "y"
  ) +
  xlim(-2, 2) +
  geom_text(
    data = mutate_if(forest.data.summary_ab_ed, is.numeric, round, 2),
    aes(label = glue("{b_Intercept} [{.lower}, {.upper}]"), y = ID, x = 1.75),
    hjust = "inward", size = 3
  ) +
  scale_y_discrete(
    limits = rev(levels(forest.data_ab_ed$ID)),
    labels = NULL,
    breaks = NULL,
    expand = expansion(add = 0.75)
  ) +
  scale_fill_manual(values = c("Study" = "#E69F00", "Pooled" = "#56B4E9")) +
  scale_color_manual(values = c("Study" = "#E69F00", "Pooled" = "#56B4E9")) +
  labs(x = "Effect size (abundance)", y = NULL) +
  my_theme

ab.forest.plot_ed

```

```{r}
ggsave(
  filename = "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/abun_cp_forest_ed.pdf",
  plot = ab.forest.plot_ed,
  device = cairo_pdf,
  width = 5,
  height = 8,
  units = "in"
)

```

## Abundance by Mean Distance (MD)
```{r}
databuncp_md <- databuncp %>% filter(Method == "MD")

```

## Model
```{r}
priors_ab_md <- c(
  prior(normal(0, 1), class = "Intercept"),
  prior(exponential(5), class = "sd")
)

abunmodel_md <- brm(
  yi | se(se, sigma = TRUE) ~ 1 + (1 | Species) + (1 | ID),
  data = databuncp_md,
  seed = 13,
  iter = 10000,
  warmup = 4000,
  chains = 4,
  control = list(adapt_delta = 0.995, max_treedepth = 15),
  prior = priors_ab_md
)

```

## Posterior predictive check
```{r}
ppc_abundance_md <- pp_check(abunmodel_md, ndraws = 1000) +
  xlab("Abundance") +
  ylab("Density") +
  ggtitle("") +
  coord_cartesian(xlim = c(-2, 2), ylim = c(0, 3)) +
  my_theme

```

## Grand mean effect
```{r}
abunmodel_md %>% Ftable(percent = TRUE)
```
## Heterogeneity (I²)
```{r}
I2_ab_md <- as_draws_df(abunmodel_md) %>%
  transmute(
    tau2_ID      = (sd_ID__Intercept)^2,
    tau2_Species = (sd_Species__Intercept)^2,
    sigma2       = (sigma)^2
  ) %>%
  mutate(
    vbar      = mean(databuncp_ed$se^2),
    var_total = tau2_ID + tau2_Species + sigma2,
    I2_total  = var_total / (var_total + vbar)
  ) %>%
  summarise(
    I2_mean     = mean(I2_total),
    I2_median   = median(I2_total),
    I2_CI_lower = quantile(I2_total, 0.025),
    I2_CI_upper = quantile(I2_total, 0.975)
  )

I2_ab_md
```

## Posterior probability intercept > 0
```{r}
post.samples_ab_md <- as_draws_df(abunmodel_md) %>%
  pull(b_Intercept)

mean(post.samples_ab_md > 0)


```

## Forest plot
```{r}
# Extract posterior draws for each study (ID) and species
study.draws_ab_md <- abunmodel_md %>%
  spread_draws(r_ID[ID, ], r_Species[Species, ], b_Intercept, ndraws = 1000) %>%
  mutate(b_Intercept = r_ID + r_Species + b_Intercept)

# Extract pooled effect (overall intercept)
pooled.effect.draws_ab_md <- abunmodel_md %>%
  spread_draws(b_Intercept, ndraws = 1000) %>%
  mutate(ID = "Pooled Effect")

# Prepare study-level data
study_data_ab_md <- study.draws_ab_md %>%
  mutate(ID = as.character(ID)) %>%
  mutate(ID = str_replace_all(ID, "[.]", " ")) %>%
  mutate(ID = reorder(ID, b_Intercept))

# Combine study and pooled effect
forest.data_ab_md <- bind_rows(study_data_ab_md, pooled.effect.draws_ab_md) %>%
  mutate(ID = as.character(ID)) %>%
  mutate(ID = fct_relevel(ID, "Pooled Effect", after = Inf)) %>%
  mutate(effect_type = ifelse(ID == "Pooled Effect", "Pooled", "Study")) %>%
  ungroup()

# Summarize for plotting points + intervals
forest.data.summary_ab_md <- forest.data_ab_md %>%
  group_by(ID) %>%
  mean_qi(b_Intercept) %>%
  mutate(ID = fct_relevel(ID, "Pooled Effect", after = Inf)) %>%
  mutate(effect_type = ifelse(ID == "Pooled Effect", "Pooled", "Study"))

```


```{r}
# Forest plot
ab.forest.plot_md <- ggplot(forest.data_ab_md, aes(x = b_Intercept, y = ID, fill = effect_type)) +
  geom_vline(xintercept = fixef(abunmodel_md)[1, 3:4], color = "grey", linetype = 2) +
  geom_vline(xintercept = 0, color = "black", size = 0.75) +
  geom_density_ridges(rel_min_height = 0.01, col = NA, scale = 1, alpha = 0.8) +
  geom_pointinterval(
    data = forest.data.summary_ab_md,
    aes(x = b_Intercept, y = ID, xmin = .lower, xmax = .upper, color = effect_type),
    size = 1, orientation = "y"
  ) +
  xlim(-2, 2) +
  geom_text(
    data = mutate_if(forest.data.summary_ab_md, is.numeric, round, 2),
    aes(label = glue("{b_Intercept} [{.lower}, {.upper}]"), y = ID, x = 1.75),
    hjust = "inward", size = 3
  ) +
  scale_y_discrete(
    limits = rev(levels(forest.data_ab_md$ID)),
    labels = NULL,
    breaks = NULL,
    expand = expansion(add = 0.75)
  ) +
  scale_fill_manual(values = c("Study" = "#E69F00", "Pooled" = "#56B4E9")) +
  scale_color_manual(values = c("Study" = "#E69F00", "Pooled" = "#56B4E9")) +
  labs(x = "Effect size (abundance)", y = NULL) +
  my_theme

ab.forest.plot_md

```

```{r}
# Save to PDF
ggsave(
  filename = "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/abun_cp_forest_md.pdf",
  plot = ab.forest.plot_md,
  device = cairo_pdf,
  width = 5,
  height = 8,
  units = "in"
)
```

### Plot for Grand Mean effect size for abundance, genetic diversity and population growth rate
```{r}
# Average effect
out_Lambda_gd <- spread_draws(Lambdamodel_gd, b_Intercept) %>% 
  mutate(study = "Population\ngrowth rate")

# Data frame of summary numbers
out_all_sum_lambda_gd <- group_by(out_Lambda_gd, study) %>% 
  mean_qi(b_Intercept)

# Draw plot
Fig1a_gd <- out_Lambda_gd %>%
  ggplot(aes(x = b_Intercept, y = study, fill = after_stat(x < 0))) +
  labs(x = "", y = "") +
  xlim(-3, 2) +
  geom_vline(xintercept = 0, linewidth = .25, lty = 2) +
  stat_halfeye(.width = c(.95)) +
  scale_fill_manual(values = c("#FFC34D", "#999999")) +
  geom_text(
    data = mutate_if(out_all_sum_lambda_gd, is.numeric, round, 2),
    aes(label = str_glue("{b_Intercept} [{.lower}, {.upper}]"),
        x = -2., y = 1.05),
    hjust = "inward"
  ) +
  my_theme

```

```{r}
# Average effect
out_Lambda_cd <- spread_draws(Lambdamodel_cd, b_Intercept) %>% 
  mutate(study = "Population\ngrowth rate")

# Data frame of summary numbers
out_all_sum_lambda_cd <- group_by(out_Lambda_cd, study) %>% 
  mean_qi(b_Intercept)

# Draw plot
Fig1b_cd <- out_Lambda_cd %>%
  ggplot(aes(x = b_Intercept, y = study, fill = after_stat(x < 0))) +
  labs(x = "", y = "") +
  xlim(-3, 2) +
  geom_vline(xintercept = 0, linewidth = .25, lty = 2) +
  stat_halfeye(.width = c(.95)) +
  scale_fill_manual(values = c("#FFC34D", "#999999")) +
  geom_text(
    data = mutate_if(out_all_sum_lambda_cd, is.numeric, round, 2),
    aes(label = str_glue("{b_Intercept} [{.lower}, {.upper}]"),
        x = -2., y = 1.05),
    hjust = "inward"
  ) +
  my_theme


```


```{r}
# Average effect
out_f_gen_gd <- spread_draws(genamodel, b_Intercept) %>% 
  mutate(study = "Genetic \n diversity")

# Data frame of summary numbers
out_all_sum_gen_gd <- group_by(out_f_gen_gd, study) %>% 
  mean_qi(b_Intercept)

# Draw plot
Fig1c_gd <- out_f_gen_gd %>%   
  ggplot(aes(x = b_Intercept, y = study, fill = after_stat(x < 0))) +
  labs(x = "", y = '') +
  xlim(-2, 2) +
  geom_vline(xintercept = 0, linewidth = .25, lty = 2) +
  stat_halfeye(.width = c(.95)) +
  scale_fill_manual(values = c("#FFC34D", "#999999")) +
  geom_text(
    data = mutate_if(out_all_sum_gen_gd, is.numeric, round, 2),
    aes(label = str_glue("{b_Intercept} [{.lower}, {.upper}]"), x = -1.6),
    hjust = "inward"
  ) +
  my_theme

```

```{r}
# Average effect
out_f_abun_gd <- spread_draws(abunmodel_gd, b_Intercept) %>% 
  mutate(study = "Abundance")

# Data frame of summary numbers
out_all_sum_abun_gd <- group_by(out_f_abun_gd, study) %>% 
  mean_qi(b_Intercept)

# Draw plot
Fig1d_gd <- out_f_abun_gd %>%   
  ggplot(aes(x = b_Intercept, y = study, fill = after_stat(x < 0))) +
  labs(x = "", y = '') +
  xlim(-2, 2) +
  geom_vline(xintercept = 0, linewidth = .25, lty = 2) +
  stat_halfeye(.width = c(.95)) +
  scale_fill_manual(values = c("#FFC34D", "#999999")) +
  geom_text(
    data = mutate_if(out_all_sum_abun_gd, is.numeric, round, 2),
    aes(label = str_glue("{b_Intercept} [{.lower}, {.upper}]"), x = -1.6),
    hjust = "inward"
  ) +
  my_theme

```

```{r}
# Average effect
out_f_abun_ed <- spread_draws(abunmodel_ed, b_Intercept) %>% 
  mutate(study = "Abundance")

# Data frame of summary numbers
out_all_sum_abun_ed <- group_by(out_f_abun_ed, study) %>% 
  mean_qi(b_Intercept)

# Draw plot
Fig1e_ed <- out_f_abun_ed %>%   
  ggplot(aes(x = b_Intercept, y = study, fill = after_stat(x < 0))) +
  labs(x = "", y = '') +
  xlim(-2, 2) +
  geom_vline(xintercept = 0, linewidth = .25, lty = 2) +
  stat_halfeye(.width = c(.95)) +
  scale_fill_manual(values = c("#FFC34D", "#999999")) +
  geom_text(
    data = mutate_if(out_all_sum_abun_ed, is.numeric, round, 2),
    aes(label = str_glue("{b_Intercept} [{.lower}, {.upper}]"), x = -1.6),
    hjust = "inward"
  ) +
  my_theme

```

# Combine all effect draws
```{r}
all_effects <- bind_rows(
  out_Lambda_gd %>% mutate(metric = "Population growth rate (GD)"),
  out_Lambda_cd %>% mutate(metric = "Population growth rate (ED)"),
  out_f_gen_gd %>% mutate(metric = "Genetic diversity (GD)"),
  out_f_abun_gd %>% mutate(metric = "Abundance (GD)"),
  out_f_abun_ed %>% mutate(metric = "Abundance (ED)"),
  out_f_abun_md %>% mutate(metric = "Abundance (MD)")
)
```
# Extract distance metric code
```{r}
# Extract distance metric codes
all_effects <- all_effects %>%
  mutate(
    metric_code = str_extract(metric, "\\((.*)\\)"),
    metric_code = str_remove_all(metric_code, "[()]")
  )

# Define factor levels for y-axis
metric_levels <- c(
  "Population growth rate (GD)",
  "Population growth rate (ED)",
  "Genetic diversity (GD)",
  "Abundance (GD)",
  "Abundance (ED)",
  "Abundance (MD)"
)

all_effects$metric <- factor(all_effects$metric, levels = metric_levels)

#  clean labels for plotting
metric_labels <- c(
  "Population growth rate (GD)" = "Population growth rate",
  "Population growth rate (ED)" = "Population growth rate",
  "Genetic diversity (GD)" = "Genetic diversity",
  "Abundance (GD)" = "Population abundance",
  "Abundance (ED)" = "Population abundance",
  "Abundance (MD)" = "Population abundance"
)

# Summarize posterior draws

all_effects_sum <- all_effects %>%
  group_by(metric, metric_code) %>%
  median_qi(b_Intercept, .width = 0.95)


# # Make sure factor levels match
# all_effects_sum$metric <- factor(all_effects_sum$metric, levels = metric_levels)

```

```{r}
all_effects_sum_sig <- all_effects_sum %>%
  mutate(
    signif = case_when(
      .lower > 0 | .upper < 0 ~ "*",
      TRUE ~ ""
    )
  )

```

```{r}
Fig1<-ggplot(all_effects, aes(x = b_Intercept, y = metric)) +
  stat_halfeye(
  aes(fill = metric_code, color = metric_code),
  .width = 0.95,
  alpha = 0.2,
  .point = median
) +
stat_interval(
  aes(color = metric_code),
  .width = 0.95,
  linewidth = 1
)+
  geom_vline(xintercept = 0, linetype = 2) +
  
  #  Error bars for 95% CI
  geom_errorbarh(
    data = all_effects_sum,
    aes(xmin = .lower, xmax = .upper, y = metric, color = metric_code),
    linewidth = 1, height = 0.2
  ) +

  # Points for posterior median
  geom_point(
    data = all_effects_sum,
    aes(x = b_Intercept, y = metric, fill = metric_code, color = metric_code),
    shape = 21, size = 4, stroke = 1.2
  ) +
  
  # Colors
  scale_fill_manual(
    name = "Distance metric",
    values = c("GD" = "#1b9e77", "ED" = "#d95f02", "MD" = "#7570b3"),
    labels = c("GD" = "Geographic", "ED" = "Euclidean", "MD" = "Mahalanobis")
  ) +
  scale_color_manual(
    name = "Distance metric",
    values = c("GD" = "#1b9e77", "ED" = "#d95f02", "MD" = "#7570b3"),
    labels = c("GD" = "Geographic", "ED" = "Euclidean", "MD" = "Mahalanobis")
  ) +
  
  # Labels and limits
  labs(x = "Effect size", y = "") +
  coord_cartesian(xlim = c(-1.2, 1.2))+
  scale_y_discrete(labels = metric_labels) +
  #   geom_text(
  #   data = mutate_if(all_effects_sum, is.numeric, round, 2),
  #   aes(label = str_glue("{b_Intercept} [{.lower}, {.upper}]"),
  #       x =-1.15),
  #   hjust = 0,
  #   size = 3,
  #   nudge_y = 0.15
  # ) +
  geom_text(
  data = all_effects_sum_sig,
  aes(
    label = signif,
    x = 0.85
  ),
  hjust = 0,
  size = 9,
  nudge_y = -0.04
)+
  # Theme
  theme_test(base_size = 12) +
  theme(
    legend.position = c(0.88, 0.95),
    legend.justification = c(0.5, 1),
    text = element_text(family = "Arial"),
    #text = element_text(color = "black"),
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 12)
  )
```

```{r}
# Save to PDF
ggsave(
  filename = "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/Fig1.pdf",
  plot = Fig1,
  device = cairo_pdf,         # embeds fonts cleanly
  width = 8,                  # in inches (adjust to journal column width)
  height = 6,
  units = "in",
  dpi = 600                   # harmless; ensures any raster parts are high‑res
)

```

## Testing if adding data from Dallas et al , 2017  improve the results

```{r}
databuncp %>% 
  filter(!Author %in% c("Dallas et al.", "Santini et al.","Osocrio-Olivera et al"))->Abundcp_SDM
```

### Grand Mean effect size and significance 
```{r}
library(dplyr)
library(brms)
library(tidyr)
library(ggdist)
library(purrr)

# Unique methods
methods <- unique(databuncp$Method)

# Create an empty list to store results
results_list <- list()

# Loop over methods
for(m in methods){
  
  # Subset data
  dat_sub <- databuncp %>% filter(Method == m)
  
  # Fit model
  model <- brm(
    yi | se(se, sigma = TRUE) ~ 1 + (1 | ID) + (1 | Species),
    data = dat_sub,
    seed = 13,
    control = list(adapt_delta = 0.95),
    silent = TRUE
  )
  
  # Extract posterior draws and add Method
  out <- spread_draws(model, b_Intercept) %>%
    mutate(Method = m,
           study = "Abundance")
  
  # Summarize median ± 95% CI and significance
  out_sum <- out %>%
    group_by(study, Method) %>%
    median_qi(b_Intercept, .width = 0.95) %>%
    ungroup() %>%
    mutate(signif = case_when(
      .lower > 0 | .upper < 0 ~ "*",
      TRUE ~ "ns"
    ))
  
  # Store
  results_list[[m]] <- list(model = model, draws = out, summary = out_sum)
}

```


```{r}
all_methods_summary <- bind_rows(
  results_list[["ED"]]$summary,
  results_list[["GD"]]$summary,
  results_list[["MD"]]$summary
)

```

```{r}
Fig1SDM<-ggplot(all_methods_summary, aes(x = Method, y = b_Intercept, fill = Method)) +
  geom_col(width = 0.6, color = "black") +
  geom_errorbar(aes(ymin = .lower, ymax = .upper), width = 0.2) +
  geom_text(aes(label = signif, y = .upper + 0.05), size = 5, fontface = "bold") +
  geom_hline(yintercept = 0, linetype = 2) +
  # Custom colors for distance metrics
  scale_fill_manual(
    name = "Distance metric",
    values = c("GD" = "#1b9e77", "ED" = "#d95f02", "MD" = "#7570b3"),
    labels = c("GD" = "Geographic", "ED" = "Euclidean", "MD" = "Mahalanobis")
  ) +
  
  theme_test(base_size = 12) +
  labs(x = "Method", y = "Abundance (median ± 95% CI)")+
  theme(legend.position = "none") 


```

```{r}
# Save to PDF
ggsave(
  filename = "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/Fig1SDM.pdf",
  plot = Fig1SDM,
  device = cairo_pdf,         # embeds fonts cleanly
  width = 5,                  # in inches (adjust to journal column width)
  height = 5,
  units = "in",
  dpi = 600                   # harmless; ensures any raster parts are high‑res
)

```

### Predictive check for all the models (abundance, population growth rate, population abundance)
```{r}
multiplot <- function(..., plotlist=NULL, cols=1) {
  require(grid)
  plots <- c(list(...), plotlist)
  numPlots = length(plots)
  
  layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                   ncol = cols, nrow = ceiling(numPlots/cols))
  
  if (numPlots==1) {
    print(plots[[1]])
  } else {
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    for (i in 1:numPlots) {
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

```

```{r}
# print figure
pdf("/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/PPC_gd.pdf",useDingbats = F,height=6,width=6)
multiplot(ppc_growthrate_gd,ppc_genetic_gd,
          ppc_abundance_gd,cols=2)
dev.off()
```

```{r}
# print figure
pdf("/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/PPC_cd.pdf",useDingbats = F,height=6,width=6)
multiplot(ppc_growthrate_cd,ppc_abundance_ed,
          ppc_abundance_md,cols=2)
dev.off()
```

### For supplementary information 
```{r}
all_effects_sum_sup_info <- all_effects %>%
  group_by(metric, metric_code) %>%
  summarise(
    median = median(b_Intercept),
    lower = quantile(b_Intercept, 0.025),
    upper = quantile(b_Intercept, 0.975),
    sd = sd(b_Intercept),
    se = sd / sqrt(n()),  # crude posterior SE
    .groups = "drop"
  )
```

# Detection of publication bias
## Population growth rate 
```{r}
Lambdamodel <- brm(
  yi | se(se,sigma=TRUE) ~ 1 + (1|ID) + (1|Species),
  data = datlambda_gd,
  seed = 13,
  family = gaussian(),
  prior = priors_lb,
  control = list(adapt_delta = 0.99, max_treedepth = 15),
  iter = 3000,
  cores = 4
)
```

```{r}
fittedlambda <- fitted(
  Lambdamodel,
  newdata = datlambda,
  allow_new_levels = TRUE
)
datlambda$meta_resid <- datlambda$yi  - fittedlambda[,1]
datlambda <- datlambda %>% 
  mutate(p_i = 1/se,
         o_i = meta_resid * p_i)
# rm(fittedlambda)
```

egger's regression
```{r, eval =F}
m_egg_lambda <- brm(data = datlambda, o_i ~ p_i, control = list(adapt_delta = 0.999), cores = 4)
```

```{r}
summary(m_egg_lambda)
```

effect size ~ year
```{r, eval =F}
m_year_lambda <- update(Lambdamodel, .~. + Year, newdata = datlambda, cores = 4)
```

```{r}
summary(m_year_lambda)
```

```{r}
Lambdamodel_rm <- rma.mv(
  yi = yi,
  V = vi,
  method = "REML",
  test = "t",
  dfs = "contain",
  random = list(~1 | ID, ~1 | Species),
  data = datlambda
)
```

## Population genetic

```{r}
genamodel <- brm(
  yi | se(se,sigma=TRUE) ~ 1 + (1 | ID) + (1|Species),
  control = list(adapt_delta = 0.9999, max_treedepth = 15),
  prior = priors_gn, iter = 8000, cores = 4,seed = 13,
  data = datgencp_gd
)
```

```{r}
fitted_gen <- fitted(genamodel, datgencp)
datgencp$meta_resid <- datgencp$yi  - fitted_gen[,1]
datgencp <- datgencp %>% 
  mutate(p_i = 1/se,
         o_i = meta_resid * p_i)
```

egger's regression
```{r, eval =F}
m_egg_gen <- brm(data = datgencp, o_i ~ p_i, control = list(adapt_delta = 0.999), cores = 4)
```

```{r}
summary(m_egg_gen)
```

effect size ~ year
```{r, eval =F}
m_year_gen <- update(genamodel, .~. + Year, newdata = datgencp, cores = 4)
```

```{r}
summary(m_year_gen)
```

```{r}
genamodelpb <- rma.mv(yi = yi , V = vi, method = "REML", 
                      test="t", dfs="contain",
                      random=list(~1|ID,  ~1|Species),  data = datgencp)
```

## Population abundance
```{r}
abunmodel <- brm(
  yi | se(se,sigma=TRUE) ~ 1 + (1 | ID) + (1 | Species),
  data = databuncp,seed = 13,
  iter = 4000, warmup = 2000, chains = 4
)
```

```{r}
fitted_abun <- fitted(abunmodel, databuncp)
databuncp$meta_resid <- databuncp$yi  - fitted_abun[,1]
databuncp <- databuncp %>% 
  mutate(p_i = 1/se,
         o_i = meta_resid * p_i)
#rm(fitted_abun)
```

egger's regression
```{r, eval =F}
m_egg_abun <- brm(data =databuncp, o_i ~ p_i, cores = 4)
```

```{r}
summary(m_egg_abun)
```

effect size ~ year
```{r, eval =F}
m_year_abun <- update(abunmodel, .~. + Year, newdata = databuncp, cores = 4)
```

```{r}
summary(m_year_abun)
```

```{r}
abunmodelpb <- rma.mv(yi = yi , V = vi, method = "REML", 
                   test="t", dfs="contain",
                   random=list(~1|ID,  ~1|Species),  data = databuncp)

```

```{r}
# Population growth rate
fitted_pop <- fitted(m_year_lambda, newdata = datlambda, re_formula = NA)
datlambda$pred <- fitted_pop[, "Estimate"]
datlambda$pred_lwr <- fitted_pop[, "Q2.5"]
datlambda$pred_upr <- fitted_pop[, "Q97.5"]

# Genetic diversity
fitted_gen <- fitted(m_year_gen, newdata = datgencp, re_formula = NA)
datgencp$pred <- fitted_gen[, "Estimate"]
datgencp$pred_lwr <- fitted_gen[, "Q2.5"]
datgencp$pred_upr <- fitted_gen[, "Q97.5"]

# Population abundance
fitted_abun <- fitted(m_year_abun, newdata = databuncp, re_formula = NA)
databuncp$pred <- fitted_abun[, "Estimate"]
databuncp$pred_lwr <- fitted_abun[, "Q2.5"]
databuncp$pred_upr <- fitted_abun[, "Q97.5"]

```

## Put all the funnels plot together

```{r}
cairo_pdf("/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/Figure22.pdf", 
          width = 12, height = 8)

# Layout: top row = trends, bottom row = funnel plots
layout_matrix <- matrix(c(1,2,3,
                          4,5,6), nrow=2, byrow=TRUE)
layout(layout_matrix, heights=c(1,1))  # top row taller

par(
  family   = "Helvetica",
  mar      = c(5, 4, 2, 2),
  cex.lab  = 1.2,   # axis labels
  cex.axis = 1.2,   # tick labels
  cex      = 1.0    # base scaling
)

# Define consistent effect size ranges
xlim_lambda <- ylim_lambda <- c(-5, 5)
xlim_gen    <- ylim_gen    <- c(-5, 5)
xlim_abun   <- ylim_abun   <- c(-2, 2)

# -------------------------
# Top row: trend panels
# -------------------------

# Panel A: Lambda
datlambda <- datlambda[order(datlambda$Year), ]
plot(datlambda$Year, datlambda$yi, 
     pch=19, col=adjustcolor("orange2", alpha.f=0.4),
     xlab="Year", ylab="Effect size (population growth rate)",
     xlim=c(min(datlambda$Year), max(datlambda$Year)), ylim=ylim_lambda,xaxt="n")
mtext("A", side=3, line=0.5, at=min(datlambda$Year), cex=1.4, font=1)
axis(1, cex.axis=0.9)
# Panel B: Genetic diversity
datgencp <- datgencp[order(datgencp$Year), ]
plot(datgencp$Year, datgencp$yi, 
     pch=19, col=adjustcolor("orange2", alpha.f=0.4),
     xlab="Year", ylab="Effect size (genetic diversity)",
     xlim=c(min(datgencp$Year), max(datgencp$Year)), ylim=ylim_gen,xaxt="n")
lines(datgencp$Year, datgencp$pred, col="orange2", lwd=3)
polygon(c(datgencp$Year, rev(datgencp$Year)),
        c(datgencp$pred_lwr, rev(datgencp$pred_upr)),
        col=adjustcolor("orange2", alpha.f=0.2), border=NA)
mtext("B", side=3, line=0.5, at=min(datgencp$Year), cex=1.4, font=1)
axis(1, cex.axis=0.9)
# Panel C: Abundance
databuncp <- databuncp[order(databuncp$Year), ]
plot(databuncp$Year, databuncp$yi, 
     pch=19, col=adjustcolor("orange2", alpha.f=0.4),
     xlab="Year", ylab="Effect size (abundance)",
     xlim=c(min(databuncp$Year), max(databuncp$Year)), ylim=ylim_abun,xaxt="n")
mtext("C", side=3, line=0.5, at=min(databuncp$Year), cex=1.4, font=1)
axis(1, cex.axis=0.9)
# -------------------------
# Bottom row: funnel plots
# -------------------------

# Panel D: Lambda funnel
funnel(Lambdamodel_rm, yaxis="seinv", level=c(90,95,99),
       shade=c("white","gray55","gray75"), ylab="Precision (1/SE)",
       xlab="Effect size (population growth rate)", cex.lab=1.2, col="orange2",
       back="white", digits=1, xlim=xlim_lambda, ylim=c(0.01,75))
mtext("D", side=3, line=0.5, at=xlim_lambda[1], cex=1.4, font=1)
box()

# Panel E: Genetic diversity funnel
funnel(genamodelpb, yaxis="seinv", level=c(90,95,99),
       shade=c("white","gray55","gray75"), ylab="", 
       xlab="Effect size (genetic diversity)", cex.lab=1.2,
       col=adjustcolor("orange2", alpha.f=0.4), back="white", digits=1, xlim=xlim_gen, ylim=c(0.01,75))
mtext("E", side=3, line=0.5, at=xlim_gen[1], cex=1.2, font=1)
box()

# Panel F: Abundance funnel
funnel(abunmodelpb, yaxis="seinv", level=c(90,95,99),
       shade=c("white","gray55","gray75"), ylab="", 
       xlab="Effect size (abundance)", cex.lab=1.2,
       col=adjustcolor("orange2", alpha.f=0.4), back="white", digits=1, xlim=xlim_abun, ylim=c(0.01,75))
mtext("F", side=3, line=0.5, at=xlim_abun[1], cex=1.2, font=1)
box()

dev.off()
```

# To what extent can heterogeneity among studies be explained by hypothesized moderators that characterize study outcomes?

## Kingdom
```{r}
datgencp %>%
  group_by(Kingdom) %>%
  summarize(n_species = n_distinct(Species))
```

### Genetic diversity

```{r}
Kingdomg.model <- brms::brm(
  yi|se(se, sigma = TRUE) ~ (Kingdom) + (1|ID) + (1|Species),
  data = datgencp,
  family = gaussian(),
  prior = c(
    prior(normal(0, 1), "b"),
    prior(student_t(3, 0, 5), "sd"),
    prior(student_t(3, 0, 5), "sigma")
  ),
  control = list(adapt_delta = 0.9999, max_treedepth = 15),
                   iter = 5000, cores = 4
)
```

```{r}
summary(Kingdomg.model)
```

### Population abundance
```{r}
# Total number of species
total_species <- databuncp %>% 
  summarize(n_species = n_distinct(Species))

# Number of species per kingdom
species_by_kingdom <- databuncp %>%
  group_by(Kingdom) %>%
  summarize(n_species = n_distinct(Species))

total_species
species_by_kingdom

```

```{r}
Kingdomab.model <- brms::brm(
  yi|se(se, sigma = TRUE) ~ (Kingdom) + (1|ID) + (1|Species),
  data = databuncp,
  family = gaussian(),
  prior = c(
    prior(normal(0, 1), "b"),
    prior(student_t(3, 0, 5), "sd"),
    prior(student_t(3, 0, 5), "sigma")
  ),
  control = list(adapt_delta = 0.9999, max_treedepth = 15),
                   iter = 5000, cores = 4
)
```

```{r}
summary(Kingdomab.model)
```

```{r}
# Genetic diversity
post_gen <- spread_draws(Kingdomg.model, b_Intercept, b_KingdomPlant) %>%
  mutate(
    Animal = b_Intercept,
    Plant = b_Intercept + b_KingdomPlant
  ) %>%
  select(Animal, Plant) %>%
  pivot_longer(cols = c(Animal, Plant), names_to = "Kingdom", values_to = "estimate") %>%
  mutate(Metric = "Genetic diversity")

# Abundance
post_ab <- spread_draws(Kingdomab.model, b_Intercept, b_KingdomPlant) %>%
  mutate(
    Animal = b_Intercept,
    Plant = b_Intercept + b_KingdomPlant
  ) %>%
  select(Animal, Plant) %>%
  pivot_longer(cols = c(Animal, Plant), names_to = "Kingdom", values_to = "estimate") %>%
  mutate(Metric = "Abundance")

# Combine
post_all <- bind_rows(post_gen, post_ab)

```

```{r}
post_sum <- post_all %>%
  group_by(Metric, Kingdom) %>%
  median_qi(estimate, .width = 0.95)

```

### Plot Kingdom vs effect size
```{r}
Figure3 <- ggplot(post_all, aes(x = estimate, y = Metric, fill = Kingdom)) +
  stat_halfeye(
    .width = 0.95,
    alpha = 0.3,
    position = position_dodge(width = 0.2)
  ) +
  geom_errorbarh(
    data = post_sum,
    aes(xmin = .lower, xmax = .upper, y = Metric),
    position = position_dodge(width = 0.2),
    height = 0.1
  ) +
  geom_point(
    data = post_sum,
    aes(x = estimate, y = Metric, fill = Kingdom),
    position = position_dodge(width = 0.2),
    size = 4,
    shape = 21
  ) +
  geom_vline(xintercept = 0, linetype = 2) +
  
  # Distinct colors for Plant vs Animal
  scale_fill_manual(
    values = c(
      "Plant" = "#17becf",   # bright teal
      "Animal" = "#e377c2"   # bright magenta
    )
  ) +
  
  labs(x = "Effect size", y = NULL, fill = "Kingdom") +
  theme_test(base_size = 12) +
  theme(
    legend.position = c(0.88, 0.95),
    legend.justification = c(0.5, 1),
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 12)
  )

```

```{r}
ggsave(
  filename = "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/Figure3.pdf",
  plot = Figure3,
  device = cairo_pdf,         # embeds fonts cleanly
  width = 6,                  # in inches (adjust to journal column width)
  height = 4,
  units = "in",
  dpi = 600                   # harmless; ensures any raster parts are high‑res
)
```

```{r}
# Genetic diversity — keep .draw
post_gen <- spread_draws(Kingdomg.model, b_Intercept, b_KingdomPlant) %>%
  mutate(
    Animal = b_Intercept,
    Plant  = b_Intercept + b_KingdomPlant
  ) %>%
  select(.draw, Animal, Plant) %>%          # <-- keep .draw
  pivot_longer(cols = c(Animal, Plant), names_to = "Kingdom", values_to = "estimate") %>%
  mutate(Metric = "Genetic diversity")

# Abundance — keep .draw
post_ab <- spread_draws(Kingdomab.model, b_Intercept, b_KingdomPlant) %>%
  mutate(
    Animal = b_Intercept,
    Plant  = b_Intercept + b_KingdomPlant
  ) %>%
  select(.draw, Animal, Plant) %>%          # <-- keep .draw
  pivot_longer(cols = c(Animal, Plant), names_to = "Kingdom", values_to = "estimate") %>%
  mutate(Metric = "Abundance")

post_all <- bind_rows(post_gen, post_ab)


# Compute Plant - Animal contrast for each metric
post_contrast <- post_all %>%
  pivot_wider(names_from = Kingdom, values_from = estimate) %>%
  mutate(contrast = Plant - Animal) %>%
  group_by(Metric) %>%
  summarise(
    median_contrast  = median(contrast),
    lower95          = quantile(contrast, 0.025),
    upper95          = quantile(contrast, 0.975),
    p_direction      = mean(contrast > 0),   # probability of direction
    .groups = "drop"
  )

post_contrast
```

```{r}
# --- Compute full contrast distributions for Panel B ---
post_contrast_full <- post_all %>%
  pivot_wider(
    id_cols = c(.draw, Metric),        # <-- explicitly keep Metric as ID
    names_from = Kingdom, 
    values_from = estimate
  ) %>%
  mutate(contrast = Plant - Animal)

dim(post_contrast_full)
post_contrast_sum <- post_contrast_full %>%
  group_by(Metric) %>%
  median_qi(contrast, .width = 0.95)

# --- Panel B: Plant - Animal contrast ---
pB <- ggplot(post_contrast_full, aes(x = contrast, y = Metric)) +
  # stat_halfeye(
  #   fill = "#7f7f7f",
  #   .width = 0.95,
  #   alpha = 0.4
  # ) +
  geom_errorbarh(
    data = post_contrast_sum,
    aes(xmin = .lower, xmax = .upper),
    height = 0.1
  ) +
  geom_point(
    data = post_contrast_sum,
    aes(x = contrast),
    size = 4, shape = 21, fill = "#7f7f7f"
  ) +
  geom_vline(xintercept = 0, linetype = 2, color = "black") +
  labs(x = "Plant \u2212 Animal contrast", y = NULL,
       # title = "(b) Inter-kingdom difference"
       ) +
  theme_test(base_size = 12)

```



```{r}

ggsave(
  filename = "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/Figure3_contrast.pdf",
  plot = pB,
  device = cairo_pdf,
  width = 5, height = 6,
  units = "in", dpi = 600
)
```

```{r}
# Add Metric code based on Metric
supp_table <- post_all %>%
  group_by(Metric, Kingdom) %>%
  summarise(
    Median = median(estimate),
    CI_lower = quantile(estimate, 0.025),
    CI_upper = quantile(estimate, 0.975),
    `P(b>0)` = mean(estimate > 0),
    .groups = "drop"
  ) %>%
  mutate(
    `95% CI` = paste0("[", round(CI_lower, 3), ", ", round(CI_upper, 3), "]"),
    Median = round(Median, 3),
    `P(b>0)` = round(`P(b>0)`, 3),
    `Metric code` = case_when(
      Metric == "Abundance" & Kingdom == "Animal" ~ "ED",
      Metric == "Abundance" & Kingdom == "Plant"  ~ "GD",
      Metric == "Genetic diversity" & Kingdom == "Animal" ~ "ED",
      Metric == "Genetic diversity" & Kingdom == "Plant"  ~ "GD"
    )
  ) %>%
  select(Metric, `Metric code`, Kingdom, Median, `95% CI`, `P(b>0)`)

supp_table

# Print nicely
kable(supp_table, caption = "Posterior median and 95% credible intervals for each metric and kingdom")

```

## Methods
```{r}
table(databuncp$Kingdom, databuncp$Method)
```

```{r}
Kingdomme.model <- brms::brm(
  yi|se(se, sigma = TRUE) ~ (Method) + (1|ID) + (1|Species),
  data = databuncp,
  family = gaussian(),
  prior = c(
    prior(normal(0, 1), "b"),
    prior(student_t(3, 0, 5), "sd"),
    prior(student_t(3, 0, 5), "sigma")
  ),
  control = list(adapt_delta = 0.9999, max_treedepth = 15),
                   iter = 5000, cores = 4
)
```

```{r}
summary(Kingdomme.model)
```

```{r}
#  Create newdata for predictions
newdata <- data.frame(
  Method = c("ED", "GD", "MD"),
  se = 1  # dummy because your model uses |se(se)
)

#  Get posterior draws on response scale
post_all <- add_predicted_draws(
  Kingdomme.model,
  newdata = newdata,
  re_formula = NA,  # ignore random effects
  ndraws = 1000
) %>%
  rename(estimate = .prediction) %>%
  mutate(
    metric = factor(Method, levels = c("ED", "GD", "MD")),
    metric_code = Method
  )

#  Summarize posterior: median + 95% credible interval
post_sum <- post_all %>%
  group_by(metric, metric_code) %>%
  median_qi(.value = estimate, .width = 0.95) %>%
  ungroup()

#  Optional: nicer metric labels
metric_labels <- c("ED" = "Euclidean", "GD" = "Geographic", "MD" = "Mahalanobis")
```

```{r}
#  Plot: Fig1-style
# Plot: Fig1-style
Figure_Method <- ggplot() +  # no default data; set data in each layer
  
  # Half-eye posterior distributions from post_all
  stat_halfeye(
    data = post_all,
    aes(x = estimate, y = metric, fill = metric_code, color = metric_code),
    .width = 0.95,
    alpha = 0.2,
    .point = median
  ) +
  
  # Interval line from post_all
  stat_interval(
    data = post_all,
    aes(x = estimate, y = metric, color = metric_code),
    .width = 0.95,
    linewidth = 1,
    orientation = "horizontal"
  ) +
  
  # Error bars for 95% CI from post_sum
  geom_errorbarh(
    data = post_sum,
    aes(xmin = .lower, xmax = .upper, y = metric, color = metric_code),
    height = 0.2,
    linewidth = 1
  ) +
  
  # Median points from post_sum
  geom_point(
    data = post_sum,
    aes(x = .value, y = metric, fill = metric_code, color = metric_code),
    shape = 21, size = 4, stroke = 1.2
  ) +
  
  # Vertical line at zero
  geom_vline(xintercept = 0, linetype = 2) +
  
  # Colors
  scale_fill_manual(
    values = c("GD" = "#1b9e77", "ED" = "#d95f02", "MD" = "#7570b3")
  ) +
  scale_color_manual(
    values = c("GD" = "#1b9e77", "ED" = "#d95f02", "MD" = "#7570b3")
  ) +
  
  # Axis labels
  labs(x = "Effect size", y = "") +
  scale_y_discrete(labels = metric_labels) +
  
  # Theme
  theme_test(base_size = 12) +
  theme(
    legend.position = "none",
    text = element_text(family = "Arial"),
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 12)
  )

Figure_Method

```


```{r}
# Save to PDF
ggsave(
  filename = "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/Figure_Method.pdf",
  plot = Figure_Method,
  device = cairo_pdf,         # embeds fonts cleanly
  width = 5,                  # in inches (adjust to journal column width)
  height = 5,
  units = "in",
  dpi = 600                   # harmless; ensures any raster parts are high‑res
)

```

```{r}
# -----------------------------------------------------------------------
# CORRECT APPROACH: use posterior_linpred draws aligned by draw index
# linpred_draws() returns the linear predictor (no residual noise added),
# which is what you want for comparing METHOD means — not predictions
# that include residual variance from add_predicted_draws()
# -----------------------------------------------------------------------
library(tidybayes)

post_all_clean <- linpred_draws(
  Kingdomme.model,
  newdata    = newdata,
  re_formula = NA,       # population-level only
  ndraws     = 1000
) %>%
  rename(estimate = .linpred) %>%
  ungroup()              # remove all grouping

# Verify: should have 3000 rows, .draw goes 1–1000 for each Method
post_all_clean %>% count(Method)
# Should be: ED=1000, GD=1000, MD=1000

# ---- Pivot wide: one column per Method, aligned by .draw ----
post_contrast_method <- post_all_clean %>%
  select(.draw, Method, estimate) %>%
  pivot_wider(
    id_cols     = .draw,        # critical: aligns draws correctly
    names_from  = Method,
    values_from = estimate
  ) %>%
  # ---- Compute all three pairwise contrasts ----
  mutate(
    `GD − ED` = GD - ED,
    `MD − ED` = MD - ED,
    `MD − GD` = MD - GD
  ) %>%
  pivot_longer(
    cols      = c(`GD − ED`, `MD − ED`, `MD − GD`),
    names_to  = "Contrast",
    values_to = "contrast"
  ) %>%
  mutate(
    Contrast = factor(Contrast,
                      levels = c("GD − ED", "MD − ED", "MD − GD"))
  )

# ---- Summarise ----
post_contrast_summary_method <- post_contrast_method %>%
  group_by(Contrast) %>%
  median_qi(contrast, .width = 0.95) %>%
  mutate(significant = (.lower > 0 | .upper < 0))

cat("=== Pairwise contrasts (median [95% CI]) ===\n")
post_contrast_summary_method %>%
  mutate(`95% CI` = sprintf("[%.3f, %.3f]", .lower, .upper)) %>%
  select(Contrast, Median = contrast, `95% CI`, significant) %>%
  print()

# ---- Plot ----
Fig_contrast_method <- ggplot(
  post_contrast_method,
  aes(x = contrast, y = Contrast)
) +
  geom_errorbarh(
    data      = post_contrast_summary_method,
    aes(xmin   = .lower,
        xmax   = .upper,
        colour = significant),
    height    = 0.15,
    linewidth = 0.9
  ) +
  geom_point(
    data  = post_contrast_summary_method,
    aes(x = contrast, colour = significant),
    size  = 4
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "black") +
  scale_colour_manual(
    values = c("TRUE" = "#d62728", "FALSE" = "grey40"),
    guide  = "none"
  ) +
  labs(
    x = "Contrast (effect size difference)",
    y = NULL
  ) +
  theme_test(base_size = 12) +
  theme(
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 12)
  )

Fig_contrast_method
```

```{r}
ggsave(
  filename = "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/Fig_contrast_method.pdf",
  plot = Fig_contrast_method,
  device = cairo_pdf,         # embeds fonts cleanly
  width = 6,                  # in inches (adjust to journal column width)
  height = 6,
  units = "in",
  dpi = 600                   # harmless; ensures any raster parts are high‑res
)
```

```{r}
King_method.model <- brms::brm(
  yi|se(se, sigma = TRUE) ~ Method*Kingdom + (1|ID) + (1|Species),
  data = databuncp,
  family = gaussian(),
  prior = c(
    prior(normal(0, 1), "b"),
    prior(student_t(3, 0, 5), "sd"),
    prior(student_t(3, 0, 5), "sigma")
  ),
  control = list(adapt_delta = 0.9999, max_treedepth = 15),
                   iter = 5000, cores = 4
)
```

```{r}
summary(King_method.model)
```

```{r}
# newdata with all Method × Kingdom combinations
newdata <- expand.grid(
  Method = c("ED", "GD", "MD"),
  Kingdom = c("Animal", "Plant"),
  se = 1
)

#  posterior draws
post_all <- add_predicted_draws(
  King_method.model,
  newdata = newdata,
  re_formula = NA,
  ndraws = 1000
) %>%
  rename(estimate = .prediction) %>%
  mutate(
    Method = factor(Method, levels = c("ED","GD","MD")),
    Kingdom = factor(Kingdom, levels = c("Animal","Plant"))
  )

#  posterior summary: median + 95% CI
post_sum <- post_all %>%
  group_by(Method, Kingdom) %>%
  median_qi(.value = estimate, .width = 0.95) %>%
  ungroup()

# Labels
method_labels <- c("ED" = "Euclidean", "GD" = "Geographic", "MD" = "Mahalanobis")
kingdom_colors <- c("Animal" = "#e377c2", "Plant" = "#17becf")  # adjust as desired
```

```{r}
# Plot: side-by-side on same panel
Figure_King_Method <- ggplot() +
  
  # Half-eye posterior distributions
  stat_halfeye(
    data = post_all,
    aes(x = estimate, y = Method, fill = Kingdom, color = Kingdom),
    .width = 0.95,
    alpha = 0.2,
    .point = median,
    position = position_dodge(width = 0.5)
  ) +
  
  # Interval lines
  stat_interval(
    data = post_all,
    aes(x = estimate, y = Method, color = Kingdom),
    .width = 0.95,
    linewidth = 1,
    orientation = "horizontal",
    position = position_dodge(width = 0.5)
  ) +
  
  # Error bars for 95% CI
  geom_errorbarh(
    data = post_sum,
    aes(xmin = .lower, xmax = .upper, y = Method, color = Kingdom),
    height = 0.2,
    linewidth = 1,
    position = position_dodge(width = 0.5)
  ) +
  
  # Median points
  geom_point(
    data = post_sum,
    aes(x = .value, y = Method, fill = Kingdom, color = Kingdom),
    shape = 21, size = 4, stroke = 1.2,
    position = position_dodge(width = 0.5)
  ) +
  
  # Vertical line at zero
  geom_vline(xintercept = 0, linetype = 2) +
  
  # Colors
  scale_fill_manual(values = kingdom_colors) +
  scale_color_manual(values = kingdom_colors) +
  
  # Axis labels
  labs(x = "Effect size ", y = "") +
  scale_y_discrete(labels = method_labels) +
  
  # Theme
  theme_test(base_size = 12) +
  theme(
    legend.position = c(0.88, 0.90),
    text = element_text(family = "Arial"),
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 12)
  )

Figure_King_Method

```

```{r}
ggsave(
  filename = "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/Figure_King_Method.pdf",
  plot = Figure_King_Method,
  device = cairo_pdf,         # embeds fonts cleanly
  width = 6,                  # in inches (adjust to journal column width)
  height = 6,
  units = "in",
  dpi = 600                   # harmless; ensures any raster parts are high‑res
)
```

## Lifespan
### Cleaning the  species traits  [life span] (https://genomics.senescence.info/species/index.html)

```{r}
databuncp %>% 
  filter(Kingdom=="Animal")->databa

# Import ANAge
datage <- read.table(
  "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Data/anage_data.txt",
  header = TRUE,
  sep = "\t",
  stringsAsFactors = FALSE
)

datage_subset <- datage %>%
  mutate(Species = paste0(as.character(Genus), " ", as.character(Species)),
         Order = Order) %>%
  select(Species, Order, Maximum.longevity..yrs., Body.mass..g., Data.quality, Specimen.origin)

# Import PanTHERIA WR05
pan05 <- read.delim(
  "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Data/PanTHERIA_1-0_WR05_Aug2008.txt",
  header = TRUE,
  stringsAsFactors = FALSE
)

pan05_subset <- pan05 %>%
  mutate(
    Species = MSW05_Binomial,
    Order = MSW05_Order,
    Maximum.longevity..yrs. = ifelse(X17.1_MaxLongevity_m > 0, X17.1_MaxLongevity_m / 12, NA),
    Body.mass..g. = ifelse(X5.1_AdultBodyMass_g > 0, X5.1_AdultBodyMass_g, NA),
    Data.quality = NA,
    Specimen.origin = NA
  ) %>%
  select(Species, Order, Maximum.longevity..yrs., Body.mass..g., Data.quality, Specimen.origin)
# Import PanTHERIA WR93
pan93 <- read.delim(
  "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Data/PanTHERIA_1-0_WR93_Aug2008.txt",
  header = TRUE,
  stringsAsFactors = FALSE
)

pan93_subset <- pan93 %>%
  mutate(
    Species = MSW93_Binomial,
    Order = MSW93_Order,
    Maximum.longevity..yrs. = ifelse(X17.1_MaxLongevity_m > 0, X17.1_MaxLongevity_m / 12, NA),
    Body.mass..g. = ifelse(X5.1_AdultBodyMass_g > 0, X5.1_AdultBodyMass_g, NA),
    Data.quality = NA,
    Specimen.origin = NA
  ) %>%
  select(Species, Order, Maximum.longevity..yrs., Body.mass..g., Data.quality, Specimen.origin)

# Import Amniote-Life-History-Database
amniote <- read.csv(
  "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Data/Amniote-Life-History-Database.csv",
  header = TRUE,
  stringsAsFactors = FALSE
)

amniote_clean <- amniote %>%
  mutate(
    adult_body_mass_g = as.numeric(gsub("Quantity\\[|\\]|\"Grams\"|,.*", "", adult_body_mass_g)),
    adult_body_mass_g = ifelse(grepl("Missing", adult_body_mass_g), NA, adult_body_mass_g),
    maximum_longevity_y = as.numeric(gsub("Quantity\\[|\\]|\"Years\"|,.*", "", maximum_longevity_y)),
    maximum_longevity_y = ifelse(grepl("Missing", maximum_longevity_y), NA, maximum_longevity_y)
  )

amniote_subset <- amniote_clean %>%
  mutate(
    Species = paste(genus, species),
    Order = order,
    Maximum.longevity..yrs. = ifelse(maximum_longevity_y > 0, maximum_longevity_y, NA),
    Body.mass..g. = ifelse(adult_body_mass_g > 0, adult_body_mass_g, NA),
    Data.quality = NA,
    Specimen.origin = NA
  ) %>%
  select(Species, Order, Maximum.longevity..yrs., Body.mass..g., Data.quality, Specimen.origin)

# Import bodymass and longevity
bodymass_data_ds <- read.csv(
  "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Data/Bodymass.csv",
  header = TRUE,
  stringsAsFactors = FALSE
)
bodymass_data_ds<-bodymass_data_ds %>% 
  rename(Body.mass..g.=BodyMass,
         Maximum.longevity..yrs.=Longevity)
# Merge all datasets
all_traits <- bind_rows(datage_subset, pan05_subset, pan93_subset, amniote_subset,bodymass_data_ds)

# Getting the species list 
specie_bm_ab <-unique(all_traits$Species)

# Run all names at once
# res_specie_bm_ab <- name_backbone_checklist(name = specie_bm_ab)
library(dplyr)
library(rgbif)

# split species into chunks of 500
library(rgbif)
library(dplyr)

# split into smaller chunks
chunks <- split(specie_bm_ab, ceiling(seq_along(specie_bm_ab)/200))

res_list <- lapply(seq_along(chunks), function(i) {

  cat("Processing chunk", i, "of", length(chunks), "\n")

  out <- name_backbone_checklist(
    name = chunks[[i]],
    bucket_size = 20,
    sleep = 2
  )

  Sys.sleep(2)   # additional pause between batches
  out
})

res_specie_bm_ab <- bind_rows(res_list)
# Extract orders
sp_orders_ab <- as.data.frame(res_specie_bm_ab[,c("canonicalName","order")])

# Match the order with the species in the original dataset. 

all_traits_ab<-left_join(all_traits,sp_orders_ab,by=c("Species"="canonicalName")) %>% 
  dplyr::select(-Order)

# Average by species
all_traits_avg <- all_traits_ab %>%
  # Count how many times each species appears
  group_by(Species) %>%
  mutate(species_count = n()) %>%
  ungroup() %>%
  # For species with duplicates, compute mean per order
  group_by(order, Species) %>%
  summarise(
    Maximum.longevity..yrs. = if (first(species_count) > 1) {
      mean(Maximum.longevity..yrs., na.rm = TRUE)
    } else {
      first(Maximum.longevity..yrs.)
    },
    Body.mass..g. = if (first(species_count) > 1) {
      mean(Body.mass..g., na.rm = TRUE)
    } else {
      first(Body.mass..g.)
    },
    .groups = "drop"
  ) %>%
  # Replace NaN (all NAs) with NA
  mutate(
    Maximum.longevity..yrs. = ifelse(is.nan(Maximum.longevity..yrs.), NA, Maximum.longevity..yrs.),
    Body.mass..g. = ifelse(is.nan(Body.mass..g.), NA, Body.mass..g.)
  )

```

```{r}
# Merge with main dataset (databa)
databuncpLS <- left_join(databa, all_traits_avg, by = "Species",
  relationship = "many-to-many")

# 8. Count species with traits
trait_counts <- databuncpLS %>%
  distinct(Species, .keep_all = TRUE) %>%
  summarise(
    n_species = n(),
    n_longevity = sum(!is.na(Maximum.longevity..yrs.)),
    n_bodymass = sum(!is.na(Body.mass..g.)),
    n_both = sum(!is.na(Maximum.longevity..yrs.) & !is.na(Body.mass..g.))
  )
trait_counts
```

```{r}
# Prepare data
databuncpLS_clean <- databuncpLS %>%
  filter(!is.na(Maximum.longevity..yrs.), !is.na(Body.mass..g.)) %>%
  distinct(Species, .keep_all = TRUE) %>%
  mutate(Body.mass_kg = Body.mass..g. / 1000)  # convert g to kg

# Fit mixed-effects model
fit_mixed_bm_ls <- lmer(log10(Maximum.longevity..yrs.) ~ log10(Body.mass_kg) + (1 | order), data = databuncpLS_clean)
summary(fit_mixed_bm_ls)
# Extract fixed-effect slope and intercept
fixef_coef <- fixef(fit_mixed_bm_ls)
intercept_fixed <- fixef_coef[1]
slope_fixed <- fixef_coef[2]

# Predict overall fixed effect line with CI
pred_df <- data.frame(
  Body.mass_kg = seq(min(databuncpLS_clean$Body.mass_kg), max(databuncpLS_clean$Body.mass_kg), length.out = 100)
)
pred_df$log10Longevity <- predict(fit_mixed_bm_ls, newdata = pred_df, re.form = NA)
pred_df$log10Longevity_upper <- pred_df$log10Longevity + 1.96 * summary(fit_mixed_bm_ls)$sigma
pred_df$log10Longevity_lower <- pred_df$log10Longevity - 1.96 * summary(fit_mixed_bm_ls)$sigma

# Get random-effect intercepts per Order
ranef_df <- ranef(fit_mixed_bm_ls)$order %>%
  tibble::rownames_to_column(var = "order") %>%
  rename(rand_intercept = `(Intercept)`)

# Generate lines for each Order (random intercept)
order_lines <- ranef_df %>%
  mutate(
    slope = slope_fixed
  ) %>%
  tidyr::crossing(Body.mass_kg = pred_df$Body.mass_kg) %>%
  mutate(log10Longevity = slope * log10(Body.mass_kg) + intercept_fixed + rand_intercept)


# # Plot
# # Plot
# Fig_bodymass_longevity <- ggplot(dat_clean, aes(x = Body.mass_kg, y = Maximum.longevity..yrs.)) +
#   geom_point(alpha = 0.6, color = "steelblue") +
#   
#   # Random-intercept lines for each Order
#   geom_line(data = order_lines, aes(x = Body.mass_kg, y = 10^log10Longevity, group = order),
#             color = "gray70", alpha = 0.6) +
#   
#   # Overall fixed-effect line
#   geom_line(data = pred_df, aes(x = Body.mass_kg, y = 10^log10Longevity),
#             color = "darkred", size = 1.2) +
#   
#   # Fixed-effect CI (note inherit.aes = FALSE)
#   geom_ribbon(data = pred_df, aes(x = Body.mass_kg, ymin = 10^log10Longevity_lower, ymax = 10^log10Longevity_upper),
#               fill = "red", alpha = 0.2, inherit.aes = FALSE) +
#   
#   scale_x_log10() +
#   scale_y_log10() +
#   labs(
#     x = "Body mass (kg, log scale)",
#     y = "Maximum longevity (years, log scale)"
#   ) +
#   theme_light()

# Export PDF
#ggsave("bodymass_longevity_mixed_effect_CI.pdf", plot = p, width = 8, height = 6)


```


```{r}
# Fit simple linear regression (log-log)
fit <- lm(log10(Maximum.longevity..yrs.) ~ log10(Body.mass_kg), data =databuncpLS_clean )
summary_fit <- summary(fit)

# Extract slope and p-value
slope <- round(summary_fit$coefficients[2, 1], 3)
pval <- signif(summary_fit$coefficients[2, 4], 3)

# Scatter plot with regression line + 95% CI
p_simple <- ggplot(dat_clean, aes(x = Body.mass_kg, y = Maximum.longevity..yrs.)) +
  geom_point(alpha = 0.6, color = "steelblue") +
  geom_smooth(method = "lm", se = TRUE, color = "darkred") +
  scale_x_log10() +
  scale_y_log10() +
  labs(
    x = "Body mass (kg, log scale)",
    y = "Maximum longevity (years, log scale)"
  ) +
  annotate("text", x = 1, y = 1, label = paste0("Slope = ", slope, "\nP = ", pval), hjust = 0, size = 4) +
  theme_light()

# Export PDF
# ggsave("/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/bodymass_longevity.pdf", plot = p_simple, width = 5, height = 5)

```

```{r}
# Filter data and resolve species names
abunadat <- databuncpLS_clean %>%
  filter(!is.na(Body.mass..g.) & !is.na(yi)) %>%
  mutate(species2 = Species)

# Resolve names via NCBI
Species_bm<-unique(abunadat$species2)

# Function to safely resolve a single species
safe_verify <- function(sp) {
  res <- tryCatch(
    gna_verifier(sp, best_match_only = TRUE, canonical = TRUE),
    error = function(e) return(data.frame(submitted_name = sp, matched_name = NA))
  )
  # If multiple rows returned, take the first (best match)
  if(nrow(res) > 1) res <- res[1, , drop = FALSE]
  return(res)
}

# Apply to all unique species
resolved_list <- lapply(Species_bm, safe_verify)

# Combine results into one data.frame
resolved_ab <- do.call(rbind, resolved_list)

# Inspect results
head(resolved_ab)

abunadat <- abunadat %>%
  left_join(resolved_ab[, c("submittedName", "matchedCanonicalSimple")],
            by = c("species2" = "submittedName")) %>%
  rename(Species_resolved = matchedCanonicalSimple)

# Retrieve taxonomy and build phylogenetic tree
sp_classif_a <- classification(unique(abunadat$Species_resolved), db = "ncbi")
abaphy <- class2tree(sp_classif_a)$phylo
# write.nexus(abaphy, file = "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Data/Phylogeny/abaphy.nex")

```

```{r}
# Clean tree and align with data
abaphy <- read.nexus(
  "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Data/Phylogeny/abaphy.nex"
)
abunadat <- abunadat %>%
  filter(Species %in% abaphy$tip.label)
abunacleanTree <- drop.tip(abaphy, abaphy$tip.label[!abaphy$tip.label %in% abunadat$Species])
A <- ape::vcv.phylo(abunacleanTree)

# Fit phylogenetic mixed model using log10(Body.mass..g.)
bodymass_model <- brm(
  yi | se(se, sigma = TRUE) ~ scale(log10(Body.mass..g.)) + (1|ID) + (1|gr(Species, cov = A)),
  data = abunadat,
  family = gaussian(),
  data2 = list(A = A),
  prior = c(
    prior(normal(0, 1), class = "b"),
    prior(normal(0, 0.5), class = "Intercept"),
    prior(student_t(3, 0, 0.2), class = "sd"),
    prior(student_t(3, 0, 0.2), class = "sigma")
  ),
  iter = 5000,
  cores = 4,
  control = list(adapt_delta = 0.95)  # increase from default 0.8
)

summary(bodymass_model)

# 5. Plot relationship with regression line and CI
abunadat$logBodyMass <- log10(abunadat$Body.mass..g.)

preds <- fitted(bodymass_model, newdata = abunadat, re_formula = NA, summary = TRUE)
abunadat$fit <- preds[, "Estimate"]
abunadat$fit_lower <- preds[, "Q2.5"]
abunadat$fit_upper <- preds[, "Q97.5"]

pdf("/Users/jm200/Library/CloudStorage/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/Figure4_bodymass.pdf", width = 5, height = 4)
ggplot(abunadat, aes(x = logBodyMass, y = yi)) +
  geom_point(color = "orange2", size = 2.5, alpha = 0.6) +
  #geom_ribbon(aes(ymin = fit_lower, ymax = fit_upper), fill = "gray80", alpha = 0.4) +
  #geom_line(aes(y = fit), color = "darkred", size = 1) +
  #geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  labs(x = "Log10 Body mass (g)", y = "Effect size") +
  theme_light(base_size = 10) +
  theme(panel.border = element_rect(fill = NA),
        axis.text.y = element_text(angle = 90, hjust = 0.5))
dev.off()

```

## Pollination mode
### Pollination and population growth rate
```{r}
Pollinationdata <- read.csv("/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Data/Pollinationinfo.csv", header = T)
Pollinationdata$Pollination <- gsub(" ", "",Pollinationdata$Pollination)
datlambdaPl<-left_join(x=datlambda,y=Pollinationdata,by="Species")
datlambdaPl <- datlambdaPl[!apply(datlambdaPl[,c("Pollination", "vi")], 1, anyNA),] 
```

```{r}
sp.lambdaplant <- datlambdaPl[, 5]
spnamesplant <- as.character(sp.lambdaplant)
# Remove the c(...) and quotes, then split by comma
spnamesplant <- gsub('^c\\(|\\)$', '', spnamesplant)           # remove c( and )
spnamesplant <- gsub('\\"', '', spnamesplant)                   # remove inner quotes
spnamesplant <- trimws(unlist(strsplit(spnamesplant, ",")))     # split by comma and trim spaces
spnamesplant
# Retrieve classifications
resolved_namesp <- classification(spnamesplant, db = "ncbi")
resolved_namesp <- unique(resolved_namesp)
# lapply(resolved_namesp, head)
# Building a phylogenetic tree
lambtree <- class2tree(resolved_namesp)
#str(gentree)
(lambtreephy <- lambtree$phylo)
```

```{r}
# Create a data frame of tip labels
spnamesplant <- data.frame(Species = lambtreephy$tip.label)

# Merge with your lambda data (keeping all tips from the tree)
podypdat <- merge(x = spnamesplant, y = datlambdaPl, by = "Species", all.x = TRUE)

# Keep only complete cases
podypdat <- podypdat[complete.cases(podypdat), ]

# Drop tips from the tree that are not in the cleaned dataset
tips_to_drop <- setdiff(lambtreephy$tip.label, podypdat$Species)
lambdcleanTree <- drop.tip(lambtreephy, tips_to_drop)

# 5. Compute the phylogenetic variance-covariance matrix
A <- ape::vcv.phylo(lambdcleanTree)

```

```{r}
datlambdaPl$Species <- factor(datlambdaPl$Species)
# Reorder A to match the levels of Species
# Keep only species present in both the data and the phylo matrix
species_in_both <- intersect(datlambdaPl$Species, rownames(A))
datlambdaPl <- datlambdaPl %>% filter(Species %in% species_in_both)

# Reorder the factor levels to match the covariance matrix
datlambdaPl$Species <- factor(datlambdaPl$Species, levels = species_in_both)

# Reorder A to match the data
A <- A[species_in_both, species_in_both]

Pollination.lambmodel <- brms::brm(
  yi|se(se, sigma = TRUE) ~ (Pollination) + (1|ID) + (1|gr(Species, cov = A)),
  data = datlambdaPl,
  family = gaussian(),
  data2 = list(A = A),
  prior = c(
    prior(normal(0, 1), "b"),
    prior(student_t(3, 0, 5), "sd"),
    prior(student_t(3, 0, 5), "sigma")
  ),
  control = list(adapt_delta = 0.9999, max_treedepth = 15),
                   iter = 5000, cores = 4
)
  
```

```{r}
summary(Pollination.lambmodel)
```

### Pollination and abundance
```{r}
databuncp %>%
  filter(Kingdom == "Plant")->databp
databuncpPl<-left_join(x=databp,y=Pollinationdata,by="Species")
databuncpPl$Pollination <- gsub("Self", "self",databuncpPl$Pollination)
# unique(databuncpPl$Pollination)
databuncpPl <- databuncpPl[!apply(databuncpPl[,c("Pollination", "vi")], 1, anyNA),]
databuncpPl %>%
  group_by(Pollination) %>%
  summarize(n_species = n_distinct(Species))
# unique(databuncpPl$Species[is.na(databuncpPl$Pollination)])
```

```{r,eval=FALSE}
sp.abp <- databuncpPl[, 5]
spnamesabp <- as.character(sp.abp)
species2 <- unique(spnamesabp)
#length(species2)
species2 <- trimws(gsub("\n", "", species2))  # clean up
batch_size <- 50  # keep well under the 100 limit
species_batches <- split(species2, ceiling(seq_along(species2) / batch_size))

# Query each batch
results <- lapply(species_batches, function(batch) {
  taxize::gna_verifier(batch, best_match_only = TRUE, canonical = TRUE)
})

# Combine all results
temp <- bind_rows(results)

dat_abp <- merge(data.frame(species2), temp[,c("submittedName", "matchedCanonicalSimple")], 
               by.x = "species2", by.y = "submittedName", all.x = TRUE)
spnamesabp <-dat_abp[,2]
# spnamesabp <- spnamesabp[complete.cases(spnamesabp)]
# length(spnamesabp)
# Retrieve classifications
spnamesabpf <- classification(spnamesabp, db = "ncbi")
resolved_names_abp <- unique(spnamesabpf)
# Building a phylogenetic tree
abptree <- class2tree(resolved_names_abp)
# str(abptree)
(abpphy <- abptree$phylo)
# length(abpphy$edge.length)
# Save to NEXUS file
# write.nexus(abpphy, file = "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Data/Phylogeny/abpphy.nex")
# Path to your file
tree_file <- "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Data/Phylogeny/abpphy.nex"

# Import the tree
abpphy <- read.nexus(file = tree_file)

```

```{r}
# Create a data frame of tip labels
speciesabunp <- data.frame(Species = abpphy$tip.label)

# Merge with your abundance data (keeping all tips from the tree)
abunpdat <- merge(x = speciesabunp, y = databuncpPl, by = "Species", all.x = TRUE)

# Keep only complete cases
abunpdat <- abunpdat[complete.cases(abunpdat), ]

# Drop tips from the tree that are not in the cleaned dataset
tips_to_drop <- setdiff(abpphy$tip.label, abunpdat$Species)
abunpcleanTree <- drop.tip(abpphy, tips_to_drop)

# Compute the phylogenetic variance-covariance matrix
A <- ape::vcv.phylo(abunpcleanTree)

```

```{r}
Pollination.abunmodel <- brms::brm(
  yi|se(se, sigma = TRUE) ~ (Pollination) + (1|ID) + (1|gr(Species, cov = A)),
  data = abunpdat,
  family = gaussian(),
  data2 = list(A = A),
  prior = c(
    prior(normal(0, 1), "b"),
    prior(student_t(3, 0, 1.5), "sd"),
    prior(student_t(3, 0, 1.5), "sigma")
  ), 
  control = list(adapt_delta = 0.9999, max_treedepth = 15),iter = 5000, cores = 4
)
```

```{r}
summary(Pollination.abunmodel)
```

```{r}
Pollination_method.abunmodel <- brms::brm(
  yi|se(se, sigma = TRUE) ~ (Pollination*Method) + (1|ID) + (1|gr(Species, cov = A)),
  data = abunpdat,
  family = gaussian(),
  data2 = list(A = A),
  prior = c(
    prior(normal(0, 1), "b"),
    prior(student_t(3, 0, 1.5), "sd"),
    prior(student_t(3, 0, 1.5), "sigma")
  ), 
  control = list(adapt_delta = 0.9999, max_treedepth = 15),iter = 5000, cores = 4
)
```

```{r}
summary(Pollination_method.abunmodel)
```


### Pollination and genetic diversity 
```{r}
datgencp %>% 
  filter(Kingdom == "Plant")->datgenp
gencpPL<-left_join(x =datgenp, y =Pollinationdata ,by="Species")
# unique(gencpPL$Species[is.na(gencpPL$Pollination)])
```

```{r}
# lookup table for pollination systems
pollination_gen <- data.frame(
  Species = c("Zostera noltii",
              "Narcissus longispathus",
              "Vaccinium stamineum",
              "Caryocar brasiliense",
              "Arabidopsis thaliana",
              "Clarkia xantiana xantiana",
              "Cirsium canescens"),
  Pollination = c("out",     # Zostera noltii
                  "out",          # Narcissus longispathus
                  "out",     # Vaccinium stamineum
                  "out",          # Caryocar brasiliense
                  "self",           # Arabidopsis thaliana
                  "out",     # Clarkia xantiana xantiana
                  "out")     # Cirsium canescens (likely outcrossed)
)

# merge into your main dataset (replace df with your object name)
gencpPL <- gencpPL %>%
  left_join(pollination_gen, by = "Species") %>%
  mutate(Pollination = coalesce(Pollination.y, Pollination.x)) %>%
  select(-Pollination.x, -Pollination.y)

```


```{r,eval=FALSE}
sp.genp <- gencpPL[, 5]
spnamesp <- as.character(sp.genp)
# Retrieve classifications
resolved_namesp <- classification(spnamesp, db = "ncbi")
resolved_namesp <- unique(resolved_namesp)
# lapply(resolved_namesp, head)
# Building a phylogenetic tree
genptree <- class2tree(resolved_namesp)
#str(gentree)
(genpphy <- genptree$phylo)
# plot(genpphy)
# write.nexus(genpphy, file = "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Data/Phylogeny/genpphy.nex")
gentree_file <- "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Data/Phylogeny/genpphy.nex"

# Import the tree
genpphy <- read.nexus(file = gentree_file)
```

```{r}
# Create a data frame of tip labels from the phylogenetic tree
speciesgen <- data.frame(Species = genpphy$tip.label)

# Merge with your trait data, keeping all species from the tree
genpdat <- merge(x = speciesgen, y = gencpPL, by = "Species", all.x = TRUE)

# Keep only complete cases
genpdat <- genpdat[complete.cases(genpdat), ]

# Drop tips from the tree that are not present in the cleaned dataset
tips_to_drop <- setdiff(genpphy$tip.label, genpdat$Species)
genpcleanTree <- drop.tip(genpphy, tips_to_drop)

# Compute the phylogenetic variance-covariance matrix
A <- ape::vcv.phylo(genpcleanTree)

# Reorder A to match the species order in the data (important for brms)
genpdat$Species <- factor(genpdat$Species, levels = rownames(A))
A <- A[levels(genpdat$Species), levels(genpdat$Species)]

```

```{r}
Pollination.genmodel <- brms::brm(
  yi|se(se, sigma = TRUE) ~ (Pollination) + (1|ID) + (1|gr(Species, cov = A)),
  data = genpdat,
  family = gaussian(),
  data2 = list(A = A),
  prior = c(
    prior(normal(0, 1), "b"),
    prior(student_t(3, 0, 3), "sd"),
    prior(student_t(3, 0, 3), "sigma")
  ),
  control = list(adapt_delta = 0.9999, max_treedepth = 15),
                   iter = 5000, cores = 4
)
  
```

```{r}
summary(Pollination.genmodel)
```

```{r}
Pollination_method.genmodel <- brms::brm(
  yi|se(se, sigma = TRUE) ~ (Pollination*Method) + (1|ID) + (1|gr(Species, cov = A)),
  data = genpdat,
  family = gaussian(),
  data2 = list(A = A),
  prior = c(
    prior(normal(0, 1), "b"),
    prior(student_t(3, 0, 3), "sd"),
    prior(student_t(3, 0, 3), "sigma")
  ),
  control = list(adapt_delta = 0.9999, max_treedepth = 15),
                   iter = 5000, cores = 4
)
  
```

```{r}
summary(Pollination_method.genmodel)
```

```{r}
# --- Genetic diversity ---
post_gen_poll <- spread_draws(Pollination.genmodel, b_Intercept, b_Pollinationself) %>%
  mutate(
    out = b_Intercept,
    self = b_Intercept + b_Pollinationself
  ) %>%
  select(out, self) %>%
  pivot_longer(cols = everything(), names_to = "Pollination", values_to = "estimate") %>%
  mutate(Metric = "Genetic diversity")

# --- Abundance ---
post_ab_poll <- spread_draws(Pollination.abunmodel, b_Intercept, b_Pollinationself) %>%
  mutate(
    out = b_Intercept,
    self = b_Intercept + b_Pollinationself
  ) %>%
  select(out, self) %>%
  pivot_longer(cols = everything(), names_to = "Pollination", values_to = "estimate") %>%
  mutate(Metric = "Abundance")

# --- Population growth rate (Lambda) ---
post_lamb_poll <- spread_draws(Pollination.lambmodel, b_Intercept, b_Pollinationself) %>%
  mutate(
    out = b_Intercept,
    self = b_Intercept + b_Pollinationself
  ) %>%
  select(out, self) %>%
  pivot_longer(cols = everything(), names_to = "Pollination", values_to = "estimate") %>%
  mutate(Metric = "Population growth rate")


```

```{r}
post_all_poll <- bind_rows(post_gen_poll, post_ab_poll, post_lamb_poll)

post_sum_poll <- post_all_poll %>%
  group_by(Metric, Pollination) %>%
  median_qi(estimate, .width = 0.95)
post_all_poll <- post_all_poll %>%
  mutate(Pollination = recode(Pollination,
                              out = "Outcrossed",
                              self = "Self-pollinated"))
post_sum_poll <- post_sum_poll %>%
  mutate(Pollination = recode(Pollination,
                              out = "Outcrossed",
                              self = "Self-pollinated"))

```

```{r}
# Define distinct colors for out/self
poll_colors <- c(
  "Outcrossed" = "#1f77b4",       # blue
  "Self-pollinated" = "#ff7f0e"   # orange
)


Figure_pollination <- ggplot(post_all_poll, aes(x = estimate, y = Metric, fill = Pollination)) +
  stat_halfeye(
    .width = 0.95,
    alpha = 0.3,
    position = position_dodge(width = 0.2)
  ) +
  geom_errorbarh(
    data = post_sum_poll,
    aes(xmin = .lower, xmax = .upper, y = Metric),
    position = position_dodge(width = 0.2),
    height = 0.1
  ) +
  geom_point(
    data = post_sum_poll,
    aes(x = estimate, y = Metric, fill = Pollination),
    position = position_dodge(width = 0.2),
    size = 4,
    shape = 21
  ) +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_fill_manual(values = poll_colors) +
  labs(x = "Effect size", y = NULL, fill = "Pollination") +
  theme_test(base_size = 12) +
  theme(
    legend.position = c(0.88, 0.95),
    legend.justification = c(0.5, 1),
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 12)
  )

Figure_pollination

```


```{r}
ggsave(
  filename = "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/Figure_pollination.pdf",
  plot = Figure_pollination,
  device = cairo_pdf,         # embeds fonts cleanly
  width = 8,                  # in inches (adjust to journal column width)
  height = 6,
  units = "in",
  dpi = 600                   # harmless; ensures any raster parts are high‑res
)
```

```{r}
# Compute posterior contrast: Outcrossed minus Self-pollinated
post_contrast_poll <- bind_rows(post_gen_poll, post_ab_poll, post_lamb_poll) %>%
  group_by(Metric, Pollination) %>%
  mutate(draw = row_number()) %>%   # unique ID per draw within each group
  ungroup() %>%
  pivot_wider(names_from = Pollination, values_from = estimate) %>%
  mutate(
    contrast = out - self,
    Metric   = factor(Metric, levels = c("Genetic diversity",
                                         "Abundance",
                                         "Population growth rate"))
  )

# Verify it worked
str(post_contrast_poll)
head(post_contrast_poll)

# Summarise contrast
post_contrast_summary <- post_contrast_poll %>%
  group_by(Metric) %>%
  median_qi(contrast, .width = 0.95) %>%
  mutate(significant = (.lower > 0 | .upper < 0))  # does 95% CI exclude zero?

# Plot
Fig4contrast<-ggplot(post_contrast_poll, aes(x = contrast, y = Metric)) +
  # stat_halfeye(
  #   .width = c(0.80, 0.95),
  #   fill   = "#1f77b4",
  #   alpha  = 0.4
  # ) +
  geom_errorbarh(
    data = post_contrast_summary,
    aes(xmin = .lower, xmax = .upper,
        colour = significant),
    height = 0.15, linewidth = 0.9
  ) +
  geom_point(
    data = post_contrast_summary,
    aes(x = contrast, colour = significant),
    size = 4
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "black") +
  scale_colour_manual(values = c("TRUE" = "#d62728", "FALSE" = "grey40"),
                      guide = "none") +
  labs(
    x = "Contrast (Outcrossed − Self-pollinated)",
    y = NULL
  ) +
  theme_test(base_size = 12) +
  theme(
    legend.position = c(0.88, 0.95),
    legend.justification = c(0.5, 1),
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 12)
  )
```

```{r}
ggsave(
  filename = "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/Fig4contrast.pdf",
  plot = Fig4contrast,
  device = cairo_pdf,         # embeds fonts cleanly
  width = 6,                  # in inches (adjust to journal column width)
  height = 6,
  units = "in",
  dpi = 600                   # harmless; ensures any raster parts are high‑res
)
```


```{r}
# Get unique species list
species_list <- as.character(unique(datlambda$Species))

# Retrieve all traits available for these species
all_traits <- BIEN_trait_species(species_list)

library(dplyr)
library(tidyr)
library(stringr)

# Function to calculate mode
get_mode <- function(x) {
  ux <- unique(x[!is.na(x)])
  if(length(ux) == 0) return(NA)
  ux[which.max(tabulate(match(x, ux)))]
}

# Clean trait values
all_traits_clean <- all_traits %>%
  mutate(trait_value_clean = str_trim(trait_value))

# Detect numeric traits
numeric_traits <- all_traits_clean %>%
  mutate(is_numeric = !is.na(as.numeric(trait_value_clean))) %>%
  group_by(trait_name) %>%
  summarise(n_numeric = sum(is_numeric, na.rm = TRUE),
            n_total = n(),
            .groups = "drop") %>%
  filter(n_numeric / n_total > 0.5) %>%   # if >50% numeric, treat as numeric
  pull(trait_name)

# Split numeric and categorical traits
traits_numeric <- all_traits_clean %>%
  filter(trait_name %in% numeric_traits) %>%
  mutate(trait_value = as.numeric(trait_value_clean)) %>%
  group_by(scrubbed_species_binomial, trait_name) %>%
  summarise(trait_mean = mean(trait_value, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = trait_name, values_from = trait_mean)

traits_categorical <- all_traits_clean %>%
  filter(!trait_name %in% numeric_traits) %>%
  group_by(scrubbed_species_binomial, trait_name) %>%
  summarise(trait_mode = get_mode(trait_value_clean), .groups = "drop") %>%
  pivot_wider(names_from = trait_name, values_from = trait_mode)
traits_categorical$`whole plant growth form`
# Combine numeric and categorical traits
traits_combined <- traits_numeric %>%
  left_join(traits_categorical, by = "scrubbed_species_binomial")
#view(traits_combined)
# Join with abundance data
datlambda_with_traits <- datlambda %>%
  left_join(traits_combined, by = c("Species" = "scrubbed_species_binomial"))

names(all_traits)
# Pivot to wide format
traits_wide <- all_traits %>%
  select(scrubbed_species_binomial, trait_name, trait_value) %>%   # select relevant columns
  pivot_wider(names_from = trait_name, values_from = trait_value)

# Inspect the wide dataframe
head(traits_wide)
#view(traits_wide)
# Compute mean of each trait per species
traits_mean <- all_traits %>%
  select(scrubbed_species_binomial, trait_name, trait_value) %>%
  # Convert trait_value to numeric if possible
  mutate(trait_value = as.numeric(trait_value)) %>%
  group_by(scrubbed_species_binomial, trait_name) %>%
  summarise(trait_mean = mean(trait_value, na.rm = TRUE), .groups = "drop") %>%
  # Pivot to wide format
  pivot_wider(names_from = trait_name, values_from = trait_mean)

#view(traits_mean)

```


```{r}
lambda_lifeform<-left_join(datlambda,traits_categorical, by=c("Species"="scrubbed_species_binomial"))
lambda_lifeform %>% 
  dplyr::select(ID,Species,`whole plant growth form`,yi,vi,se,Method)->lambda_lifeform_meta
# Create a lookup table for species with NA growth forms
growth_forms <- c(
  "Erythranthe cardinalis" = "Herb",
  "Erythranthe lewisii"    = "Herb",
  "Cypripedium calcicola"  = "Herb",   # terrestrial orchid
  "Ascophyllum nodosum"    = "Alga",
  "Fucus serratus"         = "Alga",
  "Penanthes roanensis"    = "Herb",   # likely Prenanthes roanensis
  "Ranunculus austro-oreganus" = "Herb",
  "Thunbergia atacorensis "="Herb"
)

# Replace NA values in whole.plant.growth.form
lambda_lifeform_meta$`whole plant growth form` <- ifelse(
  is.na(lambda_lifeform_meta$`whole plant growth form`),
  growth_forms[lambda_lifeform_meta$Species],
  lambda_lifeform_meta$`whole plant growth form`
)

lambda_lifeform_meta_clean <- lambda_lifeform_meta %>%
  # rename column
  rename(growth_form = `whole plant growth form`) %>%
  # standardize capitalization
  mutate(growth_form = str_to_title(growth_form)) %>%
  # remove rows with Alga
  filter(growth_form != "Alga")

```

```{r}
# Ensure variables are set correctly
lambda_lifeform_meta_clean <- lambda_lifeform_meta_clean %>%
  mutate(
    Species = factor(Species),
    growth_form = factor(growth_form)  # categorical predictor
  )

# Keep only species present in both the data and the phylo matrix
species_in_both <- intersect(lambda_lifeform_meta_clean$Species, rownames(A))
lambda_lifeform_meta_clean <- lambda_lifeform_meta_clean %>%
  filter(Species %in% species_in_both)

# Reorder Species factor levels to match covariance matrix
lambda_lifeform_meta_clean$Species <- factor(
  lambda_lifeform_meta_clean$Species,
  levels = species_in_both
)

# Reorder phylogenetic covariance matrix
A <- A[species_in_both, species_in_both]

# Run the phylogenetic meta-regression model with growth_form
GrowthForm.lambmodel <- brms::brm(
  yi | se(se, sigma = TRUE) ~ (growth_form) + (1|gr(Species, cov = A)),
  data   = lambda_lifeform_meta_clean,
  family = gaussian(),
  data2  = list(A = A),
  prior  = c(
    prior(normal(0, 1), "b"),          # fixed effects for growth_form
    prior(student_t(3, 0, 5), "sd"),   # random effects
    prior(student_t(3, 0, 5), "sigma") # residual
  ),
  control = list(adapt_delta = 0.9999, max_treedepth = 15),
  iter = 5000, cores = 4
)

```

```{r}
summary(GrowthForm.lambmodel)
```


```{r}
GrowthForm.lambmodel_method <- brms::brm(
  yi | se(se, sigma = TRUE) ~ (growth_form*Method) + (1|gr(Species, cov = A)),
  data   = lambda_lifeform_meta_clean,
  family = gaussian(),
  data2  = list(A = A),
  prior  = c(
    prior(normal(0, 1), "b"),          # fixed effects for growth_form
    prior(student_t(3, 0, 5), "sd"),   # random effects
    prior(student_t(3, 0, 5), "sigma") # residual
  ),
  control = list(adapt_delta = 0.9999, max_treedepth = 15),
  iter = 5000, cores = 4
)
```

```{r}
summary(GrowthForm.lambmodel_method)
```

```{r}
lambda_seed <- datlambda %>%
  left_join(traits_mean, by = c("Species" = "scrubbed_species_binomial")) %>%
  # Rename seed_mass if needed
  rename(seed_mass = `seed mass`) %>%  # only if the column has a space
  # Remove rows with NA in seed_mass
  filter(!is.na(seed_mass))
lambda_seed$Species <- factor(lambda_seed$Species)

```

```{r}
sp_common <- intersect(levels(lambda_seed$Species), rownames(A))
sp_common

```

```{r}
lambda_seed2 <- lambda_seed %>%
  dplyr::filter(Species %in% sp_common) %>%
  droplevels()
A2 <- A[sp_common, sp_common]

# Ensure order matches factor levels
A2 <- A2[levels(lambda_seed2$Species), levels(lambda_seed2$Species)]

all(levels(lambda_seed2$Species) == rownames(A2))
# should return TRUE

```

```{r}
seed_model_phylo <- brm(
  yi | se(se, sigma = TRUE) ~ log10(seed_mass) + (1 | gr(Species, cov = A2)),
  data = lambda_seed2,
  data2 = list(A2 = A2),
  family = gaussian(),
  prior = c(
    prior(normal(0, 1), class = "b"),
    prior(normal(0, 0.5), class = "Intercept"),
    prior(student_t(3, 0, 0.2), class = "sigma")
  ),
  iter = 5000,
  cores = 4,
  control = list(adapt_delta = 0.99)
)

```

```{r}
summary(seed_model_phylo)
```

```{r}
# Fit the seed_mass model
seed_model <- brm(
  yi | se(se, sigma = TRUE) ~ log10(seed_mass),
  data = lambda_seed,
  family = gaussian(),
  prior = c(
    prior(normal(0, 1), class = "b"),
    prior(normal(0, 0.5), class = "Intercept"),
    prior(student_t(3, 0, 0.2), class = "sigma")
  ),
  iter = 5000,
  cores = 4,
  control = list(adapt_delta = 0.99)
)

summary(seed_model)

# 2. Add log10-transformed predictor for plotting
lambda_seed$logSeedMass <- log10(lambda_seed$seed_mass)

# 3. Get fitted values with credible intervals
seed_preds <- fitted(seed_model, newdata = lambda_seed, re_formula = NA, summary = TRUE)
lambda_seed <- lambda_seed %>%
  mutate(fit = seed_preds[, "Estimate"],
         fit_lower = seed_preds[, "Q2.5"],
         fit_upper = seed_preds[, "Q97.5"])


label_lamd_seed <- lambda_seed %>%
  summarize(
    minX = min(log10(seed_mass)),
    maxX = max(log10(seed_mass)),
    lower = min(fit_lower),
    upper = max(fit_upper)
  ) %>%
  mutate(label = str_glue("CI [{round(lower,2)}, {round(upper,2)}]"))
```


```{r}
seed_model_method <- brm(
  yi | se(se, sigma = TRUE) ~ log10(seed_mass)*Method,
  data = lambda_seed,
  family = gaussian(),
  prior = c(
    prior(normal(0, 1), class = "b"),
    prior(normal(0, 0.5), class = "Intercept"),
    prior(student_t(3, 0, 0.2), class = "sigma")
  ),
  iter = 5000,
  cores = 4,
  control = list(adapt_delta = 0.99)
)
```

```{r}
summary(seed_model_method)
```

## Genetic 
```{r}
species_list_gp <- unique(as.character(datgenp$Species))
all_traits_gp <- BIEN_trait_species(species_list_gp)
all_traits_gp_clean <- all_traits_gp %>%
  mutate(trait_value_clean = str_trim(trait_value))
```

```{r}
get_mode <- function(x) {
  ux <- unique(x[!is.na(x)])
  if(length(ux) == 0) return(NA)
  ux[which.max(tabulate(match(x, ux)))]
}

numeric_gp_traits <- all_traits_gp_clean %>%
  mutate(is_numeric = !is.na(as.numeric(trait_value_clean))) %>%
  group_by(trait_name) %>%
  summarise(n_numeric = sum(is_numeric, na.rm = TRUE),
            n_total = n(),
            .groups = "drop") %>%
  filter(n_numeric / n_total > 0.5) %>%
  pull(trait_name)

```

```{r}
# Split numeric and categorical traits
traits_numeric_g <- all_traits_gp_clean %>%
  filter(trait_name %in% numeric_traits) %>%
  mutate(trait_value = as.numeric(trait_value_clean)) %>%
  group_by(scrubbed_species_binomial, trait_name) %>%
  summarise(trait_mean = mean(trait_value, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = trait_name, values_from = trait_mean)
#view(traits_numeric_g)
```

```{r}
traits_gp_categorical <- all_traits_gp_clean %>%
  filter(!trait_name %in% numeric_gp_traits) %>%
  group_by(scrubbed_species_binomial, trait_name) %>%
  summarise(trait_mode = get_mode(trait_value_clean), .groups = "drop") %>%
  pivot_wider(names_from = trait_name, values_from = trait_mode)
# view(traits_gp_categorical)
```

```{r}
gen_lifeform <- left_join(datgenp, traits_gp_categorical, by = c("Species" = "scrubbed_species_binomial"))
gen_lifeform %>% 
  dplyr::select(ID,Species,`whole plant growth form`,yi,vi,se)->gen_lifeform_meta
# view(gen_lifeform_meta)
unique(gen_lifeform_meta$`whole plant growth form`)
```

```{r}
growth_forms <- c(
  "Cymodocea nodosa"          = "Herb",   # aquatic plant
 # "Saccorhiza polyschides"    = "Alga",   # if needed, can exclude later
  "Campanulastrum americanum" = "Herb",
  "Cypripedium"               = "Herb",
  "Genista legionensis"       = "Shrub",
  "Saponaria bellidifolia"    = "Herb",
  "Neotinea maculata"         = "Herb",
  "Viola elatior"             = "Herb",
  "Viola pumila"              = "Herb",
  "Viola stagnina"            = "Herb",
  "Pinus contorta"            = "Tree",
  "Grevillea repens"          = "Shrub",
  "Anthyllis montana"         = "Shrub",
  "Carex magellanica"         = "Herb",
  "Epidendrum fulgens"        = "Herb"
)
```

```{r}
# Replace NA values in whole.plant.growth.form
gen_lifeform$`whole plant growth form` <- ifelse(
  is.na(gen_lifeform$`whole plant growth form`),
  growth_forms[gen_lifeform$Species],
  gen_lifeform$`whole plant growth form`
)

unique(gen_lifeform$`whole plant growth form`)
gen_lifeform_clean <- gen_lifeform %>%
  # rename column
  rename(growth_form = `whole plant growth form`) %>%
  # remove rows with unwanted growth forms and NAs
  filter(!is.na(growth_form) & !tolower(growth_form) %in% c("alga", "aquatic*","aquatic","forb")) %>%
  # standardize capitalization for values
  mutate(
    growth_form = case_when(
      tolower(growth_form) %in% c("tree") ~ "Tree",
      tolower(growth_form) %in% c("shrub") ~ "Shrub",
      tolower(growth_form) %in% c("herb") ~ "Herb",
      TRUE ~ growth_form
    ),
    # convert to factor
    growth_form = factor(growth_form)
  )

unique(gen_lifeform_clean$growth_form)


```

```{r}
# 1. Remove rows with missing growth_form
gen_lifeform_clean <- gen_lifeform_clean %>% filter(!is.na(growth_form))

# 2. Only keep species present in A
species_in_matrix <- intersect(gen_lifeform_clean$Species, rownames(A))
gen_lifeform_clean <- gen_lifeform_clean %>% filter(Species %in% species_in_matrix)

# 3. Reorder Species factor to match covariance matrix
gen_lifeform_clean$Species <- factor(gen_lifeform_clean$Species, levels = species_in_matrix)

# 4. Reorder phylogenetic covariance matrix
A <- A[species_in_matrix, species_in_matrix]

# 5. Fit the phylogenetic meta-regression
growthform_model_gp <- brm(
  yi | se(se, sigma = TRUE) ~ growth_form + (1|ID) + (1 | gr(Species, cov = A)),
  data = gen_lifeform_clean,
  family = gaussian(),
  data2 = list(A = A),
  prior = c(
    prior(normal(0, 1), class = "b"),
    prior(student_t(3, 0, 5), class = "sd"),
    prior(student_t(3, 0, 5), class = "sigma")
  ),
  control = list(adapt_delta = 0.9999, max_treedepth = 15),
  iter = 5000, cores = 4
)


```

```{r}
summary(growthform_model_gp)
```

```{r}
growthform_model_gp_method <- brm(
  yi | se(se, sigma = TRUE) ~ growth_form*Method + (1|ID) + (1 | gr(Species, cov = A)),
  data = gen_lifeform_clean,
  family = gaussian(),
  data2 = list(A = A),
  prior = c(
    prior(normal(0, 1), class = "b"),
    prior(student_t(3, 0, 5), class = "sd"),
    prior(student_t(3, 0, 5), class = "sigma")
  ),
  control = list(adapt_delta = 0.9999, max_treedepth = 15),
  iter = 5000, cores = 4
)

```

```{r}
summary(growthform_model_gp_method)
```

```{r}
traits_numeric_g$`seed mass`
# Compute mean of each trait per species
traits_g <- traits_numeric_g %>%
  select(scrubbed_species_binomial, `seed mass`,) %>%
  # Convert trait_value to numeric if possible
  mutate(seed_mass = as.numeric(`seed mass`)) %>% 
  na.omit()

```


```{r}
gen_seed <- gencpPL %>%
  left_join(traits_g, by = c("Species" = "scrubbed_species_binomial")) %>%
  # Remove rows with NA in seed_mass
  filter(!is.na(seed_mass))

```

```{r}
# Fit the seed_mass model
seed_model_g <- brm(
  yi | se(se, sigma = TRUE) ~ log10(seed_mass),
  data = gen_seed,
  family = gaussian(),
  prior = c(
    prior(normal(0, 1), class = "b"),
    prior(normal(0, 0.5), class = "Intercept"),
    prior(student_t(3, 0, 0.2), class = "sigma")
  ),
  iter = 5000,
  cores = 4,
  control = list(adapt_delta = 0.99)
)

summary(seed_model_g)

# 2. Add log10-transformed predictor for plotting
gen_seed$logSeedMass <- log10(gen_seed$seed_mass)

# 3. Get fitted values with credible intervals
seed_preds_g <- fitted(seed_model_g, newdata = gen_seed, re_formula = NA, summary = TRUE)
gen_seed <- gen_seed %>%
  mutate(fit = seed_preds_g[, "Estimate"],
         fit_lower = seed_preds_g[, "Q2.5"],
         fit_upper = seed_preds_g[, "Q97.5"])


```

```{r}
seed_model_g_method <- brm(
  yi | se(se, sigma = TRUE) ~ log10(seed_mass)*Method,
  data = gen_seed,
  family = gaussian(),
  prior = c(
    prior(normal(0, 1), class = "b"),
    prior(normal(0, 0.5), class = "Intercept"),
    prior(student_t(3, 0, 0.2), class = "sigma")
  ),
  iter = 5000,
  cores = 4,
  control = list(adapt_delta = 0.99)
)

```

```{r}
summary(seed_model_g_method)
```

```{r}
# Prepare trait data
species_list_ab <- unique(as.character(abunpdat$Species))

# Retrieve traits
all_traits_ab <- BIEN_trait_species(species_list_ab)

# Mode function for categorical traits
compute_mode <- function(x) {
  ux <- unique(x[!is.na(x)])
  if(length(ux) == 0) return(NA)
  ux[which.max(tabulate(match(x, ux)))]
}

# Clean trait values
all_traits_clean_ab <- all_traits_ab %>%
  mutate(trait_value_clean = str_trim(trait_value))

# Detect numeric traits
numeric_traits_ab <- all_traits_clean_ab %>%
  mutate(is_numeric = !is.na(as.numeric(trait_value_clean))) %>%
  group_by(trait_name) %>%
  summarise(n_numeric = sum(is_numeric, na.rm = TRUE),
            n_total = n(),
            .groups = "drop") %>%
  filter(n_numeric / n_total > 0.5) %>%
  pull(trait_name)

# Split numeric and categorical
traits_numeric_ab <- all_traits_clean_ab %>%
  filter(trait_name %in% numeric_traits_ab) %>%
  mutate(trait_value = as.numeric(trait_value_clean)) %>%
  group_by(scrubbed_species_binomial, trait_name) %>%
  summarise(trait_mean = mean(trait_value, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = trait_name, values_from = trait_mean)

traits_categorical_ab <- all_traits_clean_ab %>%
  filter(!trait_name %in% numeric_traits_ab) %>%
  group_by(scrubbed_species_binomial, trait_name) %>%
  summarise(trait_mode = compute_mode(trait_value_clean), .groups = "drop") %>%
  pivot_wider(names_from = trait_name, values_from = trait_mode)

# Combine numeric + categorical traits
traits_all_ab <- traits_numeric_ab %>%
  left_join(traits_categorical_ab, by = "scrubbed_species_binomial")

# Merge with abundance data
abund_traits_ab <- abunpdat %>%
  left_join(traits_all_ab, by = c("Species" = "scrubbed_species_binomial")) %>%
  rename(seed_mass_g = `seed mass`) %>%  # rename for clarity
  filter(!is.na(seed_mass_g))

# Fit seed mass meta-regression with Species random effect

```

```{r}
seed_mass_model_ab <- brm(
  yi | se(se, sigma = TRUE) ~ log10(seed_mass_g) + (1 | Species),
  data = abund_traits_ab,
  family = gaussian(),
  prior = c(
    prior(normal(0, 1), class = "b"),          # fixed effect
    prior(normal(0, 0.5), class = "Intercept"), 
    prior(student_t(3, 0, 0.2), class = "sd"), # random effect
    prior(student_t(3, 0, 0.2), class = "sigma") # residual
  ),
  iter = 5000,
  cores = 4,
  control = list(adapt_delta = 0.99)
)

summary(seed_mass_model_ab)

# Prepare fitted values for plotting
abund_traits_ab <- abund_traits_ab %>%
  mutate(logSeedMass = log10(seed_mass_g))

fitted_vals_ab <- fitted(seed_mass_model_ab, newdata = abund_traits_ab, re_formula = NA, summary = TRUE)

abund_traits_ab <- abund_traits_ab %>%
  mutate(fit = fitted_vals_ab[, "Estimate"],
         fit_lower = fitted_vals_ab[, "Q2.5"],
         fit_upper = fitted_vals_ab[, "Q97.5"])


```


```{r}
# Plot
Figure_seedmass<-ggplot(abund_traits_ab, aes(x = logSeedMass, y = yi)) +
  geom_point(color = "orange2", size = 2.5, alpha = 0.6) +
  # geom_ribbon(aes(ymin = fit_lower, ymax = fit_upper), fill = "gray80", alpha = 0.4) +
  # geom_line(aes(y = fit), color = "navy", size = 1) +
  labs(x = "Log10 Seed mass (g)", y = "Effect size ( abundance)") +
  my_theme
```



```{r}
ggsave(
  filename = "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/Figure_seedmass.pdf",
  plot = Figure_seedmass,
  device = cairo_pdf,         # embeds fonts cleanly
  width = 6,                  # in inches (adjust to journal column width)
  height = 5,
  units = "in",
  dpi = 600                   # harmless; ensures any raster parts are high‑res
)
```

```{r}
seed_mass_meth_model_ab <- brm(
  yi | se(se, sigma = TRUE) ~ log10(seed_mass_g)*Method + (1 | Species),
  data = abund_traits_ab,
  family = gaussian(),
  prior = c(
    prior(normal(0, 1), class = "b"),          # fixed effect
    prior(normal(0, 0.5), class = "Intercept"), 
    prior(student_t(3, 0, 0.2), class = "sd"), # random effect
    prior(student_t(3, 0, 0.2), class = "sigma") # residual
  ),
  iter = 5000,
  cores = 4,
  control = list(adapt_delta = 0.99)
)
summary(seed_mass_meth_model_ab)


```


```{r}
# 1. Merge abundance data with categorical traits
lifeform_ab <- left_join(abunpdat, traits_categorical_ab, 
                         by = c("Species" = "scrubbed_species_binomial"))

# Define growth forms for species with NA
growth_forms <- c(
  "Carya ovalis"         = "Tree",
  "Castanea pumila"      = "Tree",
  "Chrysolepis chrysophylla" = "Tree",
  "Crataegus crus-galli" = "Shrub",   # sometimes small tree
  "Ehretia anacua"       = "Tree",
  "Fagus sylvatica"      = "Tree",
  "Fraxinus caroliniana" = "Tree",
  "Fraxinus velutina"    = "Tree",
  "Gleditsia aquatica"   = "Tree",
  "Gordonia lasianthus"  = "Tree",
  "Halesia diptera"      = "Tree",
  "Juniperus ashei"      = "Tree",
  "Juniperus pinchotii"  = "Tree",
  "Pinus remota"         = "Tree",
  "Planera aquatica"     = "Tree",
  "Prunus americana"     = "Tree",
  "Prunus emarginata"    = "Tree",
  "Quercus agrifolia"    = "Tree",
  "Quercus grisea"       = "Tree",
  "Quercus hypoleucoides"= "Tree",
  "Quercus incana"       = "Tree",
  "Quercus marilandica"  = "Tree",
  "Quercus pagoda"       = "Tree",
  "Quercus shumardii"    = "Tree",
  "Quercus similis"      = "Tree",
  "Quercus sinuata"      = "Tree",
  "Sorbus americana"     = "Tree",
  "Sorbus decora"        = "Tree",
  "Torreya californica"  = "Tree",
  "Vernicia fordii"      = "Tree"
)

# Replace NA values in whole plant growth form
lifeform_ab$`whole plant growth form` <- ifelse(
  is.na(lifeform_ab$`whole plant growth form`),
  growth_forms[lifeform_ab$Species],
  lifeform_ab$`whole plant growth form`
)

lifeform_ab_meta <- lifeform_ab %>%
  dplyr::select(ID, Species, `whole plant growth form`, yi, vi, se,Method)

lifeform_ab_meta_clean <- lifeform_ab_meta %>%
  rename(growth_form = `whole plant growth form`) %>%
  mutate(
    growth_form = str_to_title(growth_form),               # standardize capitalization
    growth_form = ifelse(growth_form %in% c("Shrub*", "Shrub"), "Shrub", growth_form)  # combine Shrub* and Shrub
  ) %>%
  filter(
    !is.na(growth_form),                                   # remove NAs
    growth_form != "Alga",                                 # remove Alga
    !grepl("Sclerophyll", growth_form)                     # remove Sclerophyll*
  ) %>%
  mutate(
    growth_form = factor(growth_form)                      # convert back to factor
  )

# Check result

unique(lifeform_ab_meta_clean$growth_form)
species_in_matrix <- intersect(lifeform_ab_meta_clean$Species, rownames(A))
lifeform_ab_meta_clean <- lifeform_ab_meta_clean %>%
  filter(Species %in% species_in_matrix)

# Reorder Species factor to match covariance matrix
lifeform_ab_meta_clean$Species <- factor(
  lifeform_ab_meta_clean$Species,
  levels = species_in_matrix
)

# Reorder phylogenetic covariance matrix
A <- A[species_in_matrix, species_in_matrix]


# Fit phylogenetic meta-regression with growth form
growthform_model_ab <- brm(
  yi | se(se, sigma = TRUE) ~ (growth_form) + (1|ID) + (1 | gr(Species, cov = A)),
  data = lifeform_ab_meta_clean,
  family = gaussian(),
  data2 = list(A = A),
  prior = c(
    prior(normal(0, 1), class = "b"),          # fixed effects
    prior(student_t(3, 0, 5), class = "sd"),   # random effects
    prior(student_t(3, 0, 5), class = "sigma") # residual
  ),
  control = list(adapt_delta = 0.9999, max_treedepth = 15),
  iter = 5000, cores = 4
)
```

```{r}
summary(growthform_model_ab)
get_variables(growthform_model_ab)
```

```{r}
growthform_model_ab_Method <- brm(
  yi | se(se, sigma = TRUE) ~ (growth_form*Method) + (1|ID) + (1 | gr(Species, cov = A)),
  data = lifeform_ab_meta_clean,
  family = gaussian(),
  data2 = list(A = A),
  prior = c(
    prior(normal(0, 1), class = "b"),          # fixed effects
    prior(student_t(3, 0, 5), class = "sd"),   # random effects
    prior(student_t(3, 0, 5), class = "sigma") # residual
  ),
  control = list(adapt_delta = 0.9999, max_treedepth = 15),
  iter = 5000, cores = 4
)
```

```{r}
summary(growthform_model_ab_Method)
```


```{r}
# --- Population growth rate (Lambda) ---
post_lamb_gf <- spread_draws(GrowthForm.lambmodel,
                             b_Intercept,
                             b_growth_formShrub,
                             b_growth_formTree) %>%
  mutate(
    Herb  = b_Intercept,                      # Intercept = Herb
    Shrub = b_Intercept + b_growth_formShrub,
    Tree  = b_Intercept + b_growth_formTree
  ) %>%
  select(Herb, Shrub, Tree) %>%
  pivot_longer(cols = everything(),
               names_to = "GrowthForm",
               values_to = "estimate") %>%
  mutate(Metric = "Population growth rate")

# --- Genetic diversity ---
post_gen_gf <- spread_draws(growthform_model_gp,
                            b_Intercept,
                            b_growth_formShrub,
                            b_growth_formTree) %>%
  mutate(
    Herb  = b_Intercept,
    Shrub = b_Intercept + b_growth_formShrub,
    Tree  = b_Intercept + b_growth_formTree
  ) %>%
  select(Herb, Shrub, Tree) %>%
  pivot_longer(cols = everything(),
               names_to = "GrowthForm",
               values_to = "estimate") %>%
  mutate(Metric = "Genetic diversity")

# --- Abundance ---
post_ab_gf <- spread_draws(growthform_model_ab,
                           b_Intercept,
                           b_growth_formShrub,
                           b_growth_formTree) %>%
  mutate(
    Herb  = b_Intercept,
    Shrub = b_Intercept + b_growth_formShrub,
    Tree  = b_Intercept + b_growth_formTree
  ) %>%
  select(Herb, Shrub, Tree) %>%
  pivot_longer(cols = everything(),
               names_to = "GrowthForm",
               values_to = "estimate") %>%
  mutate(Metric = "Abundance")

# --- Combine all metrics ---
post_all_gf <- bind_rows(post_lamb_gf, post_gen_gf, post_ab_gf)

# --- Summarize with median + 95% CI ---
post_sum_gf <- post_all_gf %>%
  group_by(Metric, GrowthForm) %>%
  median_qi(estimate, .width = 0.95)

# --- Optional: add "ns" labels if 95% CI includes 0 ---
post_sum_gf <- post_sum_gf %>%
  mutate(sig_label = ifelse(.lower <= 0 & .upper >= 0, "ns", ""))

# --- Define colors for growth forms ---
gf_colors <- c(
  "Herb"  = "#F0E442",  # yellow-green
  "Shrub" = "#000000",  
  "Tree"  = "#A6761D"   # soft red/pink that’s different from Animal
)


# --- Plot ---
Figure_growthform <- ggplot(post_all_gf, aes(x = estimate, y = Metric, fill = GrowthForm)) +
  stat_halfeye(
    .width = 0.95,
    alpha = 0.3,
    position = position_dodge(width = 0.5)
  ) +
  geom_errorbarh(
    data = post_sum_gf,
    aes(xmin = .lower, xmax = .upper, y = Metric),
    position = position_dodge(width = 0.5),
    height = 0.1
  ) +
  geom_point(
    data = post_sum_gf,
    aes(x = estimate, y = Metric, fill = GrowthForm),
    position = position_dodge(width = 0.5),
    size = 4,
    shape = 21
  ) +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_fill_manual(values = gf_colors) +
 # scale_y_discrete(expand = c(0.5, 0.5)) +  # increases spacing between metrics
  labs(x = "Effect size", y = NULL, fill = "Growth form") +
  theme_test(base_size = 12) +
  theme(
    legend.position = c(0.88, 0.95),
    legend.justification = c(0.5, 1),
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 12)
  )

Figure_growthform

```

```{r}
ggsave(
  filename = "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/Figure_growthform.pdf",
  plot = Figure_growthform,
  device = cairo_pdf,         # embeds fonts cleanly
  width = 6,                  # in inches (adjust to journal column width)
  height = 6,
  units = "in",
  dpi = 600                   # harmless; ensures any raster parts are high‑res
)
```


```{r}
Figure_combined <- (Figure_pollination | Figure_growthform) +
  plot_annotation(tag_levels = "A") &
  theme(
    plot.tag = element_text(size = 20, face = "plain"),  # bigger size
    plot.tag.position = c(0.05, 1)                       # tweak if needed
  )

Figure_combined
```

```{r}
ggsave(
  filename = "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/Figure_growthform_pollination.pdf",
  plot = Figure_combined,
  device = cairo_pdf,         # embeds fonts cleanly
  width = 14,                  # in inches (adjust to journal column width)
  height = 8,
  units = "in",
  dpi = 600                   # harmless; ensures any raster parts are high‑res
)
```

```{r}
# --- Rebuild with .draw kept ---
post_lamb_gf <- spread_draws(GrowthForm.lambmodel,
                             b_Intercept, b_growth_formShrub, b_growth_formTree) %>%
  mutate(Herb  = b_Intercept,
         Shrub = b_Intercept + b_growth_formShrub,
         Tree  = b_Intercept + b_growth_formTree) %>%
  select(.draw, Herb, Shrub, Tree) %>%
  pivot_longer(cols = c(Herb, Shrub, Tree), names_to = "GrowthForm", values_to = "estimate") %>%
  mutate(Metric = "Population growth rate")

post_gen_gf <- spread_draws(growthform_model_gp,
                            b_Intercept, b_growth_formShrub, b_growth_formTree) %>%
  mutate(Herb  = b_Intercept,
         Shrub = b_Intercept + b_growth_formShrub,
         Tree  = b_Intercept + b_growth_formTree) %>%
  select(.draw, Herb, Shrub, Tree) %>%
  pivot_longer(cols = c(Herb, Shrub, Tree), names_to = "GrowthForm", values_to = "estimate") %>%
  mutate(Metric = "Genetic diversity")

post_ab_gf <- spread_draws(growthform_model_ab,
                           b_Intercept, b_growth_formShrub, b_growth_formTree) %>%
  mutate(Herb  = b_Intercept,
         Shrub = b_Intercept + b_growth_formShrub,
         Tree  = b_Intercept + b_growth_formTree) %>%
  select(.draw, Herb, Shrub, Tree) %>%
  pivot_longer(cols = c(Herb, Shrub, Tree), names_to = "GrowthForm", values_to = "estimate") %>%
  mutate(Metric = "Abundance")

# --- Compute all pairwise contrasts ---
post_contrast_gf <- bind_rows(post_lamb_gf, post_gen_gf, post_ab_gf) %>%
  pivot_wider(
    id_cols     = c(.draw, Metric),
    names_from  = GrowthForm,
    values_from = estimate
  ) %>%
  mutate(
    Tree_vs_Herb  = Tree  - Herb,
    Shrub_vs_Herb = Shrub - Herb,
    Tree_vs_Shrub = Tree  - Shrub
  ) %>%
  select(.draw, Metric, Tree_vs_Herb, Shrub_vs_Herb, Tree_vs_Shrub) %>%
  pivot_longer(cols = c(Tree_vs_Herb, Shrub_vs_Herb, Tree_vs_Shrub),
               names_to  = "Contrast",
               values_to = "contrast") %>%
  mutate(
    Contrast = recode(Contrast,
                      "Tree_vs_Herb"  = "Tree − Herb",
                      "Shrub_vs_Herb" = "Shrub − Herb",
                      "Tree_vs_Shrub" = "Tree − Shrub"),
    Metric = factor(Metric, levels = c("Genetic diversity",
                                       "Abundance",
                                       "Population growth rate"))
  )

# --- Summarise ---
post_contrast_gf_sum <- post_contrast_gf %>%
  group_by(Metric, Contrast) %>%
  median_qi(contrast, .width = 0.95) %>%
  mutate(significant = (.lower > 0 | .upper < 0))

# --- Plot ---
Fig5const<-ggplot(post_contrast_gf, aes(x = contrast, y = Contrast)) +
  # stat_halfeye(
  #   .width = c(0.80, 0.95),
  #   fill   = "#7f7f7f",
  #   alpha  = 0.4
  # ) +
  geom_errorbarh(
    data = post_contrast_gf_sum,
    aes(xmin = .lower, xmax = .upper, colour = significant),
    height = 0.15, linewidth = 0.9
  ) +
  geom_point(
    data = post_contrast_gf_sum,
    aes(x = contrast, colour = significant),
    size = 4
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "black") +
  scale_colour_manual(values = c("TRUE" = "#d62728", "FALSE" = "grey40"),
                      guide  = "none") +
  facet_wrap(~ Metric, ncol = 1) +
  labs(x = "Contrast (effect size difference)", y = NULL) +
  theme_test(base_size = 13) +
  theme(strip.text = element_text(face = "plain"))
```

```{r}
ggsave(
  filename = "/Users/jacobmoutouama/Dropbox/Moutouama Lab/Manuscripts/Meta-analysis-Center-periphery-hypothesis/Figure/Fig5const.pdf",
  plot = Fig5const,
  device = cairo_pdf,         # embeds fonts cleanly
  width = 6,                  # in inches (adjust to journal column width)
  height = 8,
  units = "in",
  dpi = 600                   # harmless; ensures any raster parts are high‑res
)
```

```{r}
sppa_ef<-read_csv("/Users/jacobmoutouama/Desktop/Data/datlambda.csv")
names(sppa_ef)
```

```{r}
Lambdamodel_cd_lat <- brm(
  yi | se(se, sigma = TRUE) ~ Latitude + (1 | ID) + (1 | Species),
  data = sppa_ef,
  family = gaussian(),
  prior = priors_lb,
  seed = 13,
  iter = 3000,
  cores = 4,
  control = list(adapt_delta = 0.99, max_treedepth = 15)
)
```

```{r}
summary(Lambdamodel_cd_lat)
```

